### 2022-08-27 hw1-1/eda_27082022.Rmd
added EDA

---
title: "R Notebook"
output: html_notebook
---

```{r}
#imports
library(cowplot)
library(ggplot2)
library(reshape2)
```

### Credit Card Data Intro EDA

1.  Read data

```{bash}
head -5 ./data\ 2.2/credit_card_data.txt
```

```{r}
org_cc_data <- read.table(file = './data 2.2/credit_card_data-headers.txt', sep = "\t", header = TRUE)
dim(org_cc_data)
class(org_cc_data)
org_cc_data
```

2.  Summary of Data

```{r}
summary(org_cc_data)
#class(summary(org_cc_data))
```

**A1,A9,A10,A12 are binary basis min/max values, rest are continuous**
**For target variable mean is ~45% ==> variable is not imbalanced**

3.  Distribution of Data

```{r}
my_plots <- lapply(names(org_cc_data), function(var_x){
  p <- 
    ggplot(org_cc_data) +
    aes_string(var_x)

  if(var_x %in% list("A1","A9","A10","A12","R1")) {
    p <- p + geom_bar()

  } else {
    p <- p + geom_density()
  }

})

plot_grid(plotlist = my_plots)
#class(summary(org_cc_data))
```
**A15 seems highly skewed towards right side and A11 seems slightly skewed**

4.  Correlation Heat Map
```{r}
cormat <- round(cor(org_cc_data),2)
cormat[upper.tri(cormat)] <- NA
melted_cormat <- melt(cormat)
# plotting the correlation heatmap
library(ggplot2)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2,
                                   fill=value)) +
geom_tile() +
geom_text(aes(Var1, Var2, label = value),
          color = "white", size = 3)
```
If |cor| >0.3 ==> Significant (Let)
R1: A8,A9,A10,A11
A11: A8,A9,A10
A10: A9
A9: A8
A8: A2

### 2022-08-28 hw1-1/ashish/a.svm_27082022.Rmd
added 2.1 part 1,2

---
title: "SVM modelling iterations Ashish"
output: github_document
---

```{r}
library(kernlab)
library(ggplot2)
```

### Read Data

```{r}
org_cc_data <- read.table(file = '../data 2.2/credit_card_data-headers.txt', sep = "\t", header = TRUE)
dim(org_cc_data)
names(org_cc_data)
```

### v0: Vanilla Model and Accuracy

```{r}
model_v0 = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = 10)
#model_v0

acc_func <- function(model) {
   pred_all <- predict(model,org_cc_data[,1:10])
   print (paste("Overall Acc:", round(sum(pred_all == org_cc_data[,11]) * 100 / nrow(org_cc_data),4)))

   pred_1 <- predict(model,org_cc_data[org_cc_data$R1 == 1,1:10])
   print (paste("Acc in 1's:", round(sum(pred_1 == org_cc_data[org_cc_data$R1 == 1,11]) * 100 / nrow(org_cc_data[org_cc_data$R1 == 1,1:10]),4)))

   pred_0 <- predict(model,org_cc_data[org_cc_data$R1 == 0,1:10])
   print (paste("Acc in 0's:", round(sum(pred_0 == org_cc_data[org_cc_data$R1 == 0,11]) * 100 / nrow(org_cc_data[org_cc_data$R1 == 0,1:10]),4)))
}

acc_func(model_v0)
```

### v1: Optimise C: Find best window for linear search

```{r}
C_values <- c(0.0001,0.001,0.0015,0.002,0.005,0.01,0.03)

for (C_i in C_values) {
  print (paste("For C = ",C_i))
  acc_func(ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = C_i))
  print("")
}
```

### v2: vanilla but capping A15

```{r}
hist(org_cc_data$A15, freq = FALSE, breaks = 100)
hist(log(org_cc_data$A15), freq = FALSE, breaks = 100)
#lines(density(x))

cc_data1 = org_cc_data
cc_data1$A15 = log(0.001 + cc_data1$A15)
acc_func(ksvm(x=as.matrix(cc_data1[,1:10]), y=cc_data1[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = 0.007))
```
### Final model and equation

```{r}
modelf = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = 0.0015)
acc_func(modelf)

a <- colSums(modelf@xmatrix[[1]] * modelf@coef[[1]])
a0 <- -modelf@b
a0
```

### Non Linear models: Radial Basis Kernel
```{r}
C_values <- c(0.01,1,10,50,100,1000,2000)

for (C_i in C_values) {
  print (paste("For C = ",C_i))
  modeli = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "rbfdot", C = C_i)
  acc_func(modeli)
  print(modeli@nSV)
}
```

### Non Linear models: Laplace Kernel
```{r}
C_values <- c(0.01,1,10,50,100,1000,2000)

for (C_i in C_values) {
  print (paste("For C = ",C_i))
  modeli = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "laplacedot", C = C_i)
  acc_func(modeli)
  print(modeli@nSV)
}
```

### 2022-08-28 hw1-1/eda_27082022.Rmd
added EDA
added 2.1 part 1,2

@@
 ---
-title: "R Notebook"
-output: html_notebook
+title: "eda"
+output: github_document
 ---
@@
 ```
-If |cor| >0.3 ==> Significant (Let)
+If |cor| > 0.3 ==> Significant (Let)
 R1: A8,A9,A10,A11

### 2022-08-28 hw1-1/submission.Rmd
created submission file

---
title: "Submission HW1"
output:
  html_document:
    df_print: paged
---

# Submission for HW1 \| ISYE 6501 \| Fall 22

-   Ashish Dhiman \| [ashish.dhiman\@gatech.edu](mailto:ashish.dhiman@gatech.edu){.email}

-   Abhinav Arun \|

-   Anshit Verma \|

## Question 2.1

**Describe a situation or problem from your job, everyday life, current events, etc., for which a classification model would be appropriate. List some (up to 5) predictors that you might use.**

### Example 1: Credit Risk Evaluation

Classification models find use cases across a spectrum of Credit Risk functions. A typical example being classifying a particular transaction as risky or otherwise basis which it is either declined or authorized. Here risk implies that the individual might not be able to make the required payments in the future, and therefore transactions made are riskier.

**Objective Function (**$Y$): To classify individual transactions as risky (decline) or not.\
**Predictor Variables (**$X_i$): Some of the key predictor variables given below:

-   ***Past Delinquency***: Delinquency implies that the individual was unable to keep up on his monthly payments previously and missed on his obligated payments. This is a key marker for credit risk, and past delinquent behavior hints towards potential future risk.

-   ***Credit Utilization***: Credit Utilization is defined as the ratio of current balance to overall credit limit accorded to the individual. Suppose a individual has a credit line of \$10,000 and he already has utilized \$ 9k of it. This individual is generally prone to more risk as compared to the individual who only has utilized say \$ 2k of his \$ 10k line.

-   ***Current Debt to Income ratio***: This is a measure of Income to Debt Capacity of the individual. In other words, it stacks up the overall debt obligations of the individual across mortgage, auto loan, credit cards etc., against his total income. If a larger part of an individual's income is directed towards his debt payment, than he/she again might be more susceptible to miss payments in the future and hence is riskier.

-   ***Amount of Transaction***: This is the dollar amount of the transaction. Intuitively a transaction of \$20k is riskier than \$5, since even if the individual misses his payment, the hit taken by the firm is restricted to only \$5.

## Question 2.2

**The files credit_card_data.txt (without headers) and credit_card_data-headers.txt (with headers) contain a dataset with 654 data points, 6 continuous and 4 binary predictor variables. It has anonymized credit card applications with a binary response variable (last column) indicating if the application was positive or negative.**

### Exploratory Data Analysis (EDA)

```{r}
#imports
library(cowplot)
library(ggplot2)
library(reshape2)
```

#### Read Data and Summary

```{r}
org_cc_data <- read.table(file = './data 2.2/credit_card_data-headers.txt', sep = "\t", header = TRUE)
dim(org_cc_data)
head(org_cc_data)
summary(org_cc_data)
```

**A1,A9,A10,A12 are binary basis min/max values, rest are continuous**

**For target variable mean is \~45% ==\> variable is not grossly imbalanced**

#### Data Distribution and Co relation

```{r}
### Data Distribution
my_plots <- lapply(names(org_cc_data), function(var_x){
  p <- 
    ggplot(org_cc_data) +
    aes_string(var_x)

  if(var_x %in% list("A1","A9","A10","A12","R1")) {
    p <- p + geom_bar()

  } else {
    p <- p + geom_density()
  }

})

plot_grid(plotlist = my_plots)

### Corealtion
cormat <- round(cor(org_cc_data),2)
cormat[upper.tri(cormat)] <- NA
melted_cormat <- melt(cormat)
# plotting the correlation heatmap
library(ggplot2)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2,
                                   fill=value)) +
geom_tile() +
geom_text(aes(Var1, Var2, label = value),
          color = "white", size = 3)
```

**If \|cor\| \> 0.3 ==\> Significant, then following pairs show high correlation:**

-   R1: A8,A9,A10,A11

-   A11: A8,A9,A10

-   A10: A9

-   A9: A8

-   A8: A2

## Question 2.2 (part 1)

**Using the support vector machine function ksvm contained in the R package kernlab, find a good classifier for this data. Show the equation of your classifier, and how well it classifies the data points in the full data set. (Don't worry about test/validation data yet; we'll cover that topic soon.)**

```{r}
#imports
library(kernlab)
library(ggplot2)
```

```{r}
dim(org_cc_data)
names(org_cc_data)
```

### v0: Vanilla Model and Accuracy function

```{r}
model_v0 = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = 10)

### Accuracy function: Overall and in each class
acc_func <- function(model) {
   pred_all <- predict(model,org_cc_data[,1:10])
   print (paste("Overall Acc:", round(sum(pred_all == org_cc_data[,11]) * 100 / nrow(org_cc_data),4)))

   pred_1 <- predict(model,org_cc_data[org_cc_data$R1 == 1,1:10])
   print (paste("Acc in 1's:", round(sum(pred_1 == org_cc_data[org_cc_data$R1 == 1,11]) * 100 / nrow(org_cc_data[org_cc_data$R1 == 1,1:10]),4)))

   pred_0 <- predict(model,org_cc_data[org_cc_data$R1 == 0,1:10])
   print (paste("Acc in 0's:", round(sum(pred_0 == org_cc_data[org_cc_data$R1 == 0,11]) * 100 / nrow(org_cc_data[org_cc_data$R1 == 0,1:10]),4)))
}

acc_func(model_v0)
print(paste("#Support Vectors",model_v0@nSV))
```

### v1: Optimise C

```{r}
C_values <- c(0.0001,0.001,0.0015,0.002,0.005,0.01,0.03) #Range identified with hit and trial

for (C_i in C_values) {
  print (paste("For C = ",C_i))
  acc_func(ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = C_i))
  print("")
}
```

**C=0.015 seems the best option, because even with almost equal accuracy, the accuracy in the negative class is higher, where for other C values the accuracy amongst classes is more lopsided**

### 2022-08-28 hw1-1/submission.tex
created submission file

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Submission HW1},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Submission HW1}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

Submission for HW1 \textbar{} ISYE 6501 \textbar{} Fall 22\\
- Ashish Dhiman \textbar{}
\href{mailto:ashish.dhiman@gatech.edu}{\nolinkurl{ashish.dhiman@gatech.edu}}
- Abhinav Arun \textbar{}\\
- Anshit Verma \textbar{}

\hypertarget{question-2.1}{%
\subsection{Question 2.1}\label{question-2.1}}

\textbf{Describe a situation or problem from your job, everyday life,
current events, etc., for which a classification model would be
appropriate. List some (up to 5) predictors that you might use.}

\hypertarget{example-1-credit-risk-evaluation}{%
\paragraph{Example 1: Credit Risk
Evaluation}\label{example-1-credit-risk-evaluation}}

Classification models find use cases across a spectrum of Credit Risk
functions. A typical example being classifying a particular transaction
as risky or otherwise basis which it is either declined or authorized.
Here risk implies that the individual might not be able to make the
required payments in the future, and therefore transactions made are
riskier.

Objective Function (\(Y\)): To classify individual transactions as risky
(decline) or not Predictor Variables (\(X_i\)): Some of the key
predictor variables given below:

Past Delinquency: Delinquency implies that the individual was unable to
keep up on his monthly payments previously and missed on his obligated
payments. This is a key marker for credit risk, and past delinquent
behavior hints towards potential future risk.

Credit Utilization: Credit Utilization is defined as the ratio of
current balance to overall credit limit accorded to the individual.
Suppose a individual has a credit line of \$10,000 and he already has
utilized \$ 9k of it. This individual is generally prone to more risk as
compared to the individual who only has utilized say \$ 2k of his \$ 10k
line.

Current Debt to Income ratio: This is a measure of Income to Debt
Capacity of the individual. In other words, it stacks up the overall
debt obligations of the individual across mortgage, auto loan, credit
cards etc., against his total income. If a larger part of an
individual's income is directed towards his debt payment, than he/she
again might be more susceptible to miss payments in the future and hence
is riskier.

Amount of Transaction: This is the dollar amount of the transaction.
Intuitively a transaction of \$20k is riskier than \$5, since even if
the individual misses his payment, the hit taken by the firm is
restricted to only \$5.

\end{document}

### 2022-08-28 hw1-1/abhinav/knn.Rmd
KNN classification script

---
title: "R Notebook"
output: github_document
---

```{r}
#imports
library(kknn)
```

```{r}
knn_data<-read.table("./hw1-1/data 2.2/credit_card_data-headers.txt",sep='\t',header=TRUE)
dim(knn_data)
```

```{r}
# Plain version with kernel as optimised 
knn_pred<-vector(mode="list",length = dim(knn_data)[1])
pred_table<-data.frame(k_val=numeric(0),acc_val=double(0))
for (k0 in seq(1,nrow(knn_data),by=10)){
  for (i in 1:nrow(knn_data)){
    knn_model_v0<-kknn(R1~.,train = knn_data[-i,],test = knn_data[i,],k=k0,kernel="optimal",scale = TRUE)
    knn_pred[i]<-as.integer(round(fitted(knn_model_v0),digits = 0)==knn_data$R1[i]) 
  }
  percent_accuracy<-round(as.double(sum(unlist(knn_pred)))/nrow(knn_data),digits =2)
  pred_table[nrow(pred_table)+1,]=c(k0,percent_accuracy)
}
```

```{r}
# need to iterate for lower values of K
plot(pred_table)
```

```{r}
knn_pred<-vector(mode="list",length = dim(knn_data)[1])
pred_table<-data.frame(k_val=numeric(0),acc_val=double(0))
for (k0 in seq(1,25,by=3)){
  for (i in 1:nrow(knn_data)){
    knn_model_v0<-kknn(R1~.,train = knn_data[-i,],test = knn_data[i,],k=k0,kernel="optimal",scale = TRUE)
    knn_pred[i]<-as.integer(round(fitted(knn_model_v0),digits = 0)==knn_data$R1[i]) 
  }
  percent_accuracy<-round(as.double(sum(unlist(knn_pred)))/nrow(knn_data),digits =2)
  pred_table[nrow(pred_table)+1,]=c(k0,percent_accuracy)
}
```

### 2022-08-29 hw1-1/submission.Rmd
residual commits

@@
 output:
+  pdf_document: default
   html_document:
@@
 
-# Submission for HW1 \| ISYE 6501 \| Fall 22
+# ISYE 6501 \| Fall 22
 
@@
 
--   Abhinav Arun \|
+-   Abhinav Arun \| [aarun60\@gatech.edu](mailto:aarun60gatech.edu){.email}
 
--   Anshit Verma \|
+-   Anshit Verma \| [averma373\@gatech.edu](mailto:averma373@gatech.edu){.email}
+
+[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
 
 ## Question 2.1
-
-**Describe a situation or problem from your job, everyday life, current events, etc., for which a classification model would be appropriate. List some (up to 5) predictors that you might use.**
 
@@
 
-**The files credit_card_data.txt (without headers) and credit_card_data-headers.txt (with headers) contain a dataset with 654 data points, 6 continuous and 4 binary predictor variables.
-It has anonymized credit card applications with a binary response variable (last column) indicating if the application was positive or negative.**
-
 ### Exploratory Data Analysis (EDA)
@@
 ```
+
+[**#Analysis:**]{style="color:red"}
 
@@
 
+[**#Analysis:**]{style="color:red"}
+
 **If \|cor\| \> 0.3 ==\> Significant, then following pairs show high correlation:**
 
--   R1: A8,A9,A10,A11
+-   R1: A8,A9,A10,A11 (We expect these to show up with high weights in SVM eqn.)
 
@@
 
-## Question 2.2 (part 1)
-
-**Using the support vector machine function ksvm contained in the R package kernlab, find a good classifier for this data.
-Show the equation of your classifier, and how well it classifies the data points in the full data set.
-(Don't worry about test/validation data yet; we'll cover that topic soon.)**
+## Question 2.2 (part 1: kSVM with Linear kernel)
 
@@
 
+[**#Analysis:**]{style="color:red"}
+**There are 190 support vectors, or roughly 30% of the total data points.
+
 ### v1: Optimise C
@@
 
+[**#Analysis:**]{style="color:red"}
+**With increasing C, we weight the cost function more towards misclassification (realtive to margin).
+Therefore intuitively incereasing C, should increase Accuracy albeit at the cost of margin.
+And in the above data points too, similar effect is apparent**
+
 **C=0.015 seems the best option, because even with almost equal accuracy, the accuracy in the negative class is higher, where for other C values the accuracy amongst classes is more lopsided**
+
+### v1: Equation of SVM
+$W = \sum_i(\alpha Y_i X_{support\_vector_i})$$

### 2022-08-30 hw1-1/submission.Rmd
minor analysis notes

@@
 output:
-  pdf_document: default
   html_document:
     df_print: paged
+  pdf_document: default
 ---
@@
 
-[**#Analysis:**]{style="color:red"}
-**There are 190 support vectors, or roughly 30% of the total data points.
+[**#Analysis:**]{style="color:red"} \*\*There are 190 support vectors, or roughly 30% of the total data points.
 
@@
 
-[**#Analysis:**]{style="color:red"}
-**With increasing C, we weight the cost function more towards misclassification (realtive to margin).
+[**#Analysis:**]{style="color:red"} **With increasing C, we weight the cost function more towards misclassification (realtive to margin).
 Therefore intuitively incereasing C, should increase Accuracy albeit at the cost of margin.
@@
 
-### v1: Equation of SVM
-$W = \sum_i(\alpha Y_i X_{support\_vector_i})$$
+### Final SVM Model and it's equation
+
+$W = \sum_i(\alpha Y_i X_{support\_vector_i})$
+
+```{r}
+modelf = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = 0.0015)
+acc_func(modelf)
+
+a <- colSums(modelf@xmatrix[[1]] * modelf@coef[[1]])
+a0 <- -modelf@b
+a
+a0
+```
+[**#Analysis:**]{style="color:red"}
+
+## Question 2.2 (part 2: kSVM with Non Linear kernel)
+
+We have tried 2 kernels:
+
+1.  Radial Basis, and
+
+2.  Laplace
+
+### Non Linear models: Radial Basis Kernel
+```{r}
+C_values <- c(0.01,1,10,50,100,1000,2000)
+
+for (C_i in C_values) {
+  print (paste("For C = ",C_i))
+  modeli = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "rbfdot", C = C_i)
+  acc_func(modeli)
+  print(modeli@nSV)
+}
+```
+
+### Non Linear models: Laplace Kernel
+```{r}
+C_values <- c(0.01,1,10,50,100,1000,2000)
+
+for (C_i in C_values) {
+  print (paste("For C = ",C_i))
+  modeli = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "laplacedot", C = C_i)
+  acc_func(modeli)
+  print(modeli@nSV)
+}
+```

### 2022-08-30 hw1-1/submission.Rmd
submission updated

@@
 Classification models find use cases across a spectrum of Credit Risk functions.
-A typical example being classifying a particular transaction as risky or otherwise basis which it is either declined or authorized.
-Here risk implies that the individual might not be able to make the required payments in the future, and therefore transactions made are riskier.
+A typical example is to a authorize or decline a transaction basis its classification as "risky" or not.
+Here "risky" implies that the individual might not be able to make the required payments in the future, and therefore transactions made are riskier.
 
@@
 ```{r}
-#imports
+#lib imports
 library(cowplot)
@@
 
-**A1,A9,A10,A12 are binary basis min/max values, rest are continuous**
+A1,A9,A10,A12 are binary basis min/max values, rest are continuous
 
-**For target variable mean is \~45% ==\> variable is not grossly imbalanced**
+For target variable mean is \~45% ==\> variable is not grossly imbalanced
 
-#### Data Distribution and Co relation
+```{r}
+org_cc_data1 = org_cc_data
+my_groups <- rep("Negative", nrow(org_cc_data))
+my_groups[org_cc_data$R > 0] <- "Positive"
+org_cc_data1$my_groups <- my_groups
+
+# ggp <- ggplot(org_cc_data1, aes(A3, fill = my_groups)) +  # Seprate density basis label
+#   geom_density(alpha=.34) #Make Transperent
+# ggp  
+```
+
+#### Data Distribution and Pairwise Corelation
 
@@
     p <- p + geom_density()
+  }
+
+})
+
+plot_grid(plotlist = my_plots)
+
+### Seperate Distribution by class labels
+my_plots <- lapply(names(org_cc_data1[,1:11]), function(var_x){
+  p <-
+    ggplot(org_cc_data1[,1:11], aes(var_x, fill = my_groups)) + theme(legend.position="none") +
+    aes_string(var_x)
+
+  if(var_x %in% list("A1","A9","A10","A12","R1")) {
+    p <- p + geom_bar()
+
+  }
+  else {
+    p <- p + geom_density(alpha=.34)
   }
@@
 
-**If \|cor\| \> 0.3 ==\> Significant, then following pairs show high correlation:**
+-   [Distribution:]{.underline}
+
+    From variables distribution it is apparent that a few variables like A15 are highly skewed.
+
+-   [Distribution split basis Class:]{.underline}
+
+    Variables like A9 and A10 exhibit disparity in the label distribution, and hence we expect these variables to show up with stronger weights in SVM equation
+
+-   [Pairwise Corelation:]{.underline}
+
+    If \|cor\| \> 0.3 ==\> Significant, then following pairs show high correlation:
 
@@
 
-[**#Analysis:**]{style="color:red"} \*\*There are 190 support vectors, or roughly 30% of the total data points.
+[**#Analysis:**]{style="color:red"}
 
-### v1: Optimise C
+There are 190 support vectors, or roughly 30% of the total data points, which on first instance seems pretty large.
+It could be though that many points are in a close neighborhood and therefore a significant number of these points lie near the margin lines.
+
+### Model v1: Optimise C
 
 ```{r}
-C_values <- c(0.0001,0.001,0.0015,0.002,0.005,0.01,0.03) #Range identified with hit and trial
+C_values <- c(0.0001,0.001,0.0015,0.002,0.005,0.01,0.03,0.1,0.5,1) #Range identified with hit and trial
 
@@
   acc_func(ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = C_i))
+  print(paste("#Support Vectors",model_v0@nSV))
   print("")
@@
 
-[**#Analysis:**]{style="color:red"} **With increasing C, we weight the cost function more towards misclassification (realtive to margin).
-Therefore intuitively incereasing C, should increase Accuracy albeit at the cost of margin.
-And in the above data points too, similar effect is apparent**
+[**#Analysis:**]{style="color:red"}
 
-**C=0.015 seems the best option, because even with almost equal accuracy, the accuracy in the negative class is higher, where for other C values the accuracy amongst classes is more lopsided**
+With increasing C, we weight the cost function more towards mis-classification (relative to margin).
+Therefore intuitively increasing C, should increase Accuracy albeit at the cost of margin.
+And in the above data points too, similar effect is apparent, as Accuracy first increases and then plateaus.
+
+With Accuracy plateauing 86%, there are two viable C values:
+
+-   C=0.015: For this option provides the overall accuracy is 86% but accuracy in the negative class is higher, whereas for higher C values the accuracy among class groups is more lopsided
+
+-   C=0.001: While here the overall accuracy drops to 83% but accuracy in the negative class is significantly higher at 92%, and given this is a loan application example, it might be better to gear the model more towards the negative class.
 
 ### Final SVM Model and it's equation
-
-$W = \sum_i(\alpha Y_i X_{support\_vector_i})$
 
@@
 acc_func(modelf)
+```
 
+The coefficients of SVM decision boundary is expressed as: $W = \sum_{i \in Supoort\ Vectors}(\alpha_i Y_i X_i)$
+
+```{r}
 a <- colSums(modelf@xmatrix[[1]] * modelf@coef[[1]])
 a0 <- -modelf@b
-a
+sort(a)
 a0
 ```
+
 [**#Analysis:**]{style="color:red"}
+
+As expected (from distribution plots) variables A9, A11, A8, A10 show very significant weights in the decision boundary. Similarly variables like A1,A2 have very low weights, and can also be dropped.
 
@@
 
-We have tried 2 kernels:
+Non Linear kernels work by transposing the actual data into a higher dimensional space. Here we have tried two kernels, both of which transform data to the infinite dimension space.
 
@@
 ### Non Linear models: Radial Basis Kernel
+
 ```{r}
@@
 
+[**#Analysis:**]{style="color:red"}
+
+As seen for Linear Kernels, increase in C, increases accuracy while decreasing number of support vectors.
+Because this is a very complex model, we are able to get very high accuracy, however we can't be sure if the same would generalize to unseen data.
+
 ### Non Linear models: Laplace Kernel
+
 ```{r}
@@
 ```
+
+[**#Analysis:**]{style="color:red"}
+
+Here again we are able to get 100 accuracy, however we can't be sure if the same would generalize to unseen data.
+
+## Question 2.3 (knn Model)

### 2022-08-30 hw1-1/abhinav/knn.Rmd
residual commits
KNN code updated

@@
 title: "R Notebook"
-output: github_document
+output: html_document
 ---
@@
 ```{r}
-# need to iterate for lower values of K
+# iteration over all values of K in increments of 10 to get a probable range where max accuarcy is attained
+plot(pred_table)
+```
+
+```{r}
+# Rectangular(unweighted) kernel as optimised 
+knn_pred<-vector(mode="list",length = dim(knn_data)[1])
+pred_table<-data.frame(k_val=numeric(0),acc_val=double(0))
+for (k0 in seq(1,nrow(knn_data),by=10)){
+  for (i in 1:nrow(knn_data)){
+    knn_model_v0<-kknn(R1~.,train = knn_data[-i,],test = knn_data[i,],k=k0,kernel="rectangular",scale = TRUE)
+    knn_pred[i]<-as.integer(round(fitted(knn_model_v0),digits = 0)==knn_data$R1[i]) 
+  }
+  percent_accuracy<-round(as.double(sum(unlist(knn_pred)))/nrow(knn_data),digits =2)
+  pred_table[nrow(pred_table)+1,]=c(k0,percent_accuracy)
+}
 plot(pred_table)
@@
 pred_table<-data.frame(k_val=numeric(0),acc_val=double(0))
-for (k0 in seq(1,25,by=3)){
+for (k0 in seq(1,30,by=1)){
   for (i in 1:nrow(knn_data)){
-    knn_model_v0<-kknn(R1~.,train = knn_data[-i,],test = knn_data[i,],k=k0,kernel="optimal",scale = TRUE)
+    knn_model_v0<-kknn(R1~.,train = knn_data[-i,],test = knn_data[i,],k=k0,kernel="rectangular",scale = TRUE)
     knn_pred[i]<-as.integer(round(fitted(knn_model_v0),digits = 0)==knn_data$R1[i]) 
@@
 }
+# Iteration over lower values of K
+plot(pred_table)
 ```
+
+```{r}
+#set of all k values where overall accuracy is maximum
+pred_table$k_val[pred_table$acc_val==max(pred_table$acc_val,na.rm=TRUE)]
+```
+
+```{r}
+#anshit to review from here -> loop over k and get a dataframe with k,threshold and auc value, accuracies
+k0=21
+test_pred<-data.frame(pred_val=double(0),act_class=numeric(0))
+for (i in 1:nrow(knn_data)){
+    knn_model_v1<-kknn(R1~.,train = knn_data[-i,],test = knn_data[i,],k=k0,kernel="rectangular",scale = TRUE)
+    knn_pred[i]<-as.double(fitted(knn_model_v1))
+    test_pred[nrow(test_pred)+1,]=c(knn_pred[i],knn_data$R1[i])
+}
+```
+
+```{r}
+accuracy_knn <- function(df){
+  acc_overall = sum(df$act_class == df$pred_class)/nrow(df)
+  df_1 = df[df$act_class > 0,]
+  df_0 = df[df$act_class == 0,]
+  acc_1 = sum(df_1$act_class == df_1$pred_class)/nrow(df_1)
+  acc_0 = sum(df_0$act_class == df_0$pred_class)/nrow(df_0)
+  return (c(acc_overall, acc_1, acc_0))
+}
+```
+
+```{r}
+threshold_auc<-data.frame(threshold_val=double(0),auc_val=double(0),acc_overall=double(0),acc_1=double(0),acc_0=double(0))
+for(t0 in seq(0.2,0.8,by=0.02)){
+  f0<-function(x,t=t0){
+    ifelse(x>t,1,0)
+  }
+  test_pred$pred_class<-apply(data.frame(test_pred$pred_val),1,f0)
+  # function to get overall , 0's and 1's accuracy
+  # accuracy_knn(test_pred)
+  auc_val<-auc(test_pred$act_class,test_pred$pred_class)
+  threshold_auc[nrow(threshold_auc)+1,]=c(t0,auc_val,accuracy_knn(test_pred))
+}
+```
+
+```{r}
+head(threshold_auc)
+accuracy_knn(test_pred)
+```
+
+```{r}
+#roc_score=roc(test_pred$act_class,test_pred$pred_class)
+#plot(roc_score ,main ="ROC curve --KNN ")
+print(paste("Max AUC score :" ,round(max(threshold_auc$auc_val),2)))
+print(paste("Threshold values for which max AUC score is attained", threshold_auc$threshold_val[threshold_auc$auc_val==max(threshold_auc$auc_val,na.rm=TRUE)]))
+```

### 2022-08-30 hw1-1/submission.Rmd
added part3; vanilla model

@@
 library(reshape2)
+library(kknn)
 ```
@@
 ## Question 2.3 (knn Model)
+
+### Vanilla knn model
+
+We implement vanilla knn model here, with rectangular kernel and majority voting (i.e probability threshold \>0.5) among neighbors to decide predicted label.
+
+```{r}
+# Rectangular(unweighted) kernel as optimised 
+knn_pred <- vector(mode="list",length = dim(knn_data)[1])
+pred_table <- data.frame(k_val=numeric(0),acc_val=double(0)) #dataframe to hold accuracy with k
+for (k0 in seq(1,nrow(knn_data),by=25)){ #To loop over k values
+  for (i in 1:nrow(knn_data)){ #To loop over indvidual data points
+    knn_model_v0<-kknn(R1~.,train = knn_data[-i,],test = knn_data[i,],k=k0,kernel="rectangular",scale = TRUE)
+    knn_pred[i]<-as.integer(round(fitted(knn_model_v0),digits = 0)==knn_data$R1[i]) #majority voting > 0.5
+  }
+  percent_accuracy<-round(as.double(sum(unlist(knn_pred)))/nrow(knn_data),digits =2)
+  pred_table[nrow(pred_table)+1,]=c(k0,percent_accuracy)
+}
+plot(pred_table)
+```
+
+[**#Analysis:**]{style="color:red"}
+
+The graph above here plots Accuracy on y-axis vs 'k'.
+As we increase the k, Accuracy is mostly static and after a certain threshold, it starts dropping almost linearly to 55%, where it then stabilizes.
+The 55% number here is intuitive since with as many nearest neighbors as the #data points, the prediction decision basically boils down to assigning the majority label in the dataset, i.e.
+Negative.
+Therefore, for 55% the negative class, we receive a correct prediction while the rest 45% Positive class is misclassified.
+
+### Optimisation over 'k' and probability threshold
+
+[**#Analysis:**]{style="color:red"}

### 2022-08-31 hw1-1/submission.Rmd
added abhinav ques 2.1

--- 2022-08-30 hw1-1/submission.Rmd
@@
 
+### Example 2: Readmission classification in Healthcare analytics
+
+An example of classification problem in Healthcare domain is the classification of whether a patient will get readmitted or not within 30 days of being discharged from a acute hospital setting.
+Having a good predictive model of the same can help us to curb the healthcare expenditure as the estimated average cost of readmission is more than \$10k.
+
+**Objective Function (**$Y$): To classify whether a readmission will happen within 30 days of the Inpatient discharge for a given patient.\
+**Predictor Variables (**$X_i$): Some of the key predictor variables given below:
+
+-   ***Length of stay*** : The period for which a patient stays in the Inpatient setting is a key predictor of readmission.
+A detailed analysis of the same shows that readmission probability is high for both low length of stay (when a patient is discharged earlier than expected without proper treatment) and even when length of stay is larger than a given threshold (because of patient acquiring healthcare associated infections over a prolonged period of stay).
+
+-   ***Discharge Disposition*** : The place where a patient is discharged also a good predictor of whether a person will get readmitted or not within 30 days .For some of the diagnosis categories , if a patient is discharged to home in place of a Post acute care setting , the readmission rate is higher.
+
+-   ***ED Utilization*** : The number of emergency department(ED) visits of a patient within a rolling 1 year is a good predictor of the risk profile of the patient and in turn it is a good predictor of whether a readmission will happen or not.
+
+-   ***\# of changes in medications/procedures*** : The number of medication changes is also a good variable for predicting readmission .
+Higher \# of medication changes during hospital stay is correlated with the occurence of a patient being readmitted.
+
 ## Question 2.2
@@
 ## Question 2.3 (knn Model)
+
+###

### 2022-08-31 hw1-1/submission.Rmd

@@
 
-###
+### Vanilla knn model
+
+We implement vanilla knn model here, with rectangular kernel and majority voting (i.e probability threshold \>0.5) among neighbors to decide predicted label.
+
+```{r}
+# Rectangular(unweighted) kernel as optimised 
+knn_pred <- vector(mode="list",length = dim(knn_data)[1])
+pred_table <- data.frame(k_val=numeric(0),acc_val=double(0)) #dataframe to hold accuracy with k
+for (k0 in seq(1,nrow(knn_data),by=25)){ #To loop over k values
+  for (i in 1:nrow(knn_data)){ #To loop over indvidual data points
+    knn_model_v0<-kknn(R1~.,train = knn_data[-i,],test = knn_data[i,],k=k0,kernel="rectangular",scale = TRUE)
+    knn_pred[i]<-as.integer(round(fitted(knn_model_v0),digits = 0)==knn_data$R1[i]) #majority voting > 0.5
+  }
+  percent_accuracy<-round(as.double(sum(unlist(knn_pred)))/nrow(knn_data),digits =2)
+  pred_table[nrow(pred_table)+1,]=c(k0,percent_accuracy)
+}
+plot(pred_table)
+```
+
+[**#Analysis:**]{style="color:red"}
+
+The graph above here plots Accuracy on y-axis vs 'k'.
+As we increase the k, Accuracy is mostly static and after a certain threshold, it starts dropping almost linearly to 55%, where it then stabilizes.
+The 55% number here is intuitive since with as many nearest neighbors as the #data points, the prediction decision basically boils down to assigning the majority label in the dataset, i.e.
+Negative.
+Therefore, for 55% the negative class, we receive a correct prediction while the rest 45% Positive class is misclassified.
+
+### Optimisation over 'k' and probability threshold
+
+[**#Analysis:**]{style="color:red"}

### 2022-08-31 hw1-1/abhinav/knn.Rmd
Final KNN model updated

@@
 library(kknn)
+library(pROC)
 ```
@@
 ```{r}
-knn_data<-read.table("./hw1-1/data 2.2/credit_card_data-headers.txt",sep='\t',header=TRUE)
+knn_data<-read.table("../data 2.2/credit_card_data-headers.txt",sep='\t',header=TRUE)
 dim(knn_data)
@@
 ```{r}
-#anshit to review from here -> loop over k and get a dataframe with k,threshold and auc value, accuracies
-k0=21
-test_pred<-data.frame(pred_val=double(0),act_class=numeric(0))
-for (i in 1:nrow(knn_data)){
-    knn_model_v1<-kknn(R1~.,train = knn_data[-i,],test = knn_data[i,],k=k0,kernel="rectangular",scale = TRUE)
-    knn_pred[i]<-as.double(fitted(knn_model_v1))
-    test_pred[nrow(test_pred)+1,]=c(knn_pred[i],knn_data$R1[i])
-}
-```
-
-```{r}
+# Accuracy Function
 accuracy_knn <- function(df){
@@
 ```{r}
-threshold_auc<-data.frame(threshold_val=double(0),auc_val=double(0),acc_overall=double(0),acc_1=double(0),acc_0=double(0))
-for(t0 in seq(0.2,0.8,by=0.02)){
+# Threshold Function
 f0<-function(x,t=t0){
   ifelse(x>t,1,0)
-  }
-  test_pred$pred_class<-apply(data.frame(test_pred$pred_val),1,f0)
-  # function to get overall , 0's and 1's accuracy
-  # accuracy_knn(test_pred)
-  auc_val<-auc(test_pred$act_class,test_pred$pred_class)
-  threshold_auc[nrow(threshold_auc)+1,]=c(t0,auc_val,accuracy_knn(test_pred))
 }
@@
 ```{r}
-head(threshold_auc)
-accuracy_knn(test_pred)
+# Data frame for storing values
+threshold_auc<-data.frame(k=double(0), threshold_val=double(0), auc_val=double(0), acc_overall=double(0), acc_1=double(0), acc_0=double(0))
+
+# Loop iterating over values of K for findind best value of K
+for(k in seq(20, 50, by=5)){
+  test_pred<-data.frame(pred_val=double(0),act_class=numeric(0))
+  for (i in 1:nrow(knn_data)){
+      knn_model_v1<-kknn(R1~., train = knn_data[-i,], test = knn_data[i,], k=k, kernel="rectangular", scale = TRUE)
+      knn_pred[i]<-as.double(fitted(knn_model_v1))
+      test_pred[nrow(test_pred)+1,]=c(knn_pred[i],knn_data$R1[i])
+  }
+
+  # Iterating over values of threshold to determine the threshold which maximizes AUC score and hence the better separation between the two classes
+  for(t0 in seq(0.2,0.8,by=0.02)){
+    test_pred$pred_class<-apply(data.frame(test_pred$pred_val),1,f0)
+    auc_val<-auc(test_pred$act_class,test_pred$pred_class)
+    print(auc_val)
+    print(c(k, t0, auc_val, accuracy_knn(test_pred)))
+    threshold_auc[nrow(threshold_auc)+1,]=c(k, t0,auc_val,accuracy_knn(test_pred)[1], accuracy_knn(test_pred)[2], accuracy_knn(test_pred)[3])
+  }
+}
 ```
@@
 ```{r}
-#roc_score=roc(test_pred$act_class,test_pred$pred_class)
-#plot(roc_score ,main ="ROC curve --KNN ")
-print(paste("Max AUC score :" ,round(max(threshold_auc$auc_val),2)))
-print(paste("Threshold values for which max AUC score is attained", threshold_auc$threshold_val[threshold_auc$auc_val==max(threshold_auc$auc_val,na.rm=TRUE)]))
+threshold_auc
 ```
+
+```{r}
+# Final step to determine the highest value of AUC for every K
+final_frame<-data.frame(k=double(0), threshold_val=double(0), auc_val=double(0), acc_overall=double(0), acc_1=double(0), acc_0=double(0))
+for(k_i in seq(20, 50, by=5)){
+  print(k_i)
+  data_i = threshold_auc[threshold_auc$k == k_i,]
+  print (data_i[which.max(data_i$auc_val),])
+  final_frame[nrow(final_frame)+1] = c(data_i[which.max(data_i$auc_val),])
+}
+```
+
+```{r}
+final_frame
+```

### 2022-08-31 hw1-1/anshit/HW1.Rmd

# Assignment 1

### Question 2.2

*Exploratory Data Analysis*

```{r}
library(cowplot)
library(ggplot2)
library(reshape2)
library(kernlab)
```

```{r}
data <- read.table('../data 2.2/credit_card_data-headers.txt', header = TRUE, sep = "\t", dec = ".")
data
```

```{r}
summary(data)
```

```{r}
my_plots = lapply(colnames(data), function(x){
  p = ggplot(data) + aes_string(x)

  if(x %in% list("A1", "A9", "A10", "A12", "R1")){
    p = p + geom_bar()
  } else {
    p = p + geom_density()
  }
})

plot_grid(plotlist = my_plots)
```

**Correlation Matrix**

```{r}
cor_mat = round(cor(data), 2)
cor_mat[upper.tri(cor_mat)] = NA
melted_mat = melt(cor_mat)
ggplot(data = melted_mat, aes(x=Var1, y=Var2, fill=value)) + geom_tile() + 
  geom_text(aes(Var1, Var2, label=value), color="white", size=3)
```
*Modelling*

```{r}
# call ksvm.  Vanilladot is a simple linear kernel.
model <- ksvm(as.matrix(data[,1:10]),data[,11],type="C-svc",kernel="vanilladot",C=0.015,scaled=TRUE)
model
```

```{r}
#predict using model
ypred <- predict(model, as.matrix(data[,1:10]))
table(data[,11], ypred)
```

```{r}
### Accuracy function: Overall and in each class
acc_func <- function(model) {
   pred_all <- predict(model,org_cc_data[,1:10])
   print (paste("Overall Acc:", round(sum(pred_all == org_cc_data[,11]) * 100 / nrow(org_cc_data),4)))

   pred_1 <- predict(model,org_cc_data[org_cc_data$R1 == 1,1:10])
   print (paste("Acc in 1's:", round(sum(pred_1 == org_cc_data[org_cc_data$R1 == 1,11]) * 100 / nrow(org_cc_data[org_cc_data$R1 == 1,1:10]),4)))

   pred_0 <- predict(model,org_cc_data[org_cc_data$R1 == 0,1:10])
   print (paste("Acc in 0's:", round(sum(pred_0 == org_cc_data[org_cc_data$R1 == 0,11]) * 100 / nrow(org_cc_data[org_cc_data$R1 == 0,1:10]),4)))

   print(paste("#Support Vectors",model_v0@nSV))
}
```

```{r}
model_v0 = ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = 10)

acc_func(model_v0)
```

```{r}
C_values <- c(0.0001,0.001,0.0015,0.002,0.005,0.01,0.03) #Range identified with hit and trial

for (C_i in C_values) {
  print (paste("For C = ",C_i))
  acc_func(ksvm(x=as.matrix(org_cc_data[,1:10]), y=org_cc_data[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C = C_i))
  print("")
}
```

**Hypertuning Parameters**

```{r}
#Variables not seeming significant -> A1, A12, A14, A15

new_parameters = data[, c('A1', 'A2', 'A3', 'A8', 'A9', 'A10', 'A11', 'A12', 'A14', 'A15', 'R1')]
head(new_parameters)

```

```{r}
column_range = 1:10
x_train = new_parameters[,column_range]

model_v1 = ksvm(x=as.matrix(x_train), y=new_parameters[,11], scaled =TRUE, type = "C-svc",kernel = "vanilladot", C =0.0015)

### Accuracy function: Overall and in each class
acc_func <- function(model) {
   pred_all <- predict(model, x_train)
   print (paste("Overall Acc:", round(sum(pred_all == new_parameters[,11]) * 100 / nrow(new_parameters),4)))

   pred_1 <- predict(model,new_parameters[new_parameters$R1 == 1,column_range])
   print (paste("Acc in 1's:", round(sum(pred_1 == new_parameters[new_parameters$R1 == 1,11]) * 100 / nrow(new_parameters[new_parameters$R1 == 1, column_range]),4)))

   pred_0 <- predict(model,new_parameters[new_parameters$R1 == 0,column_range])
   print (paste("Acc in 0's:", round(sum(pred_0 == new_parameters[new_parameters$R1 == 0,11]) * 100 / nrow(new_parameters[new_parameters$R1 == 0, column_range]),4)))

   print(paste("#Support Vectors",model_v1@nSV))
}

acc_func(model_v1)
```

### 2022-08-31 hw1-1/submission.Rmd
final knot for submission

@@
 # Rectangular(unweighted) kernel as optimised 
-knn_pred <- vector(mode="list",length = dim(knn_data)[1])
+knn_pred <- vector(mode="list",length = dim(org_cc_data)[1])
 pred_table <- data.frame(k_val=numeric(0),acc_val=double(0)) #dataframe to hold accuracy with k
-for (k0 in seq(1,nrow(knn_data),by=25)){ #To loop over k values
-  for (i in 1:nrow(knn_data)){ #To loop over indvidual data points
-    knn_model_v0<-kknn(R1~.,train = knn_data[-i,],test = knn_data[i,],k=k0,kernel="rectangular",scale = TRUE)
-    knn_pred[i]<-as.integer(round(fitted(knn_model_v0),digits = 0)==knn_data$R1[i]) #majority voting > 0.5
+for (k0 in seq(1,nrow(org_cc_data),by=25)){ #To loop over k values
+  for (i in 1:nrow(org_cc_data)){ #To loop over indvidual data points
+    knn_model_v0<-kknn(R1~.,train = org_cc_data[-i,],test = org_cc_data[i,],k=k0,kernel="rectangular",scale = TRUE)
+    knn_pred[i]<-as.integer(round(fitted(knn_model_v0),digits = 0)==org_cc_data$R1[i]) #majority voting > 0.5
   }
-  percent_accuracy<-round(as.double(sum(unlist(knn_pred)))/nrow(knn_data),digits =2)
+  percent_accuracy<-round(as.double(sum(unlist(knn_pred)))/nrow(org_cc_data),digits =2)
   pred_table[nrow(pred_table)+1,]=c(k0,percent_accuracy)
@@
 
+```{r, include=FALSE}
+# Threshold Function
+f0<-function(x,t=t0){
+  ifelse(x>t,1,0)
+}
+```
+
+```{r, include=FALSE}
+# Accuracy Function
+accuracy_knn <- function(df){
+  acc_overall = sum(df$act_class == df$pred_class)/nrow(df)
+  df_1 = df[df$act_class > 0,]
+  df_0 = df[df$act_class == 0,]
+  acc_1 = sum(df_1$act_class == df_1$pred_class)/nrow(df_1)
+  acc_0 = sum(df_0$act_class == df_0$pred_class)/nrow(df_0)
+  return (c(acc_overall, acc_1, acc_0))
+}
+```
+
+```{r, include=FALSE}
+# Data frame for storing values
+threshold_auc<-data.frame(k=double(0), threshold_val=double(0), auc_val=double(0), acc_overall=double(0), acc_1=double(0), acc_0=double(0))
+
+# Loop iterating over values of K for finding best value
+for(k in seq(20, 50, by=5)){
+  test_pred<-data.frame(pred_val=double(0),act_class=numeric(0))
+  for (i in 1:nrow(org_cc_data)){
+      knn_model_v1<-kknn(R1~., train = org_cc_data[-i,], test = org_cc_data[i,], k=k, kernel="rectangular", scale = TRUE)
+      knn_pred[i]<-as.double(fitted(knn_model_v1))
+      test_pred[nrow(test_pred)+1,]=c(knn_pred[i],org_cc_data$R1[i])
+  }
+
+  # Iteration over values of threshold to determine the threshold which maximizes AUC score and hence the better separation between the two classes
+  for(t0 in seq(0.2,0.8,by=0.02)){
+    test_pred$pred_class<-apply(data.frame(test_pred$pred_val),1,f0)
+    auc_val<-auc(test_pred$act_class,test_pred$pred_class)
+    print(auc_val)
+    print(c(k, t0, auc_val, accuracy_knn(test_pred)))
+    threshold_auc[nrow(threshold_auc)+1,]=c(k, t0,auc_val,accuracy_knn(test_pred)[1], accuracy_knn(test_pred)[2], accuracy_knn(test_pred)[3])
+  }
+}
+```
+
+```{r}
+threshold_auc
+```
+
+```{r}
+# Final step to determine the highest value of AUC for every K
+final_frame<-data.frame(k=double(0), threshold_val=double(0), auc_val=double(0), acc_overall=double(0), acc_1=double(0), acc_0=double(0))
+for(k_i in seq(20, 50, by=5)){
+  print(k_i)
+  data_i = threshold_auc[threshold_auc$k == k_i,]
+  print (data_i[which.max(data_i$auc_val),])
+  final_frame[nrow(final_frame)+1, ] <- data_i[which.max(data_i$auc_val),]
+}
+```
+
+```{r}
+final_frame
+```
+
 [**#Analysis:**]{style="color:red"}

### 2022-08-31 hw1-1/submission.Rmd
Final code submission

@@
 
--   ***Current Debt to Income ratio***: This is a measure of Income to Debt Capacity of the individual.
+-   ***Current Debt to Income ratio***: This is a measure of Debt to Income Capacity of the individual.
 In other words, it stacks up the overall debt obligations of the individual across mortgage, auto loan, credit cards etc., against his total income.
@@
 
-**Objective Function (**$Y$): To classify whether a readmission will happen within 30 days of the Inpatient discharge for a given patient.\
+**Objective Function (**$Y$): To classify whether a readmission will happen within 30 days of the Inpatient (Hospital) discharge for a given patient.\
 **Predictor Variables (**$X_i$): Some of the key predictor variables given below:
@@
 
-```{r}
+```{r, include=FALSE}
 #lib imports
@@
 -   C=0.001: While here the overall accuracy drops to 83% but accuracy in the negative class is significantly higher at 92%, and given this is a loan application example, it might be better to gear the model more towards the negative class.
+This is because it might be more riskier to approve the loan application for a person who ought to be denied the loan application versus misclassifying an otherwise approved applicant.
 
@@
 
-As expected (from distribution plots) variables A9, A11, A8, A10 show very significant weights in the decision boundary. Similarly variables like A1,A2 have very low weights, and can also be dropped.
+**As expected (from distribution plots) variables A9, A11, A8, A10 show very significant weights in the decision boundary.
+Similarly variables like A1,A2 have very low weights, and can also be dropped.**
 
@@
 As seen for Linear Kernels, increase in C, increases accuracy while decreasing number of support vectors.
-Because this is a very complex model, we are able to get very high accuracy, however we can't be sure if the same would generalize to unseen data.
+Because this is a very complex model, we are able to get very high accuracy on the training data, however we can't be sure if the same would generalize to unseen data.
 
@@
 
-Here again we are able to get 100 accuracy, however we can't be sure if the same would generalize to unseen data.
+Here again we are able to get 100 accuracy (it appears as if it is grossly overfitting), hence we can't be sure if the same performance would generalize to unseen data.
 
@@
 
-We implement vanilla knn model here, with rectangular kernel and majority voting (i.e probability threshold \>0.5) among neighbors to decide predicted label.
+We implement vanilla knn model here, with rectangular (standard unweighted) kernel and majority voting (i.e probability threshold \>0.5) among neighbors to decide predicted label.
 
@@
 
-The graph above here plots Accuracy on y-axis vs 'k'.
-As we increase the k, Accuracy is mostly static and after a certain threshold, it starts dropping almost linearly to 55%, where it then stabilizes.
+The plot above shows the Accuracy on y-axis vs 'k' i.e number of nearest neighbours on the x-axis.
+As we increase k, Accuracy is mostly static and after a certain threshold (~340), it starts dropping almost linearly to 55%, where it then stabilizes.
 The 55% number here is intuitive since with as many nearest neighbors as the #data points, the prediction decision basically boils down to assigning the majority label in the dataset, i.e.
 Negative.
-Therefore, for 55% the negative class, we receive a correct prediction while the rest 45% Positive class is misclassified.
+
+**At this point, we receive a correct prediction on the negative class which constitute around 55% of the overall number of data points, while the rest 45% Positive class is misclassified.**
 
 ### Optimisation over 'k' and probability threshold
+The KNN model as compared to SVM model provides us the prediction probability of each point being in a given class which can then be transformed into predicted labels based on the threshold value.
+This introduces the notion of deciding an optimal threshold value that maximizes our chosen metric. A typically used metric for classification problem is the AUC value under the ROC curve.
+
+https://en.wikipedia.org/wiki/Receiver_operating_characteristic
+
+In the code below , we have employed an approach of iterating over k and finding an optimal probability threshold corresponding to each k which maximizes the AUC score.
 
@@
 for(k_i in seq(20, 50, by=5)){
-  print(k_i)
+  #print(paste("Best Probability threshold for k=",k_i))
   data_i = threshold_auc[threshold_auc$k == k_i,]
-  print (data_i[which.max(data_i$auc_val),])
+  #print (data_i[which.max(data_i$auc_val),])
   final_frame[nrow(final_frame)+1, ] <- data_i[which.max(data_i$auc_val),]
@@
 [**#Analysis:**]{style="color:red"}
+From above table we are able to choose a combination of k and threshold values, so as to maximize accuracy.
+However, the combinations listed above exhibit very similar Accuracy ranges (overall and split by classes as well.)

### 2022-09-05 hw2/anshit/Homework_2.Rmd
Initial commit for HW2 - anshit

---
title: "Homework_2"
output: html_document
date: "2022-09-04"
---

```{r}
# imports
library(ggplot2)
library(cowplot)
library(reshape2)
# library(factoextra)
```

## Question 3.1

```{r}
cc_data = read.table('./data 3.1/credit_card_data-headers.txt', header = TRUE, sep='', dec='.')
head(cc_data)
dim(cc_data)
```

## Question 4.2

```{r}
iris_data = read.table('./iris.txt', header = TRUE, sep='', dec='.')
head(iris_data)
dim(iris_data)
```

### Exploratory Data Analysis

```{r}
summary(iris_data)
```

```{r}
table(iris_data$Species)
```

```{r}
# Density Plots
my_plots = lapply(colnames(iris_data), function(x){
  p = ggplot(iris_data) + aes_string(x)
  if(x=='Species') {
    p = p + geom_bar()
  }
  else {
    p = p + geom_density()
  }
})

plot_grid(plotlist = my_plots)
```

```{r}
# Correlation Matrix
cor_mat = round(cor(iris_data[,1:4]), 2)
melted_mat = melt(cor_mat)
ggplot(melted_mat, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile() + geom_text(aes(Var1, Var2, label=value), color='white', size=3)
```

### Modelling

```{r}
# Calculation of best 'K'
elbow_graph_data <- data.frame(K=integer(), sum_of_square=double())
for(k in seq(1, 10, by=1)){
  km.res <- kmeans(as.matrix(iris_data[,1:4]), k, nstart = 25)
  elbow_graph_data[nrow(elbow_graph_data) + 1, ] = c(k, km.res$tot.withinss)
}
```

```{r}
elbow_graph_data
```

```{r}
plot(elbow_graph_data, type='b', ylab='Total Sum of Square Within', main='Elbow Graph', )
```

Best K is 3 because we are getting marginal benefit after that

```{r}
best_km <- kmeans(as.matrix(scale(iris_data[,1:4])), 3, nstart = 25)
best_km
```

```{r}
# renaming predicted clusters according to categorical response variable
renamed_clusters <- c("versicolor", "setosa", "virginica")[best_km$cluster]
```

```{r}
kmeans_basic_table <- data.frame(best_km$size, best_km$centers)
kmeans_basic_df <- data.frame(actual_value=iris_data[,5], prediction=renamed_clusters)
```

```{r}
kmeans_basic_table
```

```{r}
plot(scale(iris_data[,1:4]), col=best_km$cluster)
```

```{r}
prediction_accuracy <- round(sum(kmeans_basic_df$actual_value == kmeans_basic_df$prediction) / nrow(kmeans_basic_df) * 100, 2)
print(paste('Accuracy Acheived ->', prediction_accuracy, '%' ))
```

### 2022-09-05 hw2/anshit/Homework_2.Rmd
added submisison.Rmd

@@
 library(reshape2)
-# library(factoextra)
+library(factoextra)
 ```
@@
 for(k in seq(1, 10, by=1)){
-  km.res <- kmeans(as.matrix(iris_data[,1:4]), k, nstart = 25)
+  km.res <- kmeans(as.matrix(scale(iris_data[,1:4])), k, nstart = 25)
   elbow_graph_data[nrow(elbow_graph_data) + 1, ] = c(k, km.res$tot.withinss)
@@
 ```{r}
+cm <- table(iris_data$Species, best_km$cluster)
+cm
+```
+
+```{r}
+best_km$centers
+```
+
+```{r}
 # renaming predicted clusters according to categorical response variable
-renamed_clusters <- c("versicolor", "setosa", "virginica")[best_km$cluster]
+# renamed_clusters <- c("versicolor", "setosa", "virginica")[best_km$cluster]
 ```
@@
 kmeans_basic_table <- data.frame(best_km$size, best_km$centers)
-kmeans_basic_df <- data.frame(actual_value=iris_data[,5], prediction=renamed_clusters)
+kmeans_basic_table
 ```
@@
 ```{r}
-kmeans_basic_table
+kmeans_basic_df <- data.frame(actual_value=iris_data[,5], prediction=best_km$cluster)
+kmeans_basic_df
 ```
@@
 ```{r}
-prediction_accuracy <- round(sum(kmeans_basic_df$actual_value == kmeans_basic_df$prediction) / nrow(kmeans_basic_df) * 100, 2)
-print(paste('Accuracy Acheived ->', prediction_accuracy, '%' ))
+# prediction_accuracy <- round(sum(kmeans_basic_df$actual_value == kmeans_basic_df$prediction) / nrow(kmeans_basic_df) * 100, 2)
+# print(paste('Accuracy Acheived ->', prediction_accuracy, '%' ))
 ```
+
+### Plotting Clusters
+
+```{r}
+fviz_cluster(best_km, data=scale(iris_data[,1:2]), geom=c("point"), ellipse.type = "euclid")
+```
+***Analysing For K=4***
+
+```{r}
+alt_kmeans <- kmeans(as.matrix(scale(iris_data[,1:4])), 4, nstart = 25)
+alt_kmeans
+```
+
+```{r}
+cm <- table(iris_data$Species, alt_kmeans$cluster)
+cm
+```
+
+```{r}
+alt_kmeans$centers
+```
+
+```{r}
+alt_kmeans_table_size_v_centers <- data.frame(alt_kmeans$size, alt_kmeans$centers)
+alt_kmeans_table_size_v_centers
+```
+
+```{r}
+alt_kmeans_row_pred <- data.frame(actual_value=iris_data[,5], prediction=alt_kmeans$cluster)
+alt_kmeans_row_pred
+```
+
+```{r}
+# clusters = data.frame(alt_kmeans$cluster)
+# clusters[clusters$alt_kmeans.cluster==4,]
+mapped_data <- data.frame(iris_data[,1:4], prediction=alt_kmeans$cluster)
+```
+
+```{r}
+# Summary for Cluster 1
+summary(mapped_data[mapped_data$prediction==1,1:4])
+```
+
+```{r}
+# Summary for Cluster 2
+summary(mapped_data[mapped_data$prediction==2,1:4])
+```
+
+```{r}
+# Summary for Cluster 3
+summary(mapped_data[mapped_data$prediction==3,1:4])
+```
+
+```{r}
+# Summary for Cluster 4
+summary(mapped_data[mapped_data$prediction==4,1:4])
+```
+
+```{r}
+plot(scale(iris_data[,1:4]), col=best_km$cluster)
+```
+
+```{r}
+fviz_cluster(best_km, data=scale(iris_data[,1:2]), geom=c("point"), ellipse.type = "euclid")
+```

### 2022-09-05 hw2/ashish/a.0_kfold_knnn.Rmd
added ques3.1 part 1

---
title: "ques 3.1"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r, include=FALSE}
setwd("~/data_projects/fall22_hw/isye6501_aaa/hw2/ashish")

#lib imports
library(cowplot)
library(ggplot2)
library(reshape2)
library(kknn)
library(pROC)
```

### Read Data and repeat vanilla knnn model from last time

```{r}
org_cc_data <- read.table(file = '../data 3.1/credit_card_data-headers.txt', sep = "\t", header = TRUE)
dim(org_cc_data)
```

```{r}
#predict ith row of dataframe
predcit_i = function(i =1, k0 = 5){
  model = kknn(R1~.,train = org_cc_data[-i,],test = org_cc_data[i,],k=k0,kernel="rectangular",scale = TRUE)
  return (fitted(model))
  #return (as.integer(fitted(model) > threshold))
}

#predictions for a given k
predcit_k = function(k0 = 5){
  predictions = lapply(seq(1,nrow(org_cc_data)),predcit_i, k0)
  return (predictions)
}

#precit probabilities for specific k
pred_prob <- predcit_k(k0=5)

# decide best threshold
best_threshold = function(pred_prob, df_test=org_cc_data){
  myroc <- roc(df_test$R1,as.numeric(unlist(pred_prob)), smooth=FALSE)
  print(paste("AUC is",auc(myroc)))
  #plot(myroc, auc.polygon=TRUE)
  return (coords(myroc,x="best")["threshold"][[1]])
}

ti = best_threshold(pred_prob)
print (ti)
#Accuracy for best threshold
acc_func = function(y_act, pred_probs, thresh = 0.5){
  y_pred = lapply(pred_probs, function(x) if(x>=thresh) 1 else 0)
  y_pred = as.numeric(unlist(y_pred))
  acc_overall = sum(y_act == y_pred)/length(y_act)
  acc_overall = round(acc_overall,4)* 100
  return (acc_overall)
}

acc_func(y_act = org_cc_data$R1, pred_probs = pred_prob, thresh = ti)
acc_func(y_act = org_cc_data$R1, pred_probs = pred_prob, thresh = 0.9)
```

```{r}
master_kknn_func = function(y_act = org_cc_data$R1, ki = 10){
  pred_prob <- predcit_k(k0=ki)
  ti = best_threshold(pred_prob)
  print (paste("Best threshold",ti))
  return (acc_func(y_act = y_act, pred_probs = pred_prob, thresh = ti))
}

master_kknn_func(ki = 5)
master_kknn_func(ki = 7)
master_kknn_func(ki = 10)
```

## Cross validation for kknn

-   Split data into 5 parts

    -   Keep first 4 for train, last 1 for validation

    -   On the train, obtain k values of nearest neighbors for max auc

    -   get acc on train and test

```{r}
#Split org_data into 5 parts
df_copy = org_cc_data
df_copy$cross_validation_i <- sample(5, size = nrow(df_copy), replace = TRUE)

aggregate(df_copy$R1,list(df_copy$cross_validation_i), FUN = mean)
```

```{r}
#Function to train model and return prediction on test
train_test_i = function(df_train,df_test, k0 = 5){
  model = kknn(R1~.,train = df_train,test = df_test,k=k0,kernel="rectangular",scale = TRUE)
  return (fitted(model))
}

#Train test for one group
cv_function_i = function(df= df_copy, test_index_choice,k_choice){
  print(test_index_choice)
  #split data into train and test
  mask_test_i = df$cross_validation_i == test_index_choice
  df_train = df[!mask_test_i,]
  df_test = df[mask_test_i,]
  print (paste("Train,Test Size =",nrow(df_train),nrow(df_test)))
  #train and get probaility predictions on test model
  test_probabilities = train_test_i(df_train,df_test, k0 = k_choice)

  #get best threshold for test probabilities
  ti = best_threshold(test_probabilities, df_test = df_test)
  print (paste("Best threshold is",ti))
  #get test accuracy
  y_act = df_test$R1
  #print (paste("Length of Test Probabilities",nrow(df_test),length(test_probabilities)))
  return (acc_func(y_act = y_act, pred_probs = test_probabilities, thresh = ti))
}

#CV function: pass k, return test accuracy for all folds
cv_function = function(df = df_copy,k_choice = 5){
  cv_accuracy = lapply(seq(1,5), function(x) cv_function_i(df=df_copy,x,k_choice))
}
```

```{r}
cv_accuracy = cv_function(df = df_copy,k_choice = 5)
mean(unlist(cv_accuracy))
```

## Iterate on different models and use CV for best model

```{r, include = FALSE}
#Iterate over k and get mean test
acc_df <- data.frame(matrix(ncol = 0, nrow = length(seq(3,25,2))))
acc_df$k_kknn = seq(3,25,2)
mean_cv_acc_k = lapply(seq(3,25,2), function(x) mean(unlist(cv_function(df = df_copy,x))))
min_cv_acc_k = lapply(seq(3,25,2), function(x) min(unlist(cv_function(df = df_copy,x))))
acc_df$mean_cv_accuracy = unlist(mean_cv_acc_k)
acc_df$min_cv_accuracy = unlist(min_cv_acc_k)
#mean(unlist(cv_accuracy))
```

```{r}
acc_df
```

### 2022-09-05 hw2/ashish/a.0_kfold_knnn.Rmd

@@
 #Iterate over k and get mean test
-acc_df <- data.frame(matrix(ncol = 0, nrow = length(seq(3,25,2))))
-acc_df$k_kknn = seq(3,25,2)
-mean_cv_acc_k = lapply(seq(3,25,2), function(x) mean(unlist(cv_function(df = df_copy,x))))
-min_cv_acc_k = lapply(seq(3,25,2), function(x) min(unlist(cv_function(df = df_copy,x))))
+acc_df <- data.frame(matrix(ncol = 0, nrow = length(seq(3,101,2))))
+acc_df$k_kknn = seq(3,101,2)
+mean_cv_acc_k = lapply(seq(3,101,2), function(x) mean(unlist(cv_function(df = df_copy,x))))
+min_cv_acc_k = lapply(seq(3,101,2), function(x) min(unlist(cv_function(df = df_copy,x))))
 acc_df$mean_cv_accuracy = unlist(mean_cv_acc_k)
@@
 ```
+
+```{r}
+ggplot(data = acc_df, aes(x = k_kknn, y = mean_cv_accuracy)) +
+  geom_line() +geom_point() +
+  geom_vline(xintercept = acc_df[which.max(acc_df$mean_cv_accuracy),1],
+             color ="magenta", size = 1.5, alpha = 0.5)
+```

### 2022-09-05 hw2/submission.Rmd

--- 2022-09-05 hw2/anshit/Homework_2.Rmd
@@
 ---
-title: "Homework_2"
+title: "Homework_2_Submission"
 output: html_document
@@
 library(reshape2)
-# library(factoextra)
+library(factoextra)
 ```
-
-## Question 3.1
-
-```{r}
-cc_data = read.table('./data 3.1/credit_card_data-headers.txt', header = TRUE, sep='', dec='.')
-head(cc_data)
-dim(cc_data)
-```
-
-## Question 4.2
-
-```{r}
-iris_data = read.table('./iris.txt', header = TRUE, sep='', dec='.')
-head(iris_data)
-dim(iris_data)
-```
-
-### Exploratory Data Analysis
-
-```{r}
-summary(iris_data)
-```
-
-```{r}
-table(iris_data$Species)
-```
-
-```{r}
-# Density Plots
-my_plots = lapply(colnames(iris_data), function(x){
-  p = ggplot(iris_data) + aes_string(x)
-  if(x=='Species') {
-    p = p + geom_bar()
-  }
-  else {
-    p = p + geom_density()
-  }
-})
-
-plot_grid(plotlist = my_plots)
-```
-
-```{r}
-# Correlation Matrix
-cor_mat = round(cor(iris_data[,1:4]), 2)
-melted_mat = melt(cor_mat)
-ggplot(melted_mat, aes(x=Var1, y=Var2, fill=value)) +
-  geom_tile() + geom_text(aes(Var1, Var2, label=value), color='white', size=3)
-```
-
-### Modelling
-
-```{r}
-# Calculation of best 'K'
-elbow_graph_data <- data.frame(K=integer(), sum_of_square=double())
-for(k in seq(1, 10, by=1)){
-  km.res <- kmeans(as.matrix(iris_data[,1:4]), k, nstart = 25)
-  elbow_graph_data[nrow(elbow_graph_data) + 1, ] = c(k, km.res$tot.withinss)
-}
-```
-
-```{r}
-elbow_graph_data
-```
-
-```{r}
-plot(elbow_graph_data, type='b', ylab='Total Sum of Square Within', main='Elbow Graph', )
-```
-
-Best K is 3 because we are getting marginal benefit after that
-
-```{r}
-best_km <- kmeans(as.matrix(scale(iris_data[,1:4])), 3, nstart = 25)
-best_km
-```
-
-```{r}
-# renaming predicted clusters according to categorical response variable
-renamed_clusters <- c("versicolor", "setosa", "virginica")[best_km$cluster]
-```
-
-```{r}
-kmeans_basic_table <- data.frame(best_km$size, best_km$centers)
-kmeans_basic_df <- data.frame(actual_value=iris_data[,5], prediction=renamed_clusters)
-```
-
-```{r}
-kmeans_basic_table
-```
-
-```{r}
-plot(scale(iris_data[,1:4]), col=best_km$cluster)
-```
-
-```{r}
-prediction_accuracy <- round(sum(kmeans_basic_df$actual_value == kmeans_basic_df$prediction) / nrow(kmeans_basic_df) * 100, 2)
-print(paste('Accuracy Acheived ->', prediction_accuracy, '%' ))
-```

### 2022-09-06 hw2/ashish/a.0_kfold_knnn.Rmd
ques 3.1 complete

--- 2022-09-05 hw2/ashish/a.0_kfold_knnn.Rmd
@@
 setwd("~/data_projects/fall22_hw/isye6501_aaa/hw2/ashish")
+
+set.seed(520)
 
@@
 train_test_i = function(df_train,df_test, k0 = 5){
-  model = kknn(R1~.,train = df_train,test = df_test,k=k0,kernel="rectangular",scale = TRUE)
+  f = R1~A1+A2+A3+A8+A9+A10+A11+A12+A14+A15
+  model = kknn(f,train = df_train,test = df_test,k=k0,kernel="rectangular",scale = TRUE)
   return (fitted(model))
@@
 cv_function_i = function(df= df_copy, test_index_choice,k_choice){
-  print(test_index_choice)
+  print (paste("test index for CV Fold",test_index_choice))
   #split data into train and test
@@
   ti = best_threshold(test_probabilities, df_test = df_test)
+  if (length(ti) >1){
+    ti = ti[[1]] #select 1st when there are multiple optimal thresholds
+  }
   print (paste("Best threshold is",ti))
@@
 cv_function = function(df = df_copy,k_choice = 5){
+  cat(paste("\nFor k value",k_choice))
   cv_accuracy = lapply(seq(1,5), function(x) cv_function_i(df=df_copy,x,k_choice))
@@
 
-```{r, include = FALSE}
+```{r, include=FALSE}
 #Iterate over k and get mean test
-acc_df <- data.frame(matrix(ncol = 0, nrow = length(seq(3,25,2))))
-acc_df$k_kknn = seq(3,25,2)
-mean_cv_acc_k = lapply(seq(3,25,2), function(x) mean(unlist(cv_function(df = df_copy,x))))
-min_cv_acc_k = lapply(seq(3,25,2), function(x) min(unlist(cv_function(df = df_copy,x))))
+acc_df <- data.frame(matrix(ncol = 0, nrow = length(seq(3,71,2))))
+acc_df$k_kknn = seq(3,71,2)
+mean_cv_acc_k = lapply(seq(3,71,2), function(x) mean(unlist(cv_function(df = df_copy,x))))
+min_cv_acc_k = lapply(seq(3,71,2), function(x) min(unlist(cv_function(df = df_copy,x))))
 acc_df$mean_cv_accuracy = unlist(mean_cv_acc_k)
@@
 ```
+
+```{r}
+ggplot(data = acc_df, aes(x = k_kknn, y = mean_cv_accuracy)) +
+  geom_line() +geom_point() +
+  geom_vline(xintercept = acc_df[which.max(acc_df$mean_cv_accuracy),1],
+             color ="magenta", size = 1.5, alpha = 0.5) +
+  ggtitle(paste("max accuracy for k =",acc_df[which.max(acc_df$mean_cv_accuracy),1]))
+
+print (paste("max accuracy for k =",acc_df[which.max(acc_df$mean_cv_accuracy),1]))
+```
+
+## Ques 3.1: part 2
+
+#split data into 70% train, 15% validation and 15% test
+
+```{r}
+#Split org_data into 5 parts
+df_copy2 = org_cc_data
+df_copy2$split_i <- sample(100, size = nrow(df_copy), replace = TRUE)
+
+df_copy2$split = "train"
+df_copy2$split[70 < df_copy2$split_i] = "valid"
+df_copy2$split[85 < df_copy2$split_i] = "test"
+
+summary_df = aggregate(df_copy2$R1,list(df_copy2$split), 
+          FUN = function(x) c(mean = mean(x), proportion = length(x)/nrow(df_copy2)))
+print.data.frame(summary_df)
+```
+
+Find optimal k using training on train and validation
+
+```{r}
+df_train = df_copy2[df_copy2$split=="train",]
+dim(df_train)
+df_valid = df_copy2[df_copy2$split=="valid",]
+dim(df_valid)
+df_test= df_copy2[df_copy2$split=="test",]
+dim(df_test)
+```
+
+```{r}
+train_valid_func = function(df_train, df_test, k_choice){
+  print (paste("For k=",k_choice))
+  #split data into train and test
+  print (paste("Train,Test Size =",nrow(df_train),nrow(df_test)))
+  #train and get probaility predictions on test model
+  test_probabilities = train_test_i(df_train,df_test, k0 = k_choice)
+  #get best threshold for test probabilities
+  ti = best_threshold(test_probabilities, df_test = df_test)
+  print (paste("Best threshold is",ti))
+  #get test accuracy
+  y_act = df_test$R1
+  #print (paste("Length of Test Probabilities",nrow(df_test),length(test_probabilities)))
+  return (acc_func(y_act = y_act, pred_probs = test_probabilities, thresh = ti))
+}
+```
+
+```{r}
+train_valid_func(df_train = df_train, df_test = df_valid, k_choice = 20)
+```
+
+```{r, include=FALSE}
+acc_df2 <- data.frame(matrix(ncol = 0, nrow = length(seq(3,71,2))))
+acc_df2$k_kknn = seq(3,71,2)
+acc = lapply(seq(3,71,2), function(x) 
+  train_valid_func(df_train = df_train, df_test = df_valid, k_choice = x))
+acc_df2$acc_validation_set = unlist(acc)
+```
+
+```{r}
+ggplot(data = acc_df2, aes(x = k_kknn, y = acc_validation_set)) +
+  geom_line() +geom_point() +
+  geom_vline(xintercept = acc_df2[which.max(acc_df2$acc_validation_set),1],
+             color ="magenta", size = 1.5, alpha = 0.5) +
+  ggtitle(paste("max validation accuracy for k =",acc_df2[which.max(acc_df2$acc_validation_set),1]))
+
+print (paste("max validation accuracy for k =",acc_df2[which.max(acc_df2$acc_validation_set),1]))
+```
+
+rebuild model with optimal k, with
+
+-   Train = Train + Validation
+
+-   Test = Test
+
+```{r}
+k_optimal = acc_df2[which.max(acc_df2$acc_validation_set),1]
+acc_optimal_k = train_valid_func(df_train = rbind(df_train,df_valid), df_test = df_test, k_choice = k_optimal)
+print (paste("With optimal k value, Accuracy on test =",acc_optimal_k))
+```
+
+```{r}
+
+```
+
+```{r}
+
+```
+
+```{r}
+
+```
+
+```{r}
+
+```

### 2022-09-06 hw2/abhinav/clustering.Rmd

```{r}
library(cowplot)
library(ggplot2)
library(reshape2)
library(kknn)
library(pROC)
library(datasets)
```

```{r}
iris_data<-read.table(file = './iris.txt', sep = "", header = TRUE)
clus_data<-iris_data[c(1,2,3,4)]
clus_data<-scale(clus_data)
head(clus_data)
```

```{r,include=FALSE}
# pkgs <- c("factoextra",  "NbClust")
# install.packages(pkgs)
library(ggplot2)
library(factoextra)
library(NbClust)
library(parameters)
```

```{r}
# Elbow method
fviz_nbclust(clus_data, kmeans, method = "wss") +
    geom_vline(xintercept = 4, linetype = 2)+
  labs(subtitle = "Elbow method")

# Silhouette method
fviz_nbclust(clus_data, kmeans, method = "silhouette")+
  labs(subtitle = "Silhouette method")

# Gap statistic
# nboot = 50 to keep the function speedy. 
# recommended value: nboot= 500 for your analysis.
# Use verbose = FALSE to hide computing progression.
set.seed(123)
fviz_nbclust(clus_data, kmeans, nstart = 25,  method = "gap_stat", nboot = 50)+
  labs(subtitle = "Gap statistic method")
```

```{r}
n_clust<-n_clusters(data.frame(clus_data),package = c("easystats", "NbClust", "mclust"),standardize=FALSE)
data.frame(clus_data)
n_clust
#n_clust
```

### 2022-09-06 hw2/submission.Rmd
submission added ques 3.1

@@
 ---
-title: "Homework_2_Submission"
-output: html_document
+title: "ISYE 6501 | Submission HW2"
+output:
+  html_document:
+    df_print: paged
+  pdf_document: default
 date: "2022-09-04"
@@
 
-```{r}
+# ISYE 6501 \| Fall 22
+
+-   Ashish Dhiman \| [ashish.dhiman\@gatech.edu](mailto:ashish.dhiman@gatech.edu){.email}
+
+-   Abhinav Arun \| [aarun60\@gatech.edu](mailto:aarun60gatech.edu){.email}
+
+-   Anshit Verma \| [averma373\@gatech.edu](mailto:averma373@gatech.edu){.email}
+
+[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
+
+```{r, include=FALSE}
+
+set.seed(61)
+
 # imports
@@
 library(reshape2)
-library(factoextra)
+#library(factoextra)
+library(kknn)
+library(pROC)
 ```
+
+## Read Data
+
+```{r}
+org_cc_data <- read.table(file = './data 3.1/credit_card_data-headers.txt', sep = "\t", header = TRUE)
+dim(org_cc_data)
+```
+
+## Question 3.1
+
+### Part 1: Cross Validation for finding best knnn model
+
+Here we perform 5 fold cross validation, and follow the following pseudo code:
+
+1.  Divide the data into 5 equal parts
+
+2.  Choose i'th part as test data and train knn on the rest 4
+
+    1.  Find the threshold which maximises AUC using ROC curve
+
+    2.  With this optimal threshold find accuracy on test set
+
+3.  Repeat step 2 but slide i to the next part
+
+4.  When all parts are rotated into test set, calculate the measure for accuracy of model:
+
+    1.  Mean test accuracy across 5 folds
+
+    2.  Min test accuracy across the 5 folds
+
+**We repeat the above 4 steps for different k values, to find the best model(or choice of k)**
+
+### Step 1:
+
+```{r}
+#Split org_data into 5 parts
+df_copy = org_cc_data
+df_copy$cross_validation_i <- sample(5, size = nrow(df_copy), replace = TRUE)
+
+summary_df = aggregate(df_copy$R1,list(df_copy$cross_validation_i), 
+          FUN = function(x) c(mean = mean(x), proportion = length(x)/nrow(df_copy)))
+print.data.frame(summary_df)
+```
+
+[**#Analysis**]{style="color:red"}
+
+The above summary shows the mean of R1 and relative size of each fold. Ideally for all the folds this should be 45% mean and 20%, but we see minor deviations due to random sampling.
+
+```{r}
+#Function to train model and return prediction on test
+train_test_i = function(df_train,df_test, k0 = 5){
+  f = R1~A1+A2+A3+A8+A9+A10+A11+A12+A14+A15
+  model = kknn(f,train = df_train,test = df_test,k=k0,kernel="rectangular",scale = TRUE)
+  return (fitted(model))
+}
+```
+
+### Step 2
+
+```{r}
+#Train model and test on ith group
+cv_function_i = function(df, test_index_choice,k_choice){
+  print("")
+  print (paste("test index for CV Fold",test_index_choice))
+  #split data into train and test
+  mask_test_i = df$cross_validation_i == test_index_choice
+  df_train = df[!mask_test_i,]
+  df_test = df[mask_test_i,]
+  print (paste("Train,Test Size =",nrow(df_train),nrow(df_test)))
+  #train and get probaility predictions on test model
+  test_probabilities = train_test_i(df_train,df_test, k0 = k_choice)
+
+  #get best threshold for test probabilities
+  ti = best_threshold(test_probabilities, df_test = df_test)
+  if (length(ti) >1){
+    ti = ti[[1]] #select 1st when there are multiple optimal thresholds
+  }
+  print (paste("Best threshold is",ti))
+  #get test accuracy
+  y_act = df_test$R1
+  #print (paste("Length of Test Probabilities",nrow(df_test),length(test_probabilities)))
+  return (acc_func(y_act = y_act, pred_probs = test_probabilities, thresh = ti))
+}
+```
+
+```{r}
+#Test fucntion with a sample i
+cv_function_i(df = df_copy, test_index_choice = 1, k_choice = 16)
+```
+
+### Step 3
+
+```{r}
+#CV function: pass k, return test accuracy for all folds
+cv_function = function(df = df_copy,k_choice = 5){
+  print("----------------------------------------------")
+  print(paste("For k value",k_choice))
+  cv_accuracy = lapply(seq(1,5), function(x) cv_function_i(df=df_copy,x,k_choice))
+  print("----------------------------------------------")
+  return (cv_accuracy)
+}
+```
+
+```{r}
+#Test function with a sample k
+cv_accuracy = cv_function(df = df_copy,k_choice = 16)
+```
+
+```{r}
+unlist(cv_accuracy)
+mean(unlist(cv_accuracy))
+```
+
+The values match the above test run values
+
+### Step 4: Iterate on different models and use CV for best model
+
+```{r, include=FALSE}
+#Iterate over k and get mean test accuracy
+
+end_k = 77 #half of test set size
+
+acc_df <- data.frame(matrix(ncol = 0, nrow = length(seq(3,end_k,2))))
+acc_df$k_kknn = seq(3,end_k,2)
+mean_cv_acc_k = lapply(seq(3,end_k,2), function(x) mean(unlist(cv_function(df = df_copy,x))))
+min_cv_acc_k = lapply(seq(3,end_k,2), function(x) min(unlist(cv_function(df = df_copy,x))))
+sd_cv_acc_k = lapply(seq(3,end_k,2), function(x) sd(unlist(cv_function(df = df_copy,x))))
+acc_df$mean_cv_accuracy = unlist(mean_cv_acc_k)
+acc_df$min_cv_accuracy = unlist(min_cv_acc_k)
+acc_df$sd_cv_accuracy = unlist(sd_cv_acc_k)
+```
+
+```{r}
+acc_df
+```
+
+[**#Analysis**]{style="color:red"}
+
+The above dataframe has average CV accuracy, min CV accuracy and Std.
+Deviation of Accuracy for each k value tried.
+While average CV Accuracy is a typical performance measure, we have also retained min CV accuracy which is like a lower bound on the performance of models across various validation sets, and we can also use max of this min bound to find optimal k.
+
+We have also kept standard deviation, such that we can able to infer the variance of accuracy across cross folds.
+This is helpful because the maximum CV accuracy for different values across k, varies within a percent (See graph below).
+
+#### Plot the accuracy with k, and find optimal k which maximizes accuracy
+
+```{r}
+ggplot(data = acc_df, aes(x = k_kknn, y = mean_cv_accuracy)) +
+  geom_line() +geom_point() +
+  geom_vline(xintercept = acc_df[which.max(acc_df$mean_cv_accuracy),1],
+             color ="magenta", size = 1.5, alpha = 0.5) +
+  ggtitle(paste("max accuracy for k =",acc_df[which.max(acc_df$mean_cv_accuracy),1]))
+
+print (paste("max accuracy for k =",acc_df[which.max(acc_df$mean_cv_accuracy),1]))
+```
+
+[**#Analysis**]{style="color:red"}
+
+Here we see that for k = 47, we get maximum average CV accuracy of 87.74%.
+
+However, as noted above the nuance to note her is that, for k \> 30, that accuracy mostly oscillates at almost the same range with delta of 0.5.
+But since the standard deviation of max accuracy among cross folds is 2.45 itself, this choice of k becomes a little tricky.
+In other words, we
+
+Another effect to note here is that, this optimal value of k, is also subject to the random sampling done to create the folds, and will change with a different random seed setting.
+
+```{r}
+acc_df[which.max(acc_df$mean_cv_accuracy),]
+```
+
+## Question 3.1
+
+### Part 2: Train, Validation and Test on knnn
+
+```{r}
+#split data into 70% train, 15% validation and 15% test
+df_copy2 = org_cc_data
+df_copy2$split_i <- sample(100, size = nrow(df_copy), replace = TRUE)
+
+df_copy2$split = "train"
+df_copy2$split[70 < df_copy2$split_i] = "valid"
+df_copy2$split[85 < df_copy2$split_i] = "test"
+
+summary_df = aggregate(df_copy2$R1,list(df_copy2$split), 
+          FUN = function(x) c(mean = mean(x), proportion = length(x)/nrow(df_copy2)))
+print.data.frame(summary_df)
+```
+
+Again slight deviation from expected values is seen due to random effects
+
+```{r}
+df_train = df_copy2[df_copy2$split=="train",]
+dim(df_train)
+df_valid = df_copy2[df_copy2$split=="valid",]
+dim(df_valid)
+df_test= df_copy2[df_copy2$split=="test",]
+dim(df_test)
+```
+
+### Find optimal k using training on train and validation
+
+```{r}
+train_valid_func = function(df_train, df_test, k_choice){
+  print (paste("For k=",k_choice))
+  #split data into train and test
+  print (paste("Train,Test Size =",nrow(df_train),nrow(df_test)))
+  #train and get probaility predictions on test model
+  test_probabilities = train_test_i(df_train,df_test, k0 = k_choice)
+  #get best threshold for test probabilities
+  ti = best_threshold(test_probabilities, df_test = df_test)
+  print (paste("Best threshold is",ti))
+  #get test accuracy
+  y_act = df_test$R1
+  #print (paste("Length of Test Probabilities",nrow(df_test),length(test_probabilities)))
+  return (acc_func(y_act = y_act, pred_probs = test_probabilities, thresh = ti))
+}
+```
+
+Test the function for a sample value
+
+```{r}
+train_valid_func(df_train = df_train, df_test = df_valid, k_choice = 20)
+```
+
+#### Run the above function for variety of k values
+
+```{r, include=FALSE}
+acc_df2 <- data.frame(matrix(ncol = 0, nrow = length(seq(3,end_k,2))))
+acc_df2$k_kknn = seq(3,end_k,2)
+acc = lapply(seq(3,end_k,2), function(x) 
+  train_valid_func(df_train = df_train, df_test = df_valid, k_choice = x))
+acc_df2$acc_validation_set = unlist(acc)
+```
+
+#### Plot the accuracy values
+
+```{r}
+ggplot(data = acc_df2, aes(x = k_kknn, y = acc_validation_set)) +
+  geom_line() +geom_point() +
+  geom_vline(xintercept = acc_df2[which.max(acc_df2$acc_validation_set),1],
+             color ="magenta", size = 1.5, alpha = 0.5) +
+  ggtitle(paste("max validation accuracy for k =",acc_df2[which.max(acc_df2$acc_validation_set),1]))
+
+print (paste("max validation accuracy for k =",acc_df2[which.max(acc_df2$acc_validation_set),1]))
+```
+
+#### Rebuild model with optimal k, with
+
+-   Train = Train + Validation
+
+-   Test = Test
+
+```{r}
+k_optimal = acc_df2[which.max(acc_df2$acc_validation_set),1]
+acc_optimal_k = train_valid_func(df_train = rbind(df_train,df_valid), df_test = df_test, k_choice = k_optimal)
+print (paste("With optimal k value, Accuracy on test =",acc_optimal_k))
+```
+
+[**#Analysis**]{style="color:red"}
+
+Here we see that for k = 23, we get maximum average test accuracy of 84%.
+
+Note, that both the optimal k value here and the optimal accuracy value are different than part 1 above.
+The test accuracy here is lower at 84% compared to 87% of accuracy in CV case, which is expected, since 84% is accuracy at true test, while cv accuracy is averaged across all the folds.

### 2022-09-06 hw2/anshit/Homework_2.Rmd
added box plots for analysis

@@
 iris_data = read.table('./iris.txt', header = TRUE, sep='', dec='.')
-head(iris_data)
+head(iris_data,50)
 dim(iris_data)
@@
 # Calculation of best 'K'
-elbow_graph_data <- data.frame(K=integer(), sum_of_square=double())
+elbow_graph_data <- data.frame(K=integer(), within_sum_of_square=double())
 for(k in seq(1, 10, by=1)){
@@
 ```{r}
-plot(elbow_graph_data, type='b', ylab='Total Sum of Square Within', main='Elbow Graph', )
+plot(elbow_graph_data, type='b', ylab='Total Sum of Square Within', main='Elbow Graph', col='blue')
+```
+```{r}
+# Calculation of best 'K'
+bw_graph_data <- data.frame(K=integer(), between_sum_of_square=double())
+for(k in seq(1, 10, by=1)){
+  km.res <- kmeans(as.matrix(scale(iris_data[,1:4])), k, nstart = 25)
+  bw_graph_data[nrow(bw_graph_data) + 1, ] = c(k, round(km.res$betweenss,2))
+}
+#options(scipen=999)
+bw_graph_data
+```
+
+```{r}
+plot(bw_graph_data, type='b', ylab='Betweeen Sum of Square (Intra Cluster)', main='BW SS Graph', col='green')
+```
+
+```{r}
+# Calculation of best 'K'
+ratio_graph_data <- data.frame(K=integer(), ratio=double())
+for(k in seq(2, 10, by=1)){
+  km.res <- kmeans(as.matrix(scale(iris_data[,1:4])), k, nstart = 25)
+  ratio_graph_data[nrow(ratio_graph_data) + 1, ] = c(k, round(km.res$tot.withinss/km.res$betweenss,2))
+}
+#options(scipen=999)
+ratio_graph_data
+```
+```{r}
+plot(ratio_graph_data, type='b', ylab='Ratio Within/Between SS', main='Quality', col='red')
 ```
@@
 best_km <- kmeans(as.matrix(scale(iris_data[,1:4])), 3, nstart = 25)
-best_km
+out<-cbind(iris_data,tagged_cluster=best_km$cluster)
+out
 ```
+```{r}
+boxplot(Sepal.Length~tagged_cluster,data=out)
+```
+
+```{r}
+out$tagged_cluster<-as.factor(out$tagged_cluster)
+ggplot(out, aes(x=tagged_cluster, y=Sepal.Length,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
+
+out$tagged_cluster<-as.factor(out$tagged_cluster)
+ggplot(out, aes(x=tagged_cluster, y=Sepal.Width,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
+
+out$tagged_cluster<-as.factor(out$tagged_cluster)
+ggplot(out, aes(x=tagged_cluster, y=Petal.Length,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
+
+out$tagged_cluster<-as.factor(out$tagged_cluster)
+ggplot(out, aes(x=tagged_cluster, y=Petal.Width,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
+```
+Petal width and Petal Length are strong indicators for one class (setosa)
 
@@
 ```{r}
-best_km$centers
+# function getStatusForClusterAndSpecie(row){
+#   if(row$tagged_cluster=="1" & row$Species=="versicolor"){'improper_versicolor'}
+#   else if(row$tagged_cluster=="2" & row$Species=="virginica"){'improper_virginica'}
+#   else if()
+#   
+# }
+```
+
+```{r}
+classifier_df = out[!out$Species=='setosa',]
+classifier_df$status = ''
+classifier_df$status[(classifier_df$tagged_cluster=="1" & classifier_df$Species=="versicolor")] = 'improper_versicolor'
+classifier_df$status[(classifier_df$tagged_cluster=="2" & classifier_df$Species=="virginica")] = 'improper_virginica'
+classifier_df$status[(classifier_df$tagged_cluster=="2" & classifier_df$Species=="versicolor")] = 'proper_versicolor'
+classifier_df$status[(classifier_df$tagged_cluster=="1" & classifier_df$Species=="virginica")] = 'proper_virginica'
+
+classifier_df
+
+```
+```{r}
+classifier_df$status<-as.factor(classifier_df$status)
+ggplot(classifier_df, aes(x=status, y=Sepal.Length,fill=status)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
+
+ggplot(classifier_df, aes(x=status, y=Sepal.Width,fill=status)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
+
+ggplot(classifier_df, aes(x=status, y=Petal.Length,fill=status)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
+
+ggplot(classifier_df, aes(x=status, y=Petal.Width,fill=status)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
+```
+
+```{r}
+out$tagged_cluster<-as.factor(out$tagged_cluster)
+ggplot(out, aes(x=tagged_cluster, y=Petal.Width,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
 ```

### 2022-09-07 hw2/submission.Rmd
final ques 3.1

@@
 
-```{r, include=FALSE}
+```{r}
 
@@
 
-2.  Choose i'th part as test data and train knn on the rest 4
+2.  Choose i'th part as test data and train knn model on the rest 4 parts
 
-    1.  Find the threshold which maximises AUC using ROC curve
+    1.  Find the optimal threshold using the ROC curve
 
@@
 
-4.  When all parts are rotated into test set, calculate the measure for accuracy of model:
+4.  When all parts are rotated into the test set, calculate the measure for accuracy of model:
 
@@
 
-    2.  Min test accuracy across the 5 folds
+    2.  Min test accuracy across the 5 folds (optional)
 
@@
 
-The above summary shows the mean of R1 and relative size of each fold. Ideally for all the folds this should be 45% mean and 20%, but we see minor deviations due to random sampling.
+The above summary shows the mean of R1 and relative size of each fold.
+Ideally for all the folds this should be 45% mean and 20% proportion, but we see minor deviations in individual folds due to random sampling.
 
 ```{r}
-#Function to train model and return prediction on test
+#Function to train model and return prediction probabilities on test
 train_test_i = function(df_train,df_test, k0 = 5){
@@
   return (fitted(model))
+}
+
+# decide best threshold
+best_threshold = function(pred_prob, df_test=org_cc_data){
+  myroc <- roc(df_test$R1,as.numeric(unlist(pred_prob)), smooth=FALSE)
+  print(paste("AUC is",auc(myroc)))
+  #plot(myroc, auc.polygon=TRUE)
+  return (coords(myroc,x="best")["threshold"][[1]])
+}
+
+#Accuracy for best threshold
+acc_func = function(y_act, pred_probs, thresh = 0.5){
+  y_pred = lapply(pred_probs, function(x) if(x>=thresh) 1 else 0)
+  y_pred = as.numeric(unlist(y_pred))
+  acc_overall = sum(y_act == y_pred)/length(y_act)
+  acc_overall = round(acc_overall,4)* 100
+  return (acc_overall)
 }
@@
 
-### Step 4: Iterate on different models and use CV for best model
+### Step 4: Iterate on different models(i.e on different values of k) and use CV for best model
 
-```{r, include=FALSE}
+Here we differentiate two knn models, only using the k hyperparameter
+
+```{r, results='hide'}
 #Iterate over k and get mean test accuracy
 
-end_k = 77 #half of test set size
+end_k = 77 #almost half of test set size
 
@@
 
-We have also kept standard deviation, such that we can able to infer the variance of accuracy across cross folds.
-This is helpful because the maximum CV accuracy for different values across k, varies within a percent (See graph below).
+We have also kept standard deviation, such that we will be able to infer the variance of accuracy across cross folds.
+This is helpful because the maximum CV accuracy for different values across k varies in a very close interval (around 87.5 %) -\> see graph below
 
-#### Plot the accuracy with k, and find optimal k which maximizes accuracy
+#### Plot the accuracy with k (hyperparameter of knn), and find optimal k which maximizes accuracy
 
@@
   geom_vline(xintercept = acc_df[which.max(acc_df$mean_cv_accuracy),1],
-             color ="magenta", size = 1.5, alpha = 0.5) +
+             color ="blue", size = 1.5, alpha = 0.5) +
   ggtitle(paste("max accuracy for k =",acc_df[which.max(acc_df$mean_cv_accuracy),1]))
@@
 
-However, as noted above the nuance to note her is that, for k \> 30, that accuracy mostly oscillates at almost the same range with delta of 0.5.
+However, as noted above the nuance to note here is that, for k \> 30, accuracy mostly oscillates at almost the same range with delta of 0.5.
 But since the standard deviation of max accuracy among cross folds is 2.45 itself, this choice of k becomes a little tricky.
-In other words, we
+In other words, any k value above 30 seems appropriate, as the performance separation between the models is minimal
 
@@
 
-### Part 2: Train, Validation and Test on knnn
+### Part 2: Train, Validation and Test on k-knn
 
 ```{r}
+set.seed(77)
+
 #split data into 70% train, 15% validation and 15% test
@@
 
-```{r, include=FALSE}
+```{r, results='hide'}
 acc_df2 <- data.frame(matrix(ncol = 0, nrow = length(seq(3,end_k,2))))
@@
 
-Here we see that for k = 23, we get maximum average test accuracy of 84%.
+Note, that both the optimal k value here and the optimal accuracy value are different than part 1 above.
 
-Note, that both the optimal k value here and the optimal accuracy value are different than part 1 above.
-The test accuracy here is lower at 84% compared to 87% of accuracy in CV case, which is expected, since 84% is accuracy at true test, while cv accuracy is averaged across all the folds.
+The test accuracy here is the true (or best available) evaluation of the model chosen by us.
+
+However one might also argue that the optimal k values from both the approaches lie in a similar range, since any k value \>30 can be chosen from cross validation too.

### 2022-09-07 hw2/anshit/Homework_2.Rmd
Analysis of K-Means

@@
 date: "2022-09-04"
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
 
-```{r}
+```{r, include=FALSE}
 # imports
@@
 
-## Question 3.1
-
-```{r}
-cc_data = read.table('./data 3.1/credit_card_data-headers.txt', header = TRUE, sep='', dec='.')
-head(cc_data)
-dim(cc_data)
-```
-
 ## Question 4.2
 
-```{r}
+```{r, results='hide'}
 iris_data = read.table('./iris.txt', header = TRUE, sep='', dec='.')
@@
 my_plots = lapply(colnames(iris_data), function(x){
-  p = ggplot(iris_data) + aes_string(x)
-  if(x=='Species') {
-    p = p + geom_bar()
-  }
-  else {
-    p = p + geom_density()
+  p = ggplot(iris_data, aes(fill = Species)) + aes_string(x)
+  if(x!='Species') {
+    p = p + geom_density(alpha=.34)
   }
@@
 
+From the above distribution, we can make the following inferences: 1.
+Petal.Length and Petal.Width provide clear separation between the setosa
+vs the other classes. In clustering we expect these variables to be
+significantly lower in the cluster dominated by the 'setosa' classes. 2.
+We observe that while Sepal.Width and Sepal.Length have different
+central tendencies, they also show high degree of overlap between the
+different classes at the fringes.
+
 ```{r}
@@
 cor_mat = round(cor(iris_data[,1:4]), 2)
+cor_mat[upper.tri(cor_mat)] <- NA
 melted_mat = melt(cor_mat)
@@
 ```
+
+From the above matrix we observed the following: 1. Petal Length and
+Width are highly correlated 2. Sepal Length and Petal Width are highly
+correlated also 3. Sepal length and Sepal Width show negligible
+dependence 4. Sepal width is negatively correlated petal length and
+width
 
@@
 ```
+
+From the above graph, we can see that increasing the value of K beyond 4
+yields miminal benefit in clustering.
+
 ```{r}
@@
 ```
+
 ```{r}
@@
 ```
+
 ```{r}
-boxplot(Sepal.Length~tagged_cluster,data=out)
+cm <- table(iris_data$Species, best_km$cluster)
+cm
 ```
+
+From the confusion matrix, we can infer the following: \* Cluster 3
+shows highest proportion of setosa class. In fact it encompasses 100% of
+that class. \* Cluster 2 has the highest proportion of versicolor class.
+\* Cluster 1 has the highest proportion of virginia class.
+
+However in comparison to cluster 3 which is purely setosa; clusters 1
+and 2 have a significant no. of data points of a different class. In
+other words, there is confusion between classes in clusters 1 and 2.
 
@@
 out$tagged_cluster<-as.factor(out$tagged_cluster)
-ggplot(out, aes(x=tagged_cluster, y=Petal.Width,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
+ggplot(out, aes(x=tagged_cluster, y=Petal.Width,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none") + xlab('Cluster')
 ```
-Petal width and Petal Length are strong indicators for one class (setosa)
 
-```{r}
-cm <- table(iris_data$Species, best_km$cluster)
-cm
-```
+-   As we expected from the density plots and confusion matrix, Petal
+    width and Petal Length are significant discriminator for cluster 3,
+    which is dominated by setosa class.
+-   While the other two variables provide relatively lower
+    discrimination between the clusters.
 
@@
 ```
+
+From the above plots, we see: \* Distribution of sepal length is very
+similar for improperly classified versicolor and proper virginica.
+Similarly, the distribution of improperly classified virginica and
+proper versicolor are very similar too. \* The same effect is apparent
+in Sepal width. \* This apparent similarity can be a probable reason for
+clustering to confuse between the classes.
 
@@
 ```
+
+-   It can be seen from the plot above that cluster 1 and 2 are closely
+    knit (i.e. intra cluster separation is less) and hence we have the
+    confusion in clustering some points for these 2 clusters . Cluster 3
+    has a better separability than the other 2 and hence it is a good
+    demarcated cluster.
+
+-   It is also apparent that cluster 3 has higher spread and hence can
+    be split up into 2 clusters depending on the use case which also
+    came up in our previous study (elbow graph to decide between k=3 and
+    k=4)
+
 ***Analysing For K=4***
@@
 ```{r}
-plot(scale(iris_data[,1:4]), col=best_km$cluster)
+plot(scale(iris_data[,1:4]), col=alt_kmeans$cluster)
 ```
@@
 ```{r}
-fviz_cluster(best_km, data=scale(iris_data[,1:2]), geom=c("point"), ellipse.type = "euclid")
+fviz_cluster(alt_kmeans, data=scale(iris_data[,1:2]), geom=c("point"), ellipse.type = "euclid")
 ```
+
+### How well the clustering fits the flower type ?
+
+```{r}
+cm <- table(iris_data$Species, best_km$cluster)
+cm
+```
+
+As seen from above the total accuracy of the clustering model is:
+
+```{r}
+1-(((11+14)/nrow(iris_data)))
+```
+
+### Accuracy in individual classes
+
+If cluster 1 is virginica, cluster 2 is versicolor, cluster 3 is setosa
+
+```{r}
+acc_versicolor = 1 - (11/50)
+acc_virginica = 1 - (14/50)
+acc_setosa = 1- (0/50)
+
+print (paste("Acc for versicolor",acc_versicolor))
+print (paste("Acc for virginica",acc_virginica))
+print (paste("Acc for setosa",acc_setosa))
+```
+
+Thus we have 100% accuracy for setosa, while for the other two flowers
+we have accuracy in 70's.

### 2022-09-07 hw2/submission.Rmd
Analysis of K-Means

@@
 However one might also argue that the optimal k values from both the approaches lie in a similar range, since any k value \>30 can be chosen from cross validation too.
+
+## Question 4.1
+
+### Example 1: Market Segmentation/Retail Marketing
+
+Clustering models are frequently used for grouping customers with similar attributes
+
+**Objective Function (**$Y$): **Predictor Variables (**$X_i$):
+
+-   ***Household Income***:
+-   ***Household Size***:
+-   ***Distance From Nearest Urban Area***:
+-   ***Head of Houehold Occupation***:
+
+## Question 4.2
+
+```{r, results='hide'}
+iris_data = read.table('./iris.txt', header = TRUE, sep='', dec='.')
+head(iris_data,50)
+dim(iris_data)
+```
+
+### Exploratory Data Analysis
+
+```{r}
+summary(iris_data)
+```
+
+```{r}
+table(iris_data$Species)
+```
+
+```{r}
+# Density Plots
+my_plots = lapply(colnames(iris_data), function(x){
+  p = ggplot(iris_data, aes(fill = Species)) + aes_string(x)
+  if(x!='Species') {
+    p = p + geom_density(alpha=.34)
+  }
+})
+
+plot_grid(plotlist = my_plots)
+```
+
+[**#Analysis**]{style="color:red"} From the above distribution, we can make the following inferences: 1.
+Petal.Length and Petal.Width provide clear separation between the setosa vs the other classes.
+In clustering we expect these variables to be significantly lower in the cluster dominated by the 'setosa' classes.
+2.
+We observe that while Sepal.Width and Sepal.Length have different central tendencies, they also show high degree of overlap between the different classes at the fringes.
+
+```{r}
+# Correlation Matrix
+cor_mat = round(cor(iris_data[,1:4]), 2)
+cor_mat[upper.tri(cor_mat)] <- NA
+melted_mat = melt(cor_mat)
+ggplot(melted_mat, aes(x=Var1, y=Var2, fill=value)) +
+  geom_tile() + geom_text(aes(Var1, Var2, label=value), color='white', size=3)
+```
+
+[**#Analysis**]{style="color:red"} From the above matrix we observed the following: 1.
+Petal Length and Width are highly correlated 2.
+Sepal Length and Petal Width are highly correlated also 3.
+Sepal length and Sepal Width show negligible dependence 4.
+Sepal width is negatively correlated petal length and width
+
+### Modelling
+
+```{r}
+# Calculation of best 'K' by elbow graph for total sum of squares within
+elbow_graph_data <- data.frame(K=integer(), within_sum_of_square=double())
+for(k in seq(1, 10, by=1)){
+  km.res <- kmeans(as.matrix(scale(iris_data[,1:4])), k, nstart = 25)
+  elbow_graph_data[nrow(elbow_graph_data) + 1, ] = c(k, km.res$tot.withinss)
+}
+plot(elbow_graph_data, type='b', ylab='Total Sum of Square Within', main='Elbow Graph', col='blue')
+```
+
+[**#Analysis**]{style="color:red"} \* From the above graph, we can see that increasing the value of 'K' beyond 4 yields very low marginal benefit.
+\* Both 3 and 4 appear to be good values of 'K' and can be used depending on the use case.
+In this example, since we know that there are 3 classes, we expect 3 to be a more appropriate choice.
+
+```{r}
+# Other Method for determining best 'K' by sum of intra-cluster distance
+bw_graph_data <- data.frame(K=integer(), between_sum_of_square=double())
+for(k in seq(1, 10, by=1)){
+  km.res <- kmeans(as.matrix(scale(iris_data[,1:4])), k, nstart = 25)
+  bw_graph_data[nrow(bw_graph_data) + 1, ] = c(k, round(km.res$betweenss,2))
+}
+plot(bw_graph_data, type='b', ylab='Betweeen Sum of Square (Intra Cluster)', main='Betweeen Sum of Square (Intra Cluster) Graph', col='green')
+```
+
+[**#Analysis**]{style="color:red"} \* Similarly to elbow-graph, this plot also shows 3 and 4 to be good choices for 'K'.
+
+```{r, include=FALSE}
+best_km <- kmeans(as.matrix(scale(iris_data[,1:4])), 3, nstart = 25)
+out<-cbind(iris_data,tagged_cluster=best_km$cluster)
+out$tagged_cluster<-as.factor(out$tagged_cluster)
+
+getBoxPlot <- function(df, y_val){
+  ggplot(df, aes(x=tagged_cluster, y=y_val,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")  + xlab('Cluster')
+}
+```
+
+```{r}
+confusion_matrix <- table(iris_data$Species, best_km$cluster)
+confusion_matrix
+```
+
+[**#Analysis**]{style="color:red"} From the confusion matrix, we can infer the following: \* Cluster 3 shows highest proportion of setosa class.
+In fact it encompasses 100% of that class.
+\* Cluster 2 has the highest proportion of versicolor class.
+\* Cluster 1 has the highest proportion of virginia class.
+
+However in comparison to cluster 3 which is purely setosa; clusters 1 and 2 have a significant no.
+of data points of a different class.
+In other words, there is confusion between classes in clusters 1 and 2.
+
+```{r}
+getBoxPlot(out, 'Sepal.Length')
+getBoxPlot(out, 'Sepal.Width')
+getBoxPlot(out, 'Petal.Length')
+getBoxPlot(out, 'Petal.Width')
+```
+
+[**#Analysis**]{style="color:red"} \* As we expected from the density plots and confusion matrix, Petal width and Petal Length are significant discriminator for cluster 3, which is dominated by setosa class.
+\* While the other two variables provide relatively lower discrimination between the clusters.
+
+### Analysis of the misclassified points
+
+```{r}
+classifier_df = out[!out$Species=='setosa',]
+classifier_df$status = ''
+classifier_df$status[(classifier_df$tagged_cluster=="1" & classifier_df$Species=="versicolor")] = 'improper_versicolor'
+classifier_df$status[(classifier_df$tagged_cluster=="2" & classifier_df$Species=="virginica")] = 'improper_virginica'
+classifier_df$status[(classifier_df$tagged_cluster=="2" & classifier_df$Species=="versicolor")] = 'proper_versicolor'
+classifier_df$status[(classifier_df$tagged_cluster=="1" & classifier_df$Species=="virginica")] = 'proper_virginica'
+classifier_df$status<-as.factor(classifier_df$status)
+```
+
+```{r}
+getBoxPlot(classifier_df, 'Sepal.Length')
+getBoxPlot(classifier_df, 'Sepal.Width')
+getBoxPlot(classifier_df, 'Petal.Length')
+getBoxPlot(classifier_df, 'Petal.Width')
+```
+
+[**#Analysis**]{style="color:red"}
+From the above plots, we see:
+* Distribution of sepal length is very similar for improperly classified versicolor and proper virginica.
+Similarly, the distribution of improperly classified virginica and proper versicolor are very similar too.
+* The same effect is apparent in Sepal width.
+* This apparent similarity can be a probable reason for clustering to confuse between the classes.
+
+```{r}
+# Centroids of Clusters
+kmeans_basic_table <- data.frame(best_km$size, best_km$centers)
+kmeans_basic_table
+```
+
+### Plotting Clusters
+
+```{r}
+fviz_cluster(best_km, data=scale(iris_data[,1:2]), geom=c("point"), ellipse.type = "euclid")
+```
+
+[**#Analysis**]{style="color:red"}
+* It can be seen from the plot above that cluster 1 and 2 are closely knit (i.e.
+intra cluster separation is less) and hence we have the confusion in clustering some points for these 2 clusters .
+Cluster 3 has a better separability than the other 2 and hence it is a good demarcated cluster.
+* It is also apparent that cluster 3 has higher spread and hence can be split up into 2 clusters depending on the use case which also came up in our previous study (elbow graph to decide between k=3 and k=4)
+
+### Just a visualization with 4 Cluster Space to highlight the above point
+
+```{r}
+alt_kmeans <- kmeans(as.matrix(scale(iris_data[,1:4])), 4, nstart = 25)
+alt_kmeans
+fviz_cluster(alt_kmeans, data=scale(iris_data[,1:2]), geom=c("point"), ellipse.type = "euclid")
+```
+
+### Accuracy in individual classes
+
+If cluster 1 is virginica, cluster 2 is versicolor, cluster 3 is setosa
+
+```{r}
+acc_versicolor = 1 - (11/50)
+acc_virginica = 1 - (14/50)
+acc_setosa = 1- (0/50)
+
+print (paste("Acc for versicolor",acc_versicolor))
+print (paste("Acc for virginica",acc_virginica))
+print (paste("Acc for setosa",acc_setosa))
+```
+
+[**#Analysis**]{style="color:red"}
+Thus we have 100% accuracy for setosa, while for the other two flowers
+we have accuracy in 70's.

### 2022-09-07 hw2/submission.Rmd
Final Submission

@@
 ---
-title: "ISYE 6501 | Submission HW2"
+title: "ISYE 6501"
 output:
@@
 
-# ISYE 6501 \| Fall 22
+# Submission HW2 \| Fall 22
 
@@
 library(reshape2)
-#library(factoextra)
+library(factoextra)
 library(kknn)
@@
 
-```{r}
+```{r, warning=FALSE, message=FALSE}
 #Test function with a sample k
@@
 
-```{r, results='hide'}
+```{r, results='hide', message=FALSE, warning=FALSE}
 #Iterate over k and get mean test accuracy
@@
 
-## Question 3.1
-
-### Part 2: Train, Validation and Test on k-knn
+### Question 3.1, Part 2: Train, Validation and Test on k-knn
 
@@
 
-```{r, results='hide'}
+```{r, results='hide', warning=FALSE, message=FALSE}
 acc_df2 <- data.frame(matrix(ncol = 0, nrow = length(seq(3,end_k,2))))
@@
 
-[**#Analysis**]{style="color:red"} From the above distribution, we can make the following inferences: 1.
+[**#Analysis**]{style="color:red"}
+
+From the above distribution, we can make the following inferences: 1.
 Petal.Length and Petal.Width provide clear separation between the setosa vs the other classes.
@@
 
-[**#Analysis**]{style="color:red"} From the above matrix we observed the following: 1.
+[**#Analysis**]{style="color:red"}
+
+From the above matrix we observed the following: 1.
 Petal Length and Width are highly correlated 2.
@@
 
-[**#Analysis**]{style="color:red"} \* From the above graph, we can see that increasing the value of 'K' beyond 4 yields very low marginal benefit.
+[**#Analysis**]{style="color:red"}
+
+\* From the above graph, we can see that increasing the value of 'K' beyond 4 yields very low marginal benefit.
 \* Both 3 and 4 appear to be good values of 'K' and can be used depending on the use case.
@@
 
-[**#Analysis**]{style="color:red"} \* Similarly to elbow-graph, this plot also shows 3 and 4 to be good choices for 'K'.
+[**#Analysis**]{style="color:red"}
+
+\* Similarly to elbow-graph, this plot also shows 3 and 4 to be good choices for 'K'.
 
@@
 
-[**#Analysis**]{style="color:red"} From the confusion matrix, we can infer the following: \* Cluster 3 shows highest proportion of setosa class.
+[**#Analysis**]{style="color:red"}
+
+From the confusion matrix, we can infer the following: \* Cluster 3 shows highest proportion of setosa class.
 In fact it encompasses 100% of that class.
@@
 ```{r}
-getBoxPlot(out, 'Sepal.Length')
-getBoxPlot(out, 'Sepal.Width')
-getBoxPlot(out, 'Petal.Length')
-getBoxPlot(out, 'Petal.Width')
+ggplot(out, aes(x=tagged_cluster, y=Sepal.Length,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")  + xlab('Cluster')
+ggplot(out, aes(x=tagged_cluster, y=Sepal.Width,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")  + xlab('Cluster')
+ggplot(out, aes(x=tagged_cluster, y=Petal.Length,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")  + xlab('Cluster')
+ggplot(out, aes(x=tagged_cluster, y=Petal.Width,fill=tagged_cluster)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")  + xlab('Cluster')
 ```
 
-[**#Analysis**]{style="color:red"} \* As we expected from the density plots and confusion matrix, Petal width and Petal Length are significant discriminator for cluster 3, which is dominated by setosa class.
+[**#Analysis**]{style="color:red"}
+
+\* As we expected from the density plots and confusion matrix, Petal width and Petal Length are significant discriminator for cluster 3, which is dominated by setosa class.
 \* While the other two variables provide relatively lower discrimination between the clusters.
@@
 ```{r}
-getBoxPlot(classifier_df, 'Sepal.Length')
-getBoxPlot(classifier_df, 'Sepal.Width')
-getBoxPlot(classifier_df, 'Petal.Length')
-getBoxPlot(classifier_df, 'Petal.Width')
+ggplot(classifier_df, aes(x=status, y=Sepal.Length,fill=status)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
+ggplot(classifier_df, aes(x=status, y=Sepal.Width,fill=status)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
+ggplot(classifier_df, aes(x=status, y=Petal.Length,fill=status)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
+ggplot(classifier_df, aes(x=status, y=Petal.Width,fill=status)) + geom_boxplot(alpha=0.3) +theme(legend.position="none")
 ```
@@
 [**#Analysis**]{style="color:red"}
-From the above plots, we see:
-* Distribution of sepal length is very similar for improperly classified versicolor and proper virginica.
+
+From the above plots, we see: \* Distribution of sepal length is very similar for improperly classified versicolor and proper virginica.
 Similarly, the distribution of improperly classified virginica and proper versicolor are very similar too.
-* The same effect is apparent in Sepal width.
-* This apparent similarity can be a probable reason for clustering to confuse between the classes.
+\* The same effect is apparent in Sepal width.
+\* This apparent similarity can be a probable reason for clustering to confuse between the classes.
 
@@
 [**#Analysis**]{style="color:red"}
-* It can be seen from the plot above that cluster 1 and 2 are closely knit (i.e.
+
+\* It can be seen from the plot above that cluster 1 and 2 are closely knit (i.e.
 intra cluster separation is less) and hence we have the confusion in clustering some points for these 2 clusters .
 Cluster 3 has a better separability than the other 2 and hence it is a good demarcated cluster.
-* It is also apparent that cluster 3 has higher spread and hence can be split up into 2 clusters depending on the use case which also came up in our previous study (elbow graph to decide between k=3 and k=4)
+\* It is also apparent that cluster 3 has higher spread and hence can be split up into 2 clusters depending on the use case which also came up in our previous study (elbow graph to decide between k=3 and k=4)
 
@@
 [**#Analysis**]{style="color:red"}
-Thus we have 100% accuracy for setosa, while for the other two flowers
-we have accuracy in 70's.
+
+Thus we have 100% accuracy for setosa, while for the other two flowers we have accuracy in 70's.

### 2022-09-07 hw2/submission.Rmd

@@
 
-```{r}
+```{r, warning=FALSE, message=FALSE}
 
@@
 
-Clustering models are frequently used for grouping customers with similar attributes
+Clustering models are used for grouping customers with similar attributes and features into buckets.
+A typical use case of this approach is to segment the customers into groups, such that they can be better targeted for a spectrum of business activities like targeted advertising.
+Segmenting the market helps the business better spend their money in the right way.
+One real life example is from Google and how it uses various predictors to cluster users on various factors to show targeted advertisement.
 
-**Objective Function (**$Y$): **Predictor Variables (**$X_i$):
+**Predictor Variables (**$X_i$): Some of the key predictors are given below:
 
--   ***Household Income***:
--   ***Household Size***:
--   ***Distance From Nearest Urban Area***:
--   ***Head of Houehold Occupation***:
+-   ***Search History***: This variable tracks user's search history and helps in identifying what the user might me interested in.
+It can also help identify what the user wants in the market, for example if they are interested in the fuel efficiency or model or performance of a car.
+
+-   ***Personal User Data***: User's personal data plays a key role in clustering user based on their age, gender, etc.
+Showing ads which are relevant to he user helps in creating a better user experience.
+
+-   ***Links Clicked***: This is a measure of user's interaction with links available on the webpage.
+It helps with analyzing user behavior which helps in clustering on the basis of activity.
+Suppose a user searches for car but only interacts with links which lead to results from a certain brand of car.
+Determining this behavior will help in showing targeted ads from that brand to that particular user.
+
+-   ***User Location***: This is a user's location.
+This helps with serving geographical appropriate ads.
+It would lead to bad user experience if a user from India is shown an ad for a product which is available in USA.
 

### 2022-09-07 hw2/submission.Rmd
correction ques 3.1

@@
 
-1.  Divide the data into 5 equal parts
+0.  Divide the data into train,test (where test is final data used for evaluation)
+
+1.  Divide the combined train and validation data into 5 equal parts
 
@@
 
+### Step 0:
+
+```{r}
+set.seed(77)
+
+#split data into 70% train, 15% validation and 15% test
+df_copy = org_cc_data
+df_copy$split_i <- sample(100, size = nrow(df_copy), replace = TRUE)
+
+df_copy$split = "train"
+df_copy$split[70 < df_copy2$split_i] = "valid"
+df_copy$split[85 < df_copy2$split_i] = "test"
+
+summary_df = aggregate(df_copy,list(df_copy$split), 
+          FUN = function(x) c(mean = mean(x), proportion = length(x)/nrow(df_copy)))
+print.data.frame(summary_df)
+```
+
+```{r}
+df_train = df_copy[df_copy$split=="train",]
+dim(df_train)
+df_valid = df_copy[df_copy$split=="valid",]
+dim(df_valid)
+df_test= df_copy[df_copy$split=="test",]
+dim(df_test)
+```
+
 ### Step 1:
@@
 ```{r}
-#Split org_data into 5 parts
-df_copy = org_cc_data
+#Split the combined (train and validation) into 5 parts for CV
+df_cv = rbind(df_train,df_valid)
+dim(df_cv)
+df_copy = df_cv
 df_copy$cross_validation_i <- sample(5, size = nrow(df_copy), replace = TRUE)
@@
 
-### Question 3.1, Part 2: Train, Validation and Test on k-knn
-
-```{r}
-set.seed(77)
-
-#split data into 70% train, 15% validation and 15% test
-df_copy2 = org_cc_data
-df_copy2$split_i <- sample(100, size = nrow(df_copy), replace = TRUE)
-
-df_copy2$split = "train"
-df_copy2$split[70 < df_copy2$split_i] = "valid"
-df_copy2$split[85 < df_copy2$split_i] = "test"
-
-summary_df = aggregate(df_copy2$R1,list(df_copy2$split), 
-          FUN = function(x) c(mean = mean(x), proportion = length(x)/nrow(df_copy2)))
-print.data.frame(summary_df)
-```
-
-Again slight deviation from expected values is seen due to random effects
-
-```{r}
-df_train = df_copy2[df_copy2$split=="train",]
-dim(df_train)
-df_valid = df_copy2[df_copy2$split=="valid",]
-dim(df_valid)
-df_test= df_copy2[df_copy2$split=="test",]
-dim(df_test)
-```
-
-### Find optimal k using training on train and validation
-
 ```{r}
@@
 
-#### Run the above function for variety of k values
+#### Rebuild model with optimal k, with
+
+-   Train = Train + Validation
+
+-   Test = Test
+
+```{r}
+k_optimal = acc_df2[which.max(acc_df$mean_cv_accuracy),1]
+acc_optimal_k = train_valid_func(df_train = df_copy, df_test = df_test, k_choice = k_optimal)
+print (paste("With optimal k value, Accuracy on test =",acc_optimal_k))
+```
+
+The above accuracy is a true(or best) measure/evaluation for our model
+
+### Question 3.1, Part 2: Train, Validation and Test
+
+### Find optimal k using training on train and validation
+
+#### Run the train valid func function for variety of k values
 

### 2022-09-07 hw2/submission.Rmd
Final changes pushed

@@
 
-**We repeat the above 4 steps for different k values, to find the best model(or choice of k)**
+**We repeat the above 1-4 steps for different k values, to find the best model(or choice of k)**
 
@@
 df_copy$split = "train"
-df_copy$split[70 < df_copy2$split_i] = "valid"
-df_copy$split[85 < df_copy2$split_i] = "test"
+df_copy$split[70 < df_copy$split_i] = "valid"
+df_copy$split[85 < df_copy$split_i] = "test"
 
-summary_df = aggregate(df_copy,list(df_copy$split), 
+summary_df = aggregate(df_copy$R1,list(df_copy$split), 
           FUN = function(x) c(mean = mean(x), proportion = length(x)/nrow(df_copy)))
@@
 We have also kept standard deviation, such that we will be able to infer the variance of accuracy across cross folds.
-This is helpful because the maximum CV accuracy for different values across k varies in a very close interval (around 87.5 %) -\> see graph below
+This is helpful because the maximum CV accuracy for different values across k varies in a very close interval (around 86.5 %) -\> see graph below
 
@@
 
-Here we see that for k = 47, we get maximum average CV accuracy of 87.74%.
+Here we see that for k = 33, we get maximum average CV accuracy of 87.18%.
 
-However, as noted above the nuance to note here is that, for k \> 30, accuracy mostly oscillates at almost the same range with delta of 0.5.
-But since the standard deviation of max accuracy among cross folds is 2.45 itself, this choice of k becomes a little tricky.
-In other words, any k value above 30 seems appropriate, as the performance separation between the models is minimal
+However, as noted above the nuance to note here is that, for k \> 25, accuracy mostly oscillates at almost the same range with delta of 1%.
+But since the standard deviation of max accuracy among cross folds is 4.47 itself, this choice of k becomes a little tricky.
+In other words, any k value above 25 seems appropriate, as the performance separation between the models is minimal
 
@@
 
-#### Rebuild model with optimal k, with
+#### Rebuild model with optimal k selected, for true test accuracy
 
@@
 ```{r}
-k_optimal = acc_df2[which.max(acc_df$mean_cv_accuracy),1]
+k_optimal = acc_df[which.max(acc_df$mean_cv_accuracy),1]
 acc_optimal_k = train_valid_func(df_train = df_copy, df_test = df_test, k_choice = k_optimal)
@@
 
-### Find optimal k using training on train and validation
+### Find optimal k using training on train and maximising accuracy on validation
 
-#### Run the train valid func function for variety of k values
+#### Run the train_valid_func function for variety of k values
 
@@
 
-However one might also argue that the optimal k values from both the approaches lie in a similar range, since any k value \>30 can be chosen from cross validation too.
+However one might also argue that the optimal k values from both the approaches lie in a similar range, since any k value \>25 can be chosen from cross validation too.
 

### 2022-09-12 hw3-1/anshit/HW_3.Rmd
Added HW3 - Initial Commit

---
name: "HW 3 Assignment"
author: "Anshit Verma"
date: "12/09/2022"
output: "html_document"
---

### Question 6.2

```{r}
data = read.table('./temps.txt', sep="\t", dec=".", header = TRUE)
data
```

```{r}
summary(data$X1996)
sd(data$X1996)
```

```{r}
mean = 83.72
st = 0
x = 8
st_df = data.frame(i=double(), st=double())
for(i in 1:nrow(data)){
  st = max(0, st + (mean - data$X1996[i] - x))
  st_df[nrow(st_df)+1,] = c(i, st)
}
plot(st_df, xlab='Day', ylab='St', type='o')
```

```{r}
yearly_mean = data.frame(DAY=data[,1], YearlyMean=rowMeans(data[,-1]))
yearly_mean
```

```{r}
summary(yearly_mean$YearlyMean)
```

```{r}
mean = mean(yearly_mean$YearlyMean)
st = 0
x = 8
st_df = data.frame(i=double(), st=double())
for(i in 1:nrow(yearly_mean)){
  st = max(0, st + (mean - yearly_mean$YearlyMean[i] - x))
  st_df[nrow(st_df)+1,] = c(i, st)
}
plot(st_df, xlab='Day', ylab='St', type='o')
```

# Inferences - 
1. Comparing the above plot with 1996, days when the temp is rising has shifted towards later part of the year, it suggests a change in season cycle.
2. Our max value of our running sum has gotten lower which also suggests warm season in later year.
3. Range of temp values has decreased (min has inc by 8 units) which has shifted temp towards mean.

# TODO - 
1. C and T hyperparameter - how to decide on values.
2. Dividing in buckets to see better changes in YoY temps.
Reference - https://rpubs.com/ssufian/658130

### 2022-09-12 hw2/submission.Rmd

@@
 
-### Part 1: Cross Validation for finding best knnn model
+### Part 1: Cross Validation for finding best knn model
 
@@
 
-### Exploratory Data Analysis
+### Exploratory Data Analysis (for getting the best predictor combinations)
 
@@
 In this example, since we know that there are 3 classes, we expect 3 to be a more appropriate choice.
+(\*Best value of k for our analysis\*)
 
@@
 \* Cluster 2 has the highest proportion of versicolor class.
-\* Cluster 1 has the highest proportion of virginia class.
+\* Cluster 1 has the highest proportion of virginica class.
 
@@
 
-### Accuracy in individual classes
+### Accuracy in individual classes (to see how well clustering predicts flower type)
 

### 2022-09-12 hw3-1/ashish/hw3_test.Rmd
ad sample

---
title: "HW3, Ashish EDA"
output: html_notebook
---

```{r}
data = read.table('../temps.txt', sep="\t", dec=".", header = TRUE)
data
```

```{r}
data_celsius = 
```

```{r}
CUSUM_ad = function(x,days = data$DAY ,C = 0.30,T = 4){
  mean_x = mean(x)
  sd_x = sd(x)
  C = C * sd_x
  #apply cusum
  x1 = lapply(x, function(xi)  (mean_x - xi - C))
  st_df = data.frame(st=double())
  st = 0
  for(i in 1:length(x)){
    #print (paste(i,class(st),class(x1[[i]])))
    st = max(0, st + x1[[i]])
    st_df[i,] = c(st)
  }
  #Find index where St > T
  st_df$Day = days
  st_df["anamoly"] = 0
  st_df$anamoly[st_df$st > T] = 1
  st_df$xt = x
  #plot x with T and if decrease see
  st_df
}
```

```{r}
check = CUSUM_ad(data$X1996)
#library(ggplot2)
par(mfrow=c(2,1))
plot(check$x)
plot(check$st)
#ggplot(check, aes(x = Day, y = st)) + geom_point() + geom_line() +
#  theme(axis.text.x = element_text(angle = 90))
```

```{r}

```

```{r}

```

```{r}

```

### 2022-09-13 hw3-1/ashish/hw3_test.Rmd
draft

@@
 title: "HW3, Ashish EDA"
-output: html_notebook
+output:
+  pdf_document: default
+  html_notebook: default
 ---
@@
 ```{r}
+library(grid)
+
 data = read.table('../temps.txt', sep="\t", dec=".", header = TRUE)
@@
 ```{r}
-data_celsius = 
-```
-
-```{r}
-CUSUM_ad = function(x,days = data$DAY ,C = 0.30,T = 4){
+CUSUM_ad = function(x, year, days = data$DAY ,C = 0.30,T = 4, if_c_relative = FALSE){
   mean_x = mean(x)
   sd_x = sd(x)
+  if (if_c_relative) {
     C = C * sd_x
+  }
   #apply cusum
@@
   #plot x with T and if decrease see
-  st_df
+  par(mfrow=c(2,1))
+  plot(st_df$x)
+  title(paste("Temperature (in Celsius) and S_t (bottom) for year",year), line = -1, outer = TRUE)
+  plot(st_df$st)
+  pushViewport(viewport())
+  grid.lines(x = c(0,1), y = grconvertY(T, "user", "ndc"), gp = gpar(col = "red"))
+  change_detected = which.max(st_df$st>T)
+  grid.lines(x = grconvertX(change_detected, "user", "ndc"), y = c(0,1), gp = gpar(col = "red"))
+  popViewport()
+  #return first point where St touches T
+  change_length = length(days)-change_detected
+  return (c(as.numeric(year),as.numeric(change_detected),as.numeric(change_length)))
 }
@@
 ```{r}
-check = CUSUM_ad(data$X1996)
+CUSUM_ad(data$X1996,"1996", C= 5, T = 50)
 #library(ggplot2)
-par(mfrow=c(2,1))
-plot(check$x)
-plot(check$st)
+#par(mfrow=c(2,1))
+#plot(check$x)
+#plot(check$st)
 #ggplot(check, aes(x = Day, y = st)) + geom_point() + geom_line() +
 #  theme(axis.text.x = element_text(angle = 90))
+```
+
+```{r}
+### Transform to celsius
+temp_cols = names(data)[!(names(data) %in% c("DAY"))]
+#Transform to celsius
+data_celsius = data
+data_celsius[temp_cols] <- lapply(data_celsius[temp_cols], function(f) (f-32)*5/9)
+```
+
+```{r}
+CUSUM_ad(data_celsius[,"X1996"],1996, C = 3, T = 25)
+```
+
+```{r}
+### Run Cusum over data_celsius for all the years
+mean(unlist(lapply(temp_cols, function(col) 
+  sd(data_celsius[,col])
+  )))
+```
+
+```{r}
+### Run Cusum over data_celsius for all the years
+winter_start_days = lapply(temp_cols, function(col) 
+  CUSUM_ad(x=data_celsius[,col], year = substr(col,2,5), C = 6, T = 25)
+  )
+
+# barplot(sapply(winter_start_days,"[[",1),
+# main = "Variation of Winter Start from 1996 to 2015",
+# xlab = "Year",
+# ylab = "Days since Jul 1, when winter change detected",
+# names.arg = temp_cols,
+# col = "darkred",
+# horiz = FALSE)
+```
+
+```{r}
+### Run Cusum over data_celsius for all the years
+winter_start_days = lapply(temp_cols, function(col) 
+  CUSUM_ad(x=data_celsius[,col], year = substr(col,2,5), C = 4, T = 25)
+  )
+```
+
+```{r}
+change_df <- as.data.frame(do.call(rbind, winter_start_days))
+colnames(change_df) = c("year","winter_start","winter_length")
+change_df["year_bucket"] = "a.1996_2000"
+change_df[change_df$year>=2000,"year_bucket"] = "b.2000_2004"
+change_df[change_df$year>=2004,"year_bucket"] = "c.2004_2008"
+change_df[change_df$year>=2008,"year_bucket"] = "d.2008_2012"
+change_df[change_df$year>=2012,"year_bucket"] = "e.2012_2016"
+change_df
+```
+
+```{r}
+aggregate(cbind(winter_start,winter_length) ~ year_bucket, change_df, mean)
+```
+
+```{r}
+
+```
+
+```{r}
+barplot(sapply(winter_start_days,"[[",2),
+main = "Variation of Winter Start from 1996 to 2015",
+xlab = "Year",
+ylab = "Days since Jul 1, when winter change detected",
+names.arg = temp_cols,
+col = "darkred",
+horiz = FALSE)
+```
+
+```{r}
+plot(sapply(winter_start_days,"[[",1))
+axis(1, at=1:length(temp_cols),labels=temp_cols, las=3)
 ```
@@
 ```
-
-```{r}
-
-```

### 2022-09-14 hw3-1/anshit/HW_3.Rmd

@@
 
+```{r}
+library(grid)
+```
+
 ### Question 6.2
@@
 ```{r}
-data = read.table('./temps.txt', sep="\t", dec=".", header = TRUE)
+data = read.table('../temps.txt', sep="\t", dec=".", header = TRUE)
 data
@@
 Reference - https://rpubs.com/ssufian/658130
+
+```{r}
+### Transform to celsius
+temp_cols = names(data)[!(names(data) %in% c("DAY"))]
+#Transform to celsius
+data_celsius = data
+data_celsius[temp_cols] <- lapply(data_celsius[temp_cols], function(f) (f-32)*5/9)
+```
+
+```{r}
+CUSUM_ad = function(x, year, days = data$DAY ,C = 0.30, T = 4, if_c_relative = FALSE){
+  mean_x = mean(x)
+  sd_x = sd(x)
+  if (if_c_relative) {
+    C = C * sd_x
+  }
+  #apply cusum
+  x1 = lapply(x, function(xi)  (mean_x - xi - C))
+  st_df = data.frame(st=double())
+  st = 0
+  for(i in 1:length(x)){
+    #print (paste(i,class(st),class(x1[[i]])))
+    st = max(0, st + x1[[i]])
+    st_df[i,] = c(st)
+  }
+  #Find index where St > T
+  st_df$Day = days
+  st_df["anamoly"] = 0
+  st_df$anamoly[st_df$st > T] = 1
+  st_df$xt = x
+  #plot x with T and if decrease see
+  par(mfrow=c(2,1))
+  plot(st_df$x)
+  title(paste("Temperature (in Celsius) and S_t (bottom) for year",year), line = -1, outer = TRUE)
+  plot(st_df$st)
+  pushViewport(viewport())
+  grid.lines(x = c(0,1), y = grconvertY(T, "user", "ndc"), gp = gpar(col = "red"))
+  change_detected = which.max(st_df$st>T)
+  grid.lines(x = grconvertX(change_detected, "user", "ndc"), y = c(0,1), gp = gpar(col = "red"))
+  popViewport()
+  #return first point where St touches T
+  change_length = length(days)-change_detected
+  return (c(as.numeric(year),as.numeric(change_detected),as.numeric(change_length)))
+}
+```
+
+```{r}
+### Run Cusum over data_celsius for all the years
+winter_start_days = lapply(temp_cols, function(col) 
+  CUSUM_ad(x=data_celsius[,col], year = substr(col,2,5), C = 4, T = 25)
+  )
+```
+
+```{r}
+change_df <- as.data.frame(do.call(rbind, winter_start_days))
+colnames(change_df) = c("year","winter_start","winter_length")
+change_df["year_bucket"] = "a.1996_2000"
+change_df[change_df$year>=2000,"year_bucket"] = "b.2000_2004"
+change_df[change_df$year>=2004,"year_bucket"] = "c.2004_2008"
+change_df[change_df$year>=2008,"year_bucket"] = "d.2008_2012"
+change_df[change_df$year>=2012,"year_bucket"] = "e.2012_2016"
+change_df
+```
+
+```{r}
+aggregate(cbind(winter_start,winter_length) ~ year_bucket, change_df, mean)
+```
+
+```{r}
+month_df <- data.frame(data)
+month_df["month"] = "month"
+month_df[grepl( "Jul", month_df$DAY), "month"] = "Jul"
+month_df[grepl( "Aug", month_df$DAY), "month"] = "Aug"
+month_df[grepl( "Sep", month_df$DAY), "month"] = "Sep"
+month_df[grepl( "Oct", month_df$DAY), "month"] = "Oct"
+month_df = month_df[,-1]
+for ( col in 1:ncol(month_df)){
+    colnames(month_df)[col] <-  sub("X", "", colnames(month_df)[col])
+}
+month_df
+```
+
+```{r}
+year_bucketing_function = function(df){
+  df["year_bucket"] = "a.1996_2000"
+  df[df$Year>=2000,"year_bucket"] = "b.2000_2004"
+  df[df$Year>=2004,"year_bucket"] = "c.2004_2008"
+  df[df$Year>=2008,"year_bucket"] = "d.2008_2012"
+  df[df$Year>=2012,"year_bucket"] = "e.2012_2016"
+  df
+}
+```
+
+```{r}
+aggregate_on_month_and_transpose = function(df, func){
+  return_df = aggregate(. ~ month, df, func)
+  n = return_df$month
+  # transpose all but the first column (month)
+  return_df <- as.data.frame(t(return_df[,-1]))
+  colnames(return_df) <- n
+  return_df = cbind(Year = rownames(return_df), return_df)
+  rownames(return_df) = 1:nrow(return_df)
+  return_df = return_df[, c(1, 3, 2, 5, 4)]
+  return_df
+}
+```
+
+```{r}
+mean_df = aggregate_on_month_and_transpose(month_df, mean)
+mean_df = year_bucketing_function(mean_df)
+mean_df = mean_df[,-1]
+mean_df = aggregate(.~year_bucket, mean_df, mean)
+mean_df
+```
+
+```{r}
+min_df = aggregate_on_month_and_transpose(month_df, min)
+min_df = year_bucketing_function(min_df)
+min_df = min_df[,-1]
+min_df = aggregate(.~year_bucket, min_df, mean)
+min_df
+```
+
+```{r}
+max_df = aggregate_on_month_and_transpose(month_df, max)
+max_df = year_bucketing_function(max_df)
+max_df = max_df[,-1]
+max_df = aggregate(.~year_bucket, max_df, mean)
+max_df
+```

### 2022-09-14 hw3-1/submission.Rmd
draft submission

---
title: "ISYE 6501"
output:
  html_document:
    df_print: paged
  pdf_document: default
date: "2022-09-12"
---

# Submission HW2 \| Fall 22

-   Ashish Dhiman \| [ashish.dhiman\@gatech.edu](mailto:ashish.dhiman@gatech.edu){.email}

-   Abhinav Arun \| [aarun60\@gatech.edu](mailto:aarun60gatech.edu){.email}

-   Anshit Verma \| [averma373\@gatech.edu](mailto:averma373@gatech.edu){.email}

[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}

```{r}
#lib imports
library(grid)
library(reshape2)
library(ggplot2)
```

### Read Data

```{r}
data = read.table('./temps.txt', sep="\t", dec=".", header = TRUE)
head(data)
```

### Transform Temperature data from Fahrenheit to Celsius

We can afford this transformation here, b'cos it only alters the scale of data, and not location of change.

```{r}
### Transform to celsius
temp_cols = names(data)[!(names(data) %in% c("DAY"))]
#Transform to celsius
data_celsius = data
data_celsius[temp_cols] <- lapply(data_celsius[temp_cols], function(f) (f-32)*5/9)
```

### EDA of Temperature Data

```{r}
summary(data_celsius)
```

### Visualize the temperature series with a boxplot

```{r}
data_celsius$month = unlist(
  lapply(data_celsius$DAY, function(x) substr(x,start = nchar(x)-3+1,stop = nchar(x)+1)))
data_mod <- melt(data_celsius, id.vars='month', 
                  measure.vars=temp_cols)
month_order = c('Jul', 'Aug', 'Sep', 'Oct')
ggplot(data_mod) +
geom_boxplot(aes(x=factor(month, level = month_order), y=value, color=variable))
```

[**#Analysis**]{style="color:red"}

-   As expected, Jul and Aug are hotter for all the years. In Sept the temperature starts cooling off a bit and cools further in October.

-   Also, Sept and October especially October has much more Temperature variance, and this effect is much more accentuated for the 2000s.

-   Because for most years the temperature starts tapering between Sept and Oct, we expect the change detection to lie somewhere in that range.

## Question 6.1

## Question 6.2

### Part 1: Start of Winter through cusum

#### Function for Cusum

```{r}
CUSUM_ad = function(x, year, days = data$DAY ,C = 0.30,T = 4, if_c_relative = FALSE){
  mean_x = mean(x)
  sd_x = sd(x)
  if (if_c_relative) {
    C = C * sd_x
  }
  #apply cusum
  x1 = lapply(x, function(xi)  (mean_x - xi - C))
  st_df = data.frame(st=double())
  st = 0
  for(i in 1:length(x)){
    #print (paste(i,class(st),class(x1[[i]])))
    st = max(0, st + x1[[i]])
    st_df[i,] = c(st)
  }
  #Find index where St > T
  st_df$Day = days
  st_df["anamoly"] = 0
  st_df$anamoly[st_df$st > T] = 1
  st_df$xt = x
  #plot x with T and if decrease see
  par(mfrow=c(2,1))
  plot(st_df$x,type='o')
  title(paste("xt (in Celsius) and St (bottom) for (year,C,T):",year,C,T), line = -1, outer = TRUE)
  plot(st_df$st,type='o')
  pushViewport(viewport())
  grid.lines(x = c(0,1), y = grconvertY(T, "user", "ndc"), gp = gpar(col = "red"))
  change_detected = which.max(st_df$st>T)
  #for zero true, which max returns 1
  if (sum(st_df$st>T) == 0){
    print ("No value greater than T")
    change_detected = NaN
  }
  grid.lines(x = grconvertX(change_detected, "user", "ndc"), y = c(0,1), gp = gpar(col = "red"))
  popViewport()
  #return first point where St touches T
  change_length = length(days)-change_detected
  return (c(as.numeric(year),as.numeric(change_detected),as.numeric(change_length)))
}
```

Test the function for 1996

```{r}
CUSUM_ad(data$X1996,"1996", C= 5, T = 10)
```

#### Understanding the effect of C on St calculation

```{r}
C_effect = lapply(seq(1,20,2),function(x) CUSUM_ad(data$X1996,"1996", C= x, T = 10))
```

[**#Analysis**]{style="color:red"}

C is like the margin for noise built into the model. From the above graphs it is apparent that increasing C, makes the end part (\>90) of St more grainy and abrupt. This is intuitive since St = max(0,St-1 + (mean - x - C), therefore any drop in xt (relative to mean) lower than C would be marginalised. This further warrants that St would be increasing only when we see a sustained drop in xt beyond C.

***Here Winter Start = Days since Jul 1 until our model detects Start of Winter (or Summer end)***

*\*Fall ignored here for brevity*

```{r}
C_effect = as.data.frame(do.call(rbind, C_effect))
colnames(C_effect) = c("year","winter_start","winter_length")
C_effect$C = seq(1,20,2)
plot(x=C_effect$C,y=C_effect$winter_start,type='o')
title("Effect of C on Winter Start")
```

[**#Analysis**]{style="color:red"}

Increasing the effect of C thus makes the model less sensitive, and delays the resolution of change, winter_start increases.

#### Understanding the effect of T on St calculation

```{r}
T_effect = lapply(seq(1,100,20),function(x) CUSUM_ad(data$X1996,"1996", C= 5, T = x))
```

[**#Analysis**]{style="color:red"}

As expected altering T, does not effect St. Simply put it is just the threshold, we have decided to identify the marker of change.

```{r}
T_effect = as.data.frame(do.call(rbind, T_effect))
colnames(T_effect) = c("year","winter_start","winter_length")
T_effect$C = seq(1,20,2)
plot(x=T_effect$C,y=T_effect$winter_start,type='o')
```

[**#Analysis**]{style="color:red"}

Increasing the effect of T too thus makes the model less sensitive, and delays the resolution of change, i.e. *winter_start* increases.

### Basis the above inferences we have picked C= and T= (basis hit and trial)

```{r}
winter_start_days = lapply(temp_cols, function(col) 
  CUSUM_ad(x=data_celsius[,col], year = substr(col,2,5), C = 3.5, T = 35)
  )
```

### Analyse the effect of Model over the years

```{r}
change_df <- as.data.frame(do.call(rbind, winter_start_days))
colnames(change_df) = c("year","winter_start","winter_length")
change_df["year_bucket"] = "a.1996_2000"
change_df[change_df$year>=2000,"year_bucket"] = "b.2000_2004"
change_df[change_df$year>=2004,"year_bucket"] = "c.2004_2008"
change_df[change_df$year>=2008,"year_bucket"] = "d.2008_2012"
change_df[change_df$year>=2012,"year_bucket"] = "e.2012_2016"
change_df
```

```{r}
summary(change_df$winter_start)
```

[**#Analysis**]{style="color:red"}

**We can from p25 and p75 here, that for majority of years, our model detects start of winter from 15th to 23rd October.**

```{r}
change_df[is.na(change_df$winter_start),]
```

For 2014, we are unable to detect the start of winter till Oct 31. It could be that September and October was unusually hot for 2014 or it showed high variability and no sustined drop in the data time period available.

```{r}
CUSUM_ad(x=data_celsius[,"X2014"], year = substr("X2014",2,5), C = 3.5, T = 35)
data_celsius$month = unlist(
  lapply(data_celsius$DAY, function(x) substr(x,start = nchar(x)-3+1,stop = nchar(x)+1)))
data_mod <- melt(data_celsius, id.vars='month', 
                  measure.vars=c("X2012","X2013","X2014","X2015"))
month_order = c('Jul', 'Aug', 'Sep', 'Oct')
ggplot(data_mod) +
geom_boxplot(aes(x=factor(month, level = month_order), y=value, color=variable))
```

[**#Analysis**]{style="color:red"}

As expected October for 2014 is relatively warmer.

#### Year with most delayed winter start

```{r}
change_df[which.max(change_df$winter_start),]
```

#### Year with quickest winter start

```{r}
change_df[which.min(change_df$winter_start),]
```

### Part 2: Has Atlanta become warmer over the years ?

We have the years into a four month window. if suppose Atlanta has become warmer, on average the start of winter should be more delayed in recent years compared to 1990's.

```{r}
summary_df = aggregate(winter_start ~ year_bucket, change_df, mean)
summary_df
```

[**#Analysis**]{style="color:red"}

We do see a slight delay in onset of winter in Atlanta from 1996 to 2015. This trend generally holds except for 2008 to 2011, which seems like a relatively cooler epoch.

```{r}

```

### 2022-09-14 hw3-1/submission.Rmd
draft v2

@@
 
+We can further test this hypothesis by looking at average temperatures of different months in the same windows.
+
 ```{r}
+aggregate_on_month_and_transpose = function(df, func){
+  return_df = aggregate(. ~ month, df, func)
+  n = return_df$month
+  # transpose all but the first column (month)
+  return_df <- as.data.frame(t(return_df[,-1]))
+  colnames(return_df) <- n
+  return_df = cbind(Year = rownames(return_df), return_df)
+  rownames(return_df) = 1:nrow(return_df)
+  return_df = return_df[, c(1, 3, 2, 5, 4)]
+  return_df
+}
 
+month_df = data_celsius[,!(names(data_celsius) %in% c("DAY"))]
+for ( col in 1:ncol(month_df)){
+    colnames(month_df)[col] <-  sub("X", "", colnames(month_df)[col])
+}
+
+year_bucketing_function = function(df){
+  df["year_bucket"] = "a.1996_2000"
+  df[df$Year>=2000,"year_bucket"] = "b.2000_2004"
+  df[df$Year>=2004,"year_bucket"] = "c.2004_2008"
+  df[df$Year>=2008,"year_bucket"] = "d.2008_2012"
+  df[df$Year>=2012,"year_bucket"] = "e.2012_2016"
+  df
+}
 ```
+
+#### Mean of Average temperature (daily high)
+
+```{r}
+mean_df = aggregate_on_month_and_transpose(month_df, mean)
+mean_df = year_bucketing_function(mean_df)
+mean_df = mean_df[,-1]
+mean_df = aggregate(.~year_bucket, mean_df, mean)
+mean_df
+```
+
+#### Mean of Min temperature
+
+```{r}
+min_df = aggregate_on_month_and_transpose(month_df, min)
+min_df = year_bucketing_function(min_df)
+min_df = min_df[,-1]
+min_df = aggregate(.~year_bucket, min_df, mean)
+min_df
+```
+
+#### Mean of Max temperature
+
+```{r}
+max_df = aggregate_on_month_and_transpose(month_df, max)
+max_df = year_bucketing_function(max_df)
+max_df = max_df[,-1]
+max_df = aggregate(.~year_bucket, max_df, mean)
+max_df
+```

### 2022-09-14 hw3-1/submission.Rmd
submission hw3 v1

@@
 
-We can afford this transformation here, b'cos it only alters the scale of data, and not location of change.
+**We can afford this transformation here, b'cos it only alters the scale of data, and not location of change.**
 
@@
 
+### Credit Risk Triggers during the Covid pandemic
+
+The Covid pandemic created a dynamic Credit Risk situation.
+It warranted preactive monitoring of risky segments of the portfolio, to be ahead of the recession curve.
+Hence, during pandemic, we monitored the "Natural Disaster" Trigger data from Bureau, to ascertain, if there was a "change picking up" and we should account for it.
+
+**Natural Disaster Trigger:** Regulations in the US, mandate different lenders to report to Credit Bureaus, accounts who are availing financial relief on account of "Natural Disasters".
+This information is then collated by the Bureau and shared with all the participating lenders.
+
+**CUSUM Application**: The trigger data was shared daily, basis the real-time events.
+This essentially translated to a time series of #Natural Disasters triggers received each day.
+The shape of this curve greatly resembled the Covid cases chart.
+*Cusum could be applied on this time series to ascertain if the triggers generated from covid were different from historical natural disaster triggers like hurricanes etc.*
+
+**C & T choice**:
+
+C & T could be chosen such that the model detects the covid change but neglects the change seen previously with other natural disasters.
+
 ## Question 6.2
@@
   }
-  #Find index where St > T
+  #Ancillary Columns
   st_df$Day = days
-  st_df["anamoly"] = 0
-  st_df$anamoly[st_df$st > T] = 1
   st_df$xt = x
@@
 
-*\*Fall ignored here for brevity*
+*\*Fall and Winter are considered equivalent for this analysis*
 
@@
 colnames(T_effect) = c("year","winter_start","winter_length")
-T_effect$C = seq(1,20,2)
-plot(x=T_effect$C,y=T_effect$winter_start,type='o')
+T_effect$T = seq(1,100,20)
+plot(x=T_effect$T,y=T_effect$winter_start,type='o')
 ```
@@
 
-### Basis the above inferences we have picked C= and T= (basis hit and trial)
+### Basis the above inferences we have picked C=3.5 and T=35 (basis hit and trial)
+
+The value of C here implies that, fluctations within 3.5 Celsius are discarded as noise.
 
@@
 
-**We can from p25 and p75 here, that for majority of years, our model detects start of winter from 15th to 23rd October.**
+**We have p25 = 107 and p75=115 here, i.e. for majority of years, our model detects start of winter between 15th to 23rd October.**
+
+**However we also see 1 NA, which requires more investigation.**
 
@@
 
-For 2014, we are unable to detect the start of winter till Oct 31.
-It could be that September and October was unusually hot for 2014 or it showed high variability and no sustined drop in the data time period available.
+For 2014, we are unable to detect the start of winter till Oct 3, and it might be the case that winter started in Nov that year.
+This might happen because because September and October was unusually hot for 2014 or temperature showed high variability and no sustained drop in the data time period was visible.
+
+If required this behaviour could be easily handled by altering the C and T values.
 
 ```{r}
-CUSUM_ad(x=data_celsius[,"X2014"], year = substr("X2014",2,5), C = 3.5, T = 35)
+#If we want altered model
+CUSUM_ad(x=data_celsius[,"X2014"], year = substr("X2014",2,5), C = 3, T = 35)
 data_celsius$month = unlist(
@@
 
-We have the years into a four month window. if suppose Atlanta has become warmer, on average the start of winter should be more delayed in recent years compared to 1990's.
+We have the bucketed years into a four year window. if suppose Atlanta has become warmer, on average the start of winter should be more delayed in recent years compared to 1990's.
 
@@
 
-We do see a slight delay in onset of winter in Atlanta from 1996 to 2015. This trend generally holds except for 2008 to 2011, which seems like a relatively cooler epoch.
+**We do see a slight delay in onset of winter in Atlanta from 1996 to 2015. This trend generally holds except for 2008 to 2011, which seems like a relatively cooler epoch.**
 
@@
 
+[**#Analysis**]{style="color:red"}
+
+What we see here is that average daily high temperature has mostly remained static over the years.
+Except for the window: 2008 to 2012 where Jul & Aug are warmer but October is cooler.
+This is also the same epoch where the winter start trend breaks.
+This implies that this epoch was very volatile.
+
 #### Mean of Min temperature
@@
 
+[**#Analysis**]{style="color:red"}
+
+When looking at the average of coldest daily high temperature, what we see is that there is little variation for most of the months, except for Sept.
+The coldest day on average was more hotter over the years, especially in the 2000s.
+
 #### Mean of Max temperature
@@
 ```
+
+[**#Analysis**]{style="color:red"}
+
+We see here that on average, August/Sept are seeing more hot days in recent years.

### 2022-09-19 hw4-1/ashish/a.outliers.Rmd
draft

---
title: "a.outliers"
author: "Ashish Dhiman"
date: "2022-09-19"
output: html_document
---

```{r}
library(outliers)
library(ggplot2)
library(qqplotr)
```

```{bash}
pwd
head -4 ../uscrime.txt
```

### EDA of Crime Data

```{r}
crime_df = read.table("../uscrime.txt", sep="\t", header= TRUE)
dim(crime_df)
```

3 states are missing in the data, we don't know which.

#### Data and variable meaning:

Criminologists are interested in the effect of punishment regimes on crime rates. This has been studied using aggregate data on 47 states of the USA for 1960.

| Variable | Description                                                                            |
|-------------------|-----------------------------------------------------|
| M        | percentage of males aged 14--24 in total state population                              |
| So       | indicator variable for a southern state                                                |
| Ed       | mean years of schooling of the population aged 25 years or over                        |
| Po1      | per capita expenditure on police protection in 1960                                    |
| Po2      | per capita expenditure on police protection in 1959                                    |
| LF       | labour force participation rate of civilian urban males in the age-group 14-24         |
| M.F      | number of males per 100 females                                                        |
| Pop      | state population in 1960 in hundred thousands                                          |
| NW       | percentage of nonwhites in the population                                              |
| U1       | unemployment rate of urban males 14--24                                                |
| U2       | unemployment rate of urban males 35--39                                                |
| Wealth   | wealth: median value of transferable assets or family income                           |
| Ineq     | income inequality: percentage of families earning below half the median income         |
| Prob     | probability of imprisonment: ratio of number of commitments to number of offenses      |
| Time     | average time in months served by offenders in state prisons before their first release |
| Crime    | crime rate: number of offenses per 100,000 population in 1960                          |

```{r}
summary(crime_df$Crime)
```

```{r}
aggregate(cbind(Pop) ~ So, crime_df, sum)

#split by Southern states
south_split = aggregate(cbind(Ineq,M,M.F,U1,Time,NW,Ed) ~ So, crime_df, mean)
south_split
```

We see that Southern states in 1960s(on average):

-   had \<50% of non So population

-   were more inequal

-   had greater proportion of NW

-   lower mean years of schooling

```{r}
#Boxplot of crime
boxplot(crime_df$Crime, horizontal = TRUE)
```

```{r}
boxplot(crime_df$Crime, horizontal = TRUE)
```

```{r}
#Plot crime by population, expect this to be straight line
plot(x=crime_df$Pop,y=crime_df$Crime)
```

### Outliers package

To apply Grubbs Test

Grubbs' test ([Grubbs 1969](https://www.itl.nist.gov/div898/handbook/eda/section4/eda43.htm#Grubbs) and [Stefansky 1972](https://www.itl.nist.gov/div898/handbook/eda/section4/eda43.htm#Stefansky)) is used to detect a single [outlier](https://www.itl.nist.gov/div898/handbook/eda/section3/eda35h.htm) in a univariate data set that follows an [approximately normal](https://www.itl.nist.gov/div898/handbook/eda/section3/eda35h.htm#Normality) distribution.

| H~0~: | There are no outliers in the data set        |
|-------|----------------------------------------------|
| H~a~: | There is exactly one outlier in the data set |

The two salient points here are:

1.  Data is Normally Distributed
2.  There is exactly one outlier

Before applying the test let us check for these two assumptions:

#### Assumption 1: Normality

```{r}
hist(crime_df$Crime,breaks=25)
```

```{r}
hist(log(crime_df$Crime),breaks=20)
```

Basis Histogram the data presents deviation from normality as the distribution is not symmetric, especially without the log transformation\

```{r}
ggplot(mapping = aes(sample = crime_df$Crime)) + stat_qq_point(size = 2) + stat_qq_line(color="green")
```

```{r}
ggplot(mapping = aes(sample = log(crime_df$Crime))) + stat_qq_point(size = 2) + stat_qq_line(color="green")
```

Similarly from QQ plot, we see that the points diverge from line especially at the extremes, hinting at non-normality in that region. While log of crime is relatively more normal

**Therefore we can not be 100% sure of the efficiency of Grubbs Test in our case.**

#### Assumption 2: Exactly one outlier

To check this assumption let us check the box plot of crime:

```{r}
#boxplot(crime_df$Crime ~ crime_df$So)
boxplot(crime_df$Crime)
```

```{r}
#boxplot(log(crime_df$Crime))
```

two points seem pretty far away from the whiskers.

Let us further examine these two points:

```{r}
mask = crime_df$Crime > 1800
crime_df[mask,"Crime"]
```

```{r}
crime_df[mask,]
```

We see that among these two points, there is disproportionate difference of population, $157 * 10^5$ vs $3 * 10^5$.

While the Crime variable is already normalised, number of offenses per 100,000 population in 1960. such high crime rate for such small population seems counter intuitive, especially so when all other relevant variables like Po1, Po2, Prob are fairly similar. Let us examine this in detail by plotting

```{r}
plot(x=crime_df$Pop,y=crime_df$Crime)
cor(crime_df$Pop,crime_df$Crime)
```

```{r}
cor(crime_df$Prob,crime_df$Crime)
```

### Grubbs Test

```{r}
grubbs.test(crime_df$Crime, type = 10, opposite = FALSE, two.sided = FALSE)
```

As per Grubbs test, we get p-value of \~8%. $\implies$ there is 8% chance that the given point is the effect of randomness, assuming Null Hypothesis is True. Generally we want p-value to be as low as possible to reject the Null hypothesis.In this case 8% value is slightly larger than typical thresholds of 5%, and warrants detailed investigation before classifying the point as outlier.

Let us test what happens to the 2nd highest point, if the highest is removed

```{r}
mask1 = crime_df$Crime<1990
grubbs.test(crime_df[mask1,]$Crime, type = 10, opposite = FALSE, two.sided = FALSE)
```

After removing 1993 point we get even lower p-value this time

```{r}
mask2 = crime_df$Crime<1960
grubbs.test(crime_df[mask2,]$Crime, type = 10, opposite = FALSE, two.sided = FALSE)
```

```{r}

```

### 2022-09-20 hw4-1/ashish/a.outliers.Rmd
outlier ques final

@@
 | Variable | Description                                                                            |
-|-------------------|-----------------------------------------------------|
+|----------|----------------------------------------------------------------------------------------|
 | M        | percentage of males aged 14--24 in total state population                              |
@@
 
+### Difference b/w Southern and Northern states
+
 ```{r}
-aggregate(cbind(Pop) ~ So, crime_df, sum)
+aggregate(Pop ~ So, crime_df, sum)
+```
 
-#split by Southern states
-south_split = aggregate(cbind(Ineq,M,M.F,U1,Time,NW,Ed) ~ So, crime_df, mean)
+```{r}
+south_split = aggregate(cbind(Crime,Ineq,M,M.F,U1,Time,NW,Ed) ~ So, crime_df, mean)
 south_split
@@
 ```{r}
-#Boxplot of crime
-boxplot(crime_df$Crime, horizontal = TRUE)
+#Plot crime by population
+plot(x=crime_df$Pop,y=crime_df$Crime)
 ```
 
-```{r}
-boxplot(crime_df$Crime, horizontal = TRUE)
-```
-
-```{r}
-#Plot crime by population, expect this to be straight line
-plot(x=crime_df$Pop,y=crime_df$Crime)
-```
+The two extreme large points seem like potentail outliers
 
@@
 
-The two salient points here are:
+The two salient assumptions here are:
 
@@
 
-Basis Histogram the data presents deviation from normality as the distribution is not symmetric, especially without the log transformation\
+Basis Histogram the data presents deviation from normality as the distribution is not symmetric, especially without the log transformation. We can further test this with QQ plot
 
@@
 
-Similarly from QQ plot, we see that the points diverge from line especially at the extremes, hinting at non-normality in that region. While log of crime is relatively more normal
+Similarly from QQ plot, we see that the points diverge from line especially at the extremes, hinting at non-normality in that region.
+While log of crime is relatively more normal, the log transformation alters the scale of data, and we might want to detect outliers before applying it.
 
@@
 ```{r}
-#boxplot(crime_df$Crime ~ crime_df$So)
 boxplot(crime_df$Crime)
@@
 
-```{r}
-#boxplot(log(crime_df$Crime))
-```
-
-two points seem pretty far away from the whiskers.
+As seen in scatter plot, there are two points seem pretty far away from the whiskers.
 
@@
 such high crime rate for such small population seems counter intuitive, especially so when all other relevant variables like Po1, Po2, Prob are fairly similar.
-Let us examine this in detail by plotting
+Let us examine this in detail by checking corelation:
 
 ```{r}
-plot(x=crime_df$Pop,y=crime_df$Crime)
-cor(crime_df$Pop,crime_df$Crime)
+paste(cor(crime_df$Pop,crime_df$Crime),cor(crime_df$Prob,crime_df$Crime))
 ```
 
-```{r}
-cor(crime_df$Prob,crime_df$Crime)
-```
+Basis population figures in 1960's, there are two candidate states for 1993 point: Wyoming & Nevada.
+
+And comparing data from [Disatster Center](https://www.disastercenter.com/crime/wycrime.htm), Wyoming does have roughly twice the national average crime rate, which is similar to our data as well (1993/ 905.1).
+Maybe there was some particular event in 1960 in Wyoming, which is causing such behavior.
+To understand this better, we need some expertise of Crime rate Trends in Wyoming.
+
+<https://en.wikipedia.org/wiki/1960_United_States_census#:~:text=The%20United%20States%20census%20of,enumerated%20during%20the%201950%20census.>
 
@@
 
-After removing 1993 point we get even lower p-value this time
+After removing 1993 point we get even lower p-value this time, hinting that 1969 in absence of 1993 is a much stronger candidate for being a outlier.
 
-```{r}
-mask2 = crime_df$Crime<1960
-grubbs.test(crime_df[mask2,]$Crime, type = 10, opposite = FALSE, two.sided = FALSE)
-```
-
-```{r}
-
-```
+Thus in conclusion, we have two candidate outliers in the data, 1993 and 1969 point.
+1993 point is the highest crime rate and presents weak evidence to discard it as ouitlier, however, to be definitive, we need to study the Wyoming angle in detail, especially in relevance to exact modelling needs at hand.

### 2022-09-20 hw4-1/abhinav/mlr.Rmd
MLR code push

---
title: "mlr_assg4"
output: html_document
date: "2022-09-20"
---

```{r setup, include=FALSE}
library(cowplot)
library(ggplot2)
library(reshape2)
library(grid)
```

```{r crimes}
crime_data<-read.table(file = "./uscrime.txt", sep = "\t",header=TRUE)
head(crime_data)
dim(crime_data)
summary(crime_data)
```

#### Data Distribution and Pairwise Corelation
### Data Distribution
```{r}
my_plots <- lapply(names(crime_data), function(var_x){
  p <- 
    ggplot(crime_data) +
    aes_string(var_x)

  if(var_x %in% list("So")) {
    p <- p + geom_bar()

  } else {
    p <- p + geom_density()
  }

})

plot_grid(plotlist = my_plots)

### Corealtion
cormat <- round(cor(crime_data),2)
cormat[upper.tri(cormat)] <- NA
melted_cormat <- melt(cormat)
# plotting the correlation heatmap
library(ggplot2)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2,
                                   fill=value)) +
geom_tile() +
geom_text(aes(Var1, Var2, label = value),
          color = "white", size = 3)
```
## Observations: 
Wealth and Income Inequality are 2 highly correlated variables (-ve correlation) ,  hence we'll pick only 1 out of 2 
Po1 and Po2 are 2 correlated variables , hence we'll pick only 1 out of 2 
Same goes for U1 and U2, , hence we'll pick only 1 out of 2

The dependent variable Crime Rate has high correlation with the following variables :
Po1 and Po2 -> surprising to see +ve correlation
Prob -> makes sense to have -ve correlation

```{r}
crime_data$So<-as.factor(crime_data$So)
# Linear Regression model with all features
lmodel_1<-lm(Crime~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data)
summary(lmodel_1)

# Linear Regression Model with Limited features (removing highly correlated features)
lmodel_2<-lm(Crime~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data)
summary(lmodel_2)

```

```{r}
par(mfrow = c(2, 2))
plot(lmodel_2)
```

```{r}
crime_data_wo<-crime_data[-c(11,19,29),]
dim(crime_data_wo)
# Linear Regression model with all features
lmodel_1<-lm(Crime~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data_wo)
summary(lmodel_1)
# Linear Regression Model with Limited features (removing highly correlated features)
lmodel_4<-lm(Crime~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data_wo)
summary(lmodel_4)

```

```{r}
# Plot after removing the outliers/leverage points
par(mfrow = c(2, 2))
plot(lmodel_4)
```

```{r}
crime_data_wo<-crime_data[-c(11,19,29,46),]
dim(crime_data_wo)
# Linear Regression model with all features
lmodel_5<-lm(Crime~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data_wo)
summary(lmodel_5)
# Linear Regression Model with Limited features (removing highly correlated features)
lmodel_6<-lm(Crime~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data_wo)
summary(lmodel_6)

```

### 2022-09-20 hw4-1/abhinav/mlr.Rmd
MLR code update

@@
 ## Observations: 
-Wealth and Income Inequality are 2 highly correlated variables (-ve correlation) ,  hence we'll pick only 1 out of 2 
-Po1 and Po2 are 2 correlated variables , hence we'll pick only 1 out of 2 
-Same goes for U1 and U2, , hence we'll pick only 1 out of 2
+Wealth and Income Inequality are 2 highly correlated variables (-ve correlation) ,hence we can 1 variable out of both
+Po1 and Po2 are 2 correlated variables , hence we can pick only 1 out of 2 
+Same goes for U1 and U2, , hence we can pick only 1 out of 2
 
@@
 ```{r}
+options(scipen=999)
 crime_data$So<-as.factor(crime_data$So)
@@
 summary(lmodel_1)
+# plotting the residuals to check for assumptions in the plain vanilla model with all features
+par(mfrow = c(2, 2))
+plot(lmodel_1)
+```
 
+```{r}
+# removing all the potential outliers/influential points
+crime_data_wo<-crime_data[-c(11,46,29),]
+dim(crime_data_wo)
+# Linear Regression model with all features on the reduced dataset
+lmodel_1<-lm(Crime~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data_wo)
+summary(lmodel_1)
+```
+
+```{r}
 # Linear Regression Model with Limited features (removing highly correlated features)
@@
 
-```
-
-```{r}
 par(mfrow = c(2, 2))

### 2022-09-20 hw4-1/abhinav/mlr.Rmd
MLR code updated

@@
 
-### Corealtion
+### Correlation
 cormat <- round(cor(crime_data),2)
@@
 ## Observations: 
-Wealth and Income Inequality are 2 highly correlated variables (-ve correlation) ,hence we can 1 variable out of both
+Wealth and Income Inequality are 2 highly correlated variables (-ve correlation) ,hence we can pick only 1 variable out of the 2
 Po1 and Po2 are 2 correlated variables , hence we can pick only 1 out of 2 
@@
 The dependent variable Crime Rate has high correlation with the following variables :
-Po1 and Po2 -> surprising to see +ve correlation
-Prob -> makes sense to have -ve correlation
+Po1 and Po2 -> we see a surprising +ve correlation with crime rate , As regression cannot be used for defining a causal relation , we can infer the other way round that since the crime rate is higher , that could be the reason of a higher per capita expenditure on police protection.
+
+Prob -> makes sense to have -ve correlation in this case . As the crime rate is bound to decrease if probability of imprisonment increases.
 
@@
 ```{r}
-# removing all the potential outliers/influential points
+# removing all the potential outliers/influential points based on the diagnostic plots of linear regression model
 crime_data_wo<-crime_data[-c(11,46,29),]
@@
 summary(lmodel_1)
+```
+
+```{r}
+new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
+crime_rate_pred<-predict(lmodel_1,newdata = new_datapoint)
+crime_rate_pred
 ```

### 2022-09-20 hw4-1/abhinav/mlr.Rmd
WIP-abhinav

@@
 plot(lmodel_1)
+
+new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
+crime_rate_pred<-predict(lmodel_1,newdata = new_datapoint)
+crime_rate_pred
 ```
@@
 ```{r}
-# removing all the potential outliers/influential points based on the diagnostic plots of linear regression model
-crime_data_wo<-crime_data[-c(11,46,29),]
+options(scipen=999)
+crime_data$So<-as.factor(crime_data$So)
+# Linear Regression model with all features
+lmodel_2<-lm(log(Crime)~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data)
+summary(lmodel_2)
+# plotting the residuals to check for assumptions in the plain vanilla model with all features
+par(mfrow = c(2, 2))
+plot(lmodel_2)
+
+new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
+crime_rate_pred<-predict(lmodel_2,newdata = new_datapoint)
+crime_rate_pred<- exp(crime_rate_pred)
+crime_rate_pred
+```
+
+```{r}
+# removing the potential outliers/influential points based on the diagnostic plots of linear regression model
+crime_data_wo<-crime_data[-c(11, 19, 46),]
 dim(crime_data_wo)
 # Linear Regression model with all features on the reduced dataset
-lmodel_1<-lm(Crime~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data_wo)
-summary(lmodel_1)
+lmodel_2<-lm(log(Crime)~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data_wo)
+summary(lmodel_2)
 ```
@@
 new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
-crime_rate_pred<-predict(lmodel_1,newdata = new_datapoint)
+crime_rate_pred<-predict(lmodel_4,newdata = new_datapoint)
 crime_rate_pred
+```
+```{r}
+crime_data[c(4,26),]
 ```
@@
 # Linear Regression Model with Limited features (removing highly correlated features)
-lmodel_4<-lm(Crime~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data_wo)
+lmodel_4<-lm(Crime~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data)
 summary(lmodel_4)
@@
 summary(lmodel_6)
-
 ```

### 2022-09-20 hw4-1/abhinav/mlr.Rmd
Code for Review

@@
 ```
+# Interpreting the output of the linear regression model :
 

### 2022-09-21 hw4-1/abhinav/mlr.Rmd
Final Code push

--- 2022-09-20 hw4-1/abhinav/mlr.Rmd
@@
 plot(lmodel_1)
+
+new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
+crime_rate_pred<-predict(lmodel_1,newdata = new_datapoint)
+crime_rate_pred<-round(crime_rate_pred,2)
+print(paste("Predicted Value:", crime_rate_pred))
+```
+# Interpreting the output of the linear regression model : 
+Based on the model output , it appears that the features Ed, Ineq, Prob,M , Po1 are significant variables based on the p-values . 
+* The lower the p-value , the more significant the variable  because the p-value signifies the probability of the estimated coefficients $\hat{\beta_{i}}'s$ coming from the underlying assumed distribution of null hypothesis (Null Hypothesis states that $\beta_{i}'s=0$).
+Thus , lower the p-value , higher the probability that the coefficient of the variable is not equal to 0 and hence that variable is a good predictor.
+* For this model , the $R^{2}$ value is 0.8 which means that 80% of the variance in the dependent variable is explained by the set of these independent variables .
+* The predicted crime_rate for the given data point is `r crime_rate_pred`.
+* From the 4 diagnostic plots associated with the linear regression model , we can see that the assumptions are mostly in place , error terms are normally distributed and residuals are homogeneous around the fitted line.
+However , we can see slight bends in the left and right end in the plot of residuals vs fitted values.
+Thus , a possible remedy can be thought of as modelling the log of dependent variable based on independent variable.
+(because taking a log transformation reduces the range of possible variables.)
+
+# Points for further exploration  : 
+* Taking a log transformation of the dependent variable and then create a regression model.
+* Create a model with lower number of features (remove highly correlated features).
+* Look for possible outliers/leverage points (based on the residuals vs leverage plot - cook's distance) and remove them to see changes in the model output.
+
+```{r}
+# Taking log transformation of the dependent variable with all features
+options(scipen=999)
+crime_data$So<-as.factor(crime_data$So)
+# Linear Regression model with all features
+lmodel_2<-lm(log(Crime)~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data)
+summary(lmodel_2)
+# plotting the residuals to check for assumptions in the plain vanilla model with all features
+par(mfrow = c(2, 2))
+plot(lmodel_2)
+
+new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
+crime_rate_pred<-predict(lmodel_2,newdata = new_datapoint)
+crime_rate_pred<- exp(crime_rate_pred)
+print(paste("Predicted Value:", round(crime_rate_pred,2)))
+```
+Although the $R^{2}$ takes a slight dip , but it appears that the model assumptions are better satisfied here . 
+Points 11,19,46 seems to have around 1.5*standard deviation , thus these could be tested as potential outliers.
+
+```{r}
+# Taking log transformation of the dependent variable with all features
+options(scipen=999)
+crime_data$So<-as.factor(crime_data$So)
+# Linear Regression model with reduced # of features
+lmodel_3<-lm(log(Crime)~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data)
+summary(lmodel_3)
+# plotting the residuals to check for assumptions in the plain vanilla model with all features
+par(mfrow = c(2, 2))
+plot(lmodel_3)
+
+new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
+crime_rate_pred<-predict(lmodel_3,newdata = new_datapoint)
+crime_rate_pred<- exp(crime_rate_pred)
+print(paste("Predicted Value:", round(crime_rate_pred,2)))
 ```
@@
 ```{r}
-# removing all the potential outliers/influential points based on the diagnostic plots of linear regression model
-crime_data_wo<-crime_data[-c(11,46,29),]
-dim(crime_data_wo)
-# Linear Regression model with all features on the reduced dataset
-lmodel_1<-lm(Crime~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data_wo)
-summary(lmodel_1)
-```
+# removing the potential outliers/influential points based on the diagnostic plots of linear regression model
+crime_data_wo<-crime_data[-c(11, 19, 46),]
 
-```{r}
+# Taking log transformation of the dependent variable with all features
+options(scipen=999)
+crime_data$So<-as.factor(crime_data$So)
+# Linear Regression model with reduced # of features and removed potential influential points.
+lmodel_4<-lm(log(Crime)~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data)
+summary(lmodel_4)
+# plotting the residuals to check for assumptions in the plain vanilla model with all features
+par(mfrow = c(2, 2))
+plot(lmodel_4)
+
 new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
-crime_rate_pred<-predict(lmodel_1,newdata = new_datapoint)
-crime_rate_pred
-```
-
-```{r}
-# Linear Regression Model with Limited features (removing highly correlated features)
-lmodel_2<-lm(Crime~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data)
-summary(lmodel_2)
-
-par(mfrow = c(2, 2))
-plot(lmodel_2)
-```
-
-```{r}
-crime_data_wo<-crime_data[-c(11,19,29),]
-dim(crime_data_wo)
-# Linear Regression model with all features
-lmodel_1<-lm(Crime~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data_wo)
-summary(lmodel_1)
-# Linear Regression Model with Limited features (removing highly correlated features)
-lmodel_4<-lm(Crime~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data_wo)
-summary(lmodel_4)
+crime_rate_pred<-predict(lmodel_4,newdata = new_datapoint)
+crime_rate_pred<- exp(crime_rate_pred)
+print(paste("Predicted Value:", round(crime_rate_pred,2)))
 
 ```
-
-```{r}
-# Plot after removing the outliers/leverage points
-par(mfrow = c(2, 2))
-plot(lmodel_4)
-```
-
-```{r}
-crime_data_wo<-crime_data[-c(11,19,29,46),]
-dim(crime_data_wo)
-# Linear Regression model with all features
-lmodel_5<-lm(Crime~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data_wo)
-summary(lmodel_5)
-# Linear Regression Model with Limited features (removing highly correlated features)
-lmodel_6<-lm(Crime~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data_wo)
-summary(lmodel_6)
-
-```

### 2022-09-21 hw4-1/abhinav/mlr.Rmd
Final Code push-1

@@
 # Interpreting the output of the linear regression model : 
-Based on the model output , it appears that the features Ed, Ineq, Prob,M , Po1 are significant variables based on the p-values . 
+Based on the model output , it appears that the features Ed, Ineq, Prob,M , Po1 are significant variables based on the p-values . \
 * The lower the p-value , the more significant the variable  because the p-value signifies the probability of the estimated coefficients $\hat{\beta_{i}}'s$ coming from the underlying assumed distribution of null hypothesis (Null Hypothesis states that $\beta_{i}'s=0$).
-Thus , lower the p-value , higher the probability that the coefficient of the variable is not equal to 0 and hence that variable is a good predictor.
-* For this model , the $R^{2}$ value is 0.8 which means that 80% of the variance in the dependent variable is explained by the set of these independent variables .
-* The predicted crime_rate for the given data point is `r crime_rate_pred`.
+Thus , lower the p-value , higher the probability that the coefficient of the variable is not equal to 0 and hence that variable is a good predictor.\
+
+* For this model , the $R^{2}$ value is 0.8 which means that 80% of the variance in the dependent variable is explained by the set of these independent variables .\
+
+* The predicted crime_rate for the given data point is `r crime_rate_pred`.\
+
 * From the 4 diagnostic plots associated with the linear regression model , we can see that the assumptions are mostly in place , error terms are normally distributed and residuals are homogeneous around the fitted line.
@@
 ```
-Although the $R^{2}$ takes a slight dip , but it appears that the model assumptions are better satisfied here . 
+Although the $R^{2}$ takes a slight dip , but it appears that the model assumptions are better satisfied here .\ 
 Points 11,19,46 seems to have around 1.5*standard deviation , thus these could be tested as potential outliers.

### 2022-09-21 hw4-1/abhinav/mlr.Rmd
Final Code push-2

@@
 library(grid)
+library(MASS)
 ```
@@
 
-* From the 4 diagnostic plots associated with the linear regression model , we can see that the assumptions are mostly in place , error terms are normally distributed and residuals are homogeneous around the fitted line.
+* From the 4 diagnostic plots associated with the linear regression model , we can see that the assumptions are mostly in place although there are some slight deviations, error terms are normally distributed (deviation for a couple of points) and residuals are mostly homogeneous around the fitted line.
 However , we can see slight bends in the left and right end in the plot of residuals vs fitted values.
@@
 * Look for possible outliers/leverage points (based on the residuals vs leverage plot - cook's distance) and remove them to see changes in the model output. 
+* We can use Box-Cox transformation to identify the best possible transformation of y to regress on . This is usually done to handle the points deviating a bit from the normal Q-Q plot.
+
+```{r}
+# Box Cox Transformation (we can try this out to find the best possible function of y to regress on , the value of lambda is determined through MLE)
+bc<-boxcox(lmodel_1, lambda =seq(-3,3))
+#extract the best lambda 
+best.lam<-bc$x[which(bc$y==max(bc$y))]
+
+```
+# Observations: 
+Since , the above graph gives us a value of $\lambda \approx 0$ , therefore taking a log transformation of the dependent variable y makes more sense.
 
@@
 Although the $R^{2}$ takes a slight dip , but it appears that the model assumptions are better satisfied here .\ 
-Points 11,19,46 seems to have around 1.5*standard deviation , thus these could be tested as potential outliers.
+Points 11,19,46 seems to have around 1.5*standard deviation , thus these could be tested as potential outliers/influential points.
 

### 2022-09-21 hw4-1/abhinav/mlr.Rmd
Final Code push-final

@@
 # Observations: 
-Since , the above graph gives us a value of $\lambda \approx 0$ , therefore taking a log transformation of the dependent variable y makes more sense.
+Since , the above graph gives us the best value of $\lambda$ to be around 0 only , therefore taking a log transformation of the dependent variable y makes more sense.
 

### 2022-09-21 hw4-1/Submission.Rmd
Final Submission File Created

---
title: "Submission HW2"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

# ISYE 6501 \| Fall 22

-   Ashish Dhiman \| [ashish.dhiman\@gatech.edu](mailto:ashish.dhiman@gatech.edu){.email}

-   Abhinav Arun \| [aarun60\@gatech.edu](mailto:aarun60gatech.edu){.email}

-   Anshit Verma \| [averma373\@gatech.edu](mailto:averma373@gatech.edu){.email}

[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}

## Question 5.1

```{r setup, include=FALSE}
library(outliers)
library(ggplot2)
library(qqplotr)
library(cowplot)
library(reshape2)
library(grid)
```

```{bash}
pwd
head -4 ./uscrime.txt
```

### EDA of Crime Data

```{r}
crime_df = read.table("./uscrime.txt", sep="\t", header= TRUE)
dim(crime_df)
```

3 states are missing in the data, we don't know which.

#### Data and variable meaning:

Criminologists are interested in the effect of punishment regimes on crime rates. This has been studied using aggregate data on 47 states of the USA for 1960.

| Variable | Description                                                                            |
|-------------------|-----------------------------------------------------|
| M        | percentage of males aged 14--24 in total state population                              |
| So       | indicator variable for a southern state                                                |
| Ed       | mean years of schooling of the population aged 25 years or over                        |
| Po1      | per capita expenditure on police protection in 1960                                    |
| Po2      | per capita expenditure on police protection in 1959                                    |
| LF       | labour force participation rate of civilian urban males in the age-group 14-24         |
| M.F      | number of males per 100 females                                                        |
| Pop      | state population in 1960 in hundred thousands                                          |
| NW       | percentage of nonwhites in the population                                              |
| U1       | unemployment rate of urban males 14--24                                                |
| U2       | unemployment rate of urban males 35--39                                                |
| Wealth   | wealth: median value of transferable assets or family income                           |
| Ineq     | income inequality: percentage of families earning below half the median income         |
| Prob     | probability of imprisonment: ratio of number of commitments to number of offenses      |
| Time     | average time in months served by offenders in state prisons before their first release |
| Crime    | crime rate: number of offenses per 100,000 population in 1960                          |

```{r}
summary(crime_df$Crime)
```

### Difference b/w Southern and Northern states

```{r}
aggregate(Pop ~ So, crime_df, sum)
```

```{r}
south_split = aggregate(cbind(Crime,Ineq,M,M.F,U1,Time,NW,Ed) ~ So, crime_df, mean)
south_split
```

We see that Southern states in 1960s(on average):

-   had \<50% of non So population

-   were more inequal

-   had greater proportion of NW

-   lower mean years of schooling

```{r}
#Plot crime by population
plot(x=crime_df$Pop,y=crime_df$Crime)
```

The two extreme large points seem like potentail outliers

### Outliers package

Grubbs' test ([Grubbs 1969](https://www.itl.nist.gov/div898/handbook/eda/section4/eda43.htm#Grubbs) and [Stefansky 1972](https://www.itl.nist.gov/div898/handbook/eda/section4/eda43.htm#Stefansky)) is used to detect a single [outlier](https://www.itl.nist.gov/div898/handbook/eda/section3/eda35h.htm) in a univariate data set that follows an [approximately normal](https://www.itl.nist.gov/div898/handbook/eda/section3/eda35h.htm#Normality) distribution.

| H~0~: | There are no outliers in the data set        |
|-------|----------------------------------------------|
| H~a~: | There is exactly one outlier in the data set |

The two salient assumptions here are:

1.  Data is Normally Distributed
2.  There is exactly one outlier

Before applying the test let us check for these two assumptions:

#### Assumption 1: Normality

```{r}
hist(crime_df$Crime,breaks=25)
```

```{r}
hist(log(crime_df$Crime),breaks=20)
```

Basis Histogram the data presents deviation from normality as the distribution is not symmetric, especially without the log transformation. We can further test this with QQ plot

```{r}
ggplot(mapping = aes(sample = crime_df$Crime)) + stat_qq_point(size = 2) + stat_qq_line(color="green")
```

```{r}
ggplot(mapping = aes(sample = log(crime_df$Crime))) + stat_qq_point(size = 2) + stat_qq_line(color="green")
```

Similarly from QQ plot, we see that the points diverge from line especially at the extremes, hinting at non-normality in that region. While log of crime is relatively more normal, the log transformation alters the scale of data, and we might want to detect outliers before applying it.

**Since assumption 1 is not fully satisfied, we can not be 100% sure of the efficiency of Grubbs Test in our case.**

#### Assumption 2: Exactly one outlier

To check this assumption let us check the box plot of crime:

```{r}
boxplot(crime_df$Crime)
```

As seen in scatter plot, there are two points that seem pretty far away from the whiskers.

Let us further examine these two points:

```{r}
mask = crime_df$Crime > 1800
crime_df[mask,"Crime"]
```

```{r}
crime_df[mask,]
```

We see that among these two points, there is disproportionate difference of population, $157 * 10^5$ vs $3 * 10^5$.

While the Crime variable is already normalised, number of offenses per 100,000 population in 1960. such high crime rate for such small population seems counter intuitive, especially so when all other relevant variables like Po1, Po2, Prob are fairly similar. Let us examine this in detail by checking corelation:

```{r}
paste(cor(crime_df$Pop,crime_df$Crime),cor(crime_df$Prob,crime_df$Crime))
```

Basis population figures in 1960's, there are two candidate states for 1993 point: Wyoming & Nevada.

And comparing data from [Disatster Center](https://www.disastercenter.com/crime/wycrime.htm), Wyoming does have roughly twice the national average crime rate, which is similar to our data as well (1993/ 905.1). Maybe there was some particular event in 1960 in Wyoming, which is causing such behavior. To understand this better, we need some expertise of Crime rate Trends in Wyoming.

**Hence this point maybe outlier but there is high likelihood that the underlying data is true, and the decision of dropping it would depend on our exact modelling needs.**

<https://en.wikipedia.org/wiki/1960_United_States_census#:~:text=The%20United%20States%20census%20of,enumerated%20during%20the%201950%20census.>

**Since assumption 2 is again not fully satisfied, we can not be 100% sure of the efficiency of Grubbs Test in our case.**

### Apply Grubbs Test

```{r}
grubbs.test(crime_df$Crime, type = 10, opposite = FALSE, two.sided = FALSE)
```

As per Grubbs test, we get p-value of \~8%. $\implies$ there is 8% chance that the given point is the effect of randomness, assuming Null Hypothesis is True. Generally we want p-value to be as low as possible to reject the Null hypothesis.In this case 8% value is slightly larger than typical thresholds of 5%, and warrants detailed investigation before classifying the point as outlier, which is in sync with analysis above.

Let us also test what happens to the 2nd highest point, if the 1993 pt is removed

```{r}
mask1 = crime_df$Crime<1990
grubbs.test(crime_df[mask1,]$Crime, type = 10, opposite = FALSE, two.sided = FALSE)
```

After removing 1993 point we get even lower p-value this time, hinting that 1969 in absence of 1993 is a much stronger candidate for being a outlier.

Thus in conclusion, we have two candidate outliers in the data, 1993 and 1969 point. 1993 point is the highest crime rate and presents weak evidence to discard it as outlier, however, to be definitive, we need to study the Wyoming angle in detail, especially in relevance to exact modelling needs at hand.

## Question 8.1

### Example: Real Estate Price

Linear Regression models find use case in the real estate industry to predict selling price for a particular property depending on the prediction variables.

**Objective Function (Y)**: To predict the selling price of a property based in the predicting variables for that property

**Predicting Variables (X_i)**: Some of the key predicting variables are given below -

-   ***Location***: The location of the property is an important predictor for predicting the selling price. As properties in a good neighborhood will fetch higher prices than those in a relatively worse neighborhoods.

-   ***Lot Size***: Lot size is defined as the area of the property. A large lot size is worth more than a smaller lot.

-   ***Age of the House (Condition)***: Age of the house is the number of years since the construction of the house. It can also be considered as the condition metric of the house. A house with good condition grade can worth more than one with lower condition grade as it might incur lower maintenance cost.

-   ***Garage Capacity***: Capacity of garage is defines as the number of cars a garage can hold. The ability of the garage to hold more cars is an important factor which appreciates the value of a property

## Question 8.2

```{r crimes}
crime_data<-read.table(file = "./uscrime.txt", sep = "\t",header=TRUE)
head(crime_data)
dim(crime_data)
summary(crime_data)
```

### Data Distribution and Pairwise Corelation
```{r}
my_plots <- lapply(names(crime_data), function(var_x){
  p <- 
    ggplot(crime_data) +
    aes_string(var_x)

  if(var_x %in% list("So")) {
    p <- p + geom_bar()

  } else {
    p <- p + geom_density()
  }

})

plot_grid(plotlist = my_plots)

### Correlation
cormat <- round(cor(crime_data),2)
cormat[upper.tri(cormat)] <- NA
melted_cormat <- melt(cormat)
# plotting the correlation heatmap

ggplot(data = melted_cormat, aes(x=Var1, y=Var2,
                                   fill=value)) +
geom_tile() +
geom_text(aes(Var1, Var2, label = value),
          color = "white", size = 3)
```

[**#Analysis**]{style="color:red"}

Wealth and Income Inequality are 2 highly correlated variables (-ve correlation), hence we can pick only 1 variable out of the 2.
Po1 and Po2 are 2 highly correlated variables (+ve correlation), hence we can pick only 1 out of 2. 
Same goes for U1 and U2, hence we can pick only 1 out of 2

The dependent variable Crime Rate has high correlation with the following variables :
Po1 and Po2 -> we see a surprising +ve correlation with crime rate. As regression cannot be used for defining a causal relation , we can infer the other way round that since the crime rate is higher, that could be the reason of a higher per capita expenditure on police protection.

Prob -> makes sense to have -ve correlation in this case. As the crime rate is bound to decrease if probability of imprisonment increases.

###Linearity in data between independent and dependent variables

```{r}
plot(crime_data$Po1 ~ crime_data$Crime)
```

[**#Analysis**]{style="color:red"}

From the aove plot we can observe that there seems to be a linear relationship between the dependent variable and one of the highly correlated feature. Thus our assumption of linearity holds true and we can implement linear regression model.

###Initial Model

```{r}
options(scipen=999)
crime_data$So<-as.factor(crime_data$So)
# Linear Regression model with all features
lmodel_1<-lm(Crime~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data)
summary(lmodel_1)
# plotting the residuals to check for assumptions in the plain vanilla model with all features
par(mfrow = c(2, 2))
plot(lmodel_1)

new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
crime_rate_pred<-predict(lmodel_1,newdata = new_datapoint)
crime_rate_pred<-round(crime_rate_pred,2)
print(paste("Predicted Value:", crime_rate_pred))
```

[**#Analysis**]{style="color:red"}

**Interpreting the output of the linear regression model** : 
Based on the model output , it appears that the features Ed, Ineq, Prob, M, Po1 are significant variables based on the p-values . \
* The lower the p-value , the more significant the variable  because the p-value signifies the probability of the estimated coefficients $\hat{\beta_{i}}'s$ coming from the underlying assumed distribution of null hypothesis (Null Hypothesis states that $\beta_{i}'s=0$). Thus , lower the p-value , higher the probability that the coefficient of the variable is not equal to 0 and hence that variable is a good predictor.\

* For this model, the $R^{2}$ value is 0.8 which means that 80% of the variance in the dependent variable is explained by the set of these independent variables .\

* The predicted crime_rate for the given data point is `r crime_rate_pred`. Since this prediction lies outside the range of or dependent variable, we can further tune the model as the features of the predicting data point lie withing the domain of independent variables. \

* From the 4 diagnostic plots associated with the linear regression model, we can see that the assumptions are mostly in place, error terms are normally distributed i.e. constant variance assumption and residuals are homogeneous around the fitted line. However , we can see slight bends in the left and right end in the plot of residuals vs fitted values. Thus, a possible remedy can be thought of as transforming the dependent variable (taking a log transformation reduces the range of possible variables).

* Further from the QQ Plot it can be inferred that the data is normally distributed. Thus our normality assumption also holds true.

**Points for further exploration** : 
* Taking a log transformation of the dependent variable and then create a regression model.
* Create a model with lower number of features (remove highly correlated features).
* Look for possible outliers/leverage points (based on the residuals vs leverage plot - cook's distance) and remove them to see changes in the model output.

###Taking log transformation of the dependent variable with all features

```{r}
# Taking log transformation of the dependent variable with all features
options(scipen=999)
crime_data$So<-as.factor(crime_data$So)
# Linear Regression model with all features
lmodel_2<-lm(log(Crime)~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data)
summary(lmodel_2)
# plotting the residuals to check for assumptions in the plain vanilla model with all features
par(mfrow = c(2, 2))
plot(lmodel_2)

crime_rate_pred<-predict(lmodel_2,newdata = new_datapoint)
crime_rate_pred<- exp(crime_rate_pred)
print(paste("Predicted Value:", round(crime_rate_pred,2)))
```

[**#Analysis**]{style="color:red"}

Although the $R^{2}$ takes a slight dip, but it appears that the model assumptions are better satisfied here .\ 
Points 11,19,46 seems to have around 1.5*standard deviation, thus these could be tested as potential outliers. \

Again, the predicted crime_rate for the given data point is `r crime_rate_pred`. The value has gotten within the range of dependent variable. But we can still work on our model.

###Taking log transformation of the dependent variable with selected features

```{r}
# Taking log transformation of the dependent variable with selected features
options(scipen=999)
crime_data$So<-as.factor(crime_data$So)
# Linear Regression model with reduced # of features
lmodel_3<-lm(log(Crime)~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data)
summary(lmodel_3)
# plotting the residuals to check for assumptions in the plain vanilla model with all features
par(mfrow = c(2, 2))
plot(lmodel_3)

crime_rate_pred<-predict(lmodel_3,newdata = new_datapoint)
crime_rate_pred<- exp(crime_rate_pred)
print(paste("Predicted Value:", round(crime_rate_pred,2)))
```

[**#Analysis**]{style="color:red"}

* We can see that by removing a few predicting features, the value of $R_{2}$ has gotten worse. We are not able to explain much on the basis of our predicting feature.
* We also observe that our predicted value for the new data point has gotten closer to the mean response.
* Another thing to observe is the P-value for one feature ("Po1") has gotten significantly small and we can reject the null hypothesis. Thus it is a significant predictor for our model.

```{r}
plot(crime_data$Po1 ~ log(crime_data$Crime), main="Plot for one significant feature (Po1) v/s response (log(Crime))")
abline(lm(crime_data$Po1 ~ log(crime_data$Crime)))
```

###Removing the potential outliers/influential points based on the diagnostic plots of linear regression model

```{r}
# removing the potential outliers/influential points based on the diagnostic plots of linear regression model
crime_data_wo<-crime_data[-c(11, 19, 46),]

# Taking log transformation of the dependent variable with all features
options(scipen=999)
crime_data$So<-as.factor(crime_data$So)
# Linear Regression model with reduced # of features and removed potential influential points.
lmodel_4<-lm(log(Crime)~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data)
summary(lmodel_4)
# plotting the residuals to check for assumptions in the plain vanilla model with all features
par(mfrow = c(2, 2))
plot(lmodel_4)

crime_rate_pred<-predict(lmodel_4,newdata = new_datapoint)
crime_rate_pred<- exp(crime_rate_pred)
print(paste("Predicted Value:", round(crime_rate_pred,2)))

```

[**#Analysis**]{style="color:red"}

* As we can observe that after removing the potential outliers, there is no change in the model. Thus we can conclude that our assumption to classify the removed points as outliers is wrong.

### 2022-09-21 hw4-1/anshit/anshit.Rmd

--- 2022-09-21 hw4-1/Submission.Rmd
@@
 ---
-title: "Submission HW2"
-output:
-  html_document:
-    df_print: paged
-  pdf_document: default
+title: "Homework-4"
+author: "Anshit Verma"
+date: "28-09-2022"
 ---
-
-# ISYE 6501 \| Fall 22
-
--   Ashish Dhiman \| [ashish.dhiman\@gatech.edu](mailto:ashish.dhiman@gatech.edu){.email}
-
--   Abhinav Arun \| [aarun60\@gatech.edu](mailto:aarun60gatech.edu){.email}
-
--   Anshit Verma \| [averma373\@gatech.edu](mailto:averma373@gatech.edu){.email}
-
-[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
-
-## Question 5.1
-
-```{r setup, include=FALSE}
-library(outliers)
-library(ggplot2)
-library(qqplotr)
-library(cowplot)
-library(reshape2)
-library(grid)
-```
-
-```{bash}
-pwd
-head -4 ./uscrime.txt
-```
-
-### EDA of Crime Data
-
-```{r}
-crime_df = read.table("./uscrime.txt", sep="\t", header= TRUE)
-dim(crime_df)
-```
-
-3 states are missing in the data, we don't know which.
-
-#### Data and variable meaning:
-
-Criminologists are interested in the effect of punishment regimes on crime rates. This has been studied using aggregate data on 47 states of the USA for 1960.
-
-| Variable | Description                                                                            |
-|-------------------|-----------------------------------------------------|
-| M        | percentage of males aged 14--24 in total state population                              |
-| So       | indicator variable for a southern state                                                |
-| Ed       | mean years of schooling of the population aged 25 years or over                        |
-| Po1      | per capita expenditure on police protection in 1960                                    |
-| Po2      | per capita expenditure on police protection in 1959                                    |
-| LF       | labour force participation rate of civilian urban males in the age-group 14-24         |
-| M.F      | number of males per 100 females                                                        |
-| Pop      | state population in 1960 in hundred thousands                                          |
-| NW       | percentage of nonwhites in the population                                              |
-| U1       | unemployment rate of urban males 14--24                                                |
-| U2       | unemployment rate of urban males 35--39                                                |
-| Wealth   | wealth: median value of transferable assets or family income                           |
-| Ineq     | income inequality: percentage of families earning below half the median income         |
-| Prob     | probability of imprisonment: ratio of number of commitments to number of offenses      |
-| Time     | average time in months served by offenders in state prisons before their first release |
-| Crime    | crime rate: number of offenses per 100,000 population in 1960                          |
-
-```{r}
-summary(crime_df$Crime)
-```
-
-### Difference b/w Southern and Northern states
-
-```{r}
-aggregate(Pop ~ So, crime_df, sum)
-```
-
-```{r}
-south_split = aggregate(cbind(Crime,Ineq,M,M.F,U1,Time,NW,Ed) ~ So, crime_df, mean)
-south_split
-```
-
-We see that Southern states in 1960s(on average):
-
--   had \<50% of non So population
-
--   were more inequal
-
--   had greater proportion of NW
-
--   lower mean years of schooling
-
-```{r}
-#Plot crime by population
-plot(x=crime_df$Pop,y=crime_df$Crime)
-```
-
-The two extreme large points seem like potentail outliers
-
-### Outliers package
-
-Grubbs' test ([Grubbs 1969](https://www.itl.nist.gov/div898/handbook/eda/section4/eda43.htm#Grubbs) and [Stefansky 1972](https://www.itl.nist.gov/div898/handbook/eda/section4/eda43.htm#Stefansky)) is used to detect a single [outlier](https://www.itl.nist.gov/div898/handbook/eda/section3/eda35h.htm) in a univariate data set that follows an [approximately normal](https://www.itl.nist.gov/div898/handbook/eda/section3/eda35h.htm#Normality) distribution.
-
-| H~0~: | There are no outliers in the data set        |
-|-------|----------------------------------------------|
-| H~a~: | There is exactly one outlier in the data set |
-
-The two salient assumptions here are:
-
-1.  Data is Normally Distributed
-2.  There is exactly one outlier
-
-Before applying the test let us check for these two assumptions:
-
-#### Assumption 1: Normality
-
-```{r}
-hist(crime_df$Crime,breaks=25)
-```
-
-```{r}
-hist(log(crime_df$Crime),breaks=20)
-```
-
-Basis Histogram the data presents deviation from normality as the distribution is not symmetric, especially without the log transformation. We can further test this with QQ plot
-
-```{r}
-ggplot(mapping = aes(sample = crime_df$Crime)) + stat_qq_point(size = 2) + stat_qq_line(color="green")
-```
-
-```{r}
-ggplot(mapping = aes(sample = log(crime_df$Crime))) + stat_qq_point(size = 2) + stat_qq_line(color="green")
-```
-
-Similarly from QQ plot, we see that the points diverge from line especially at the extremes, hinting at non-normality in that region.
-While log of crime is relatively more normal, the log transformation alters the scale of data, and we might want to detect outliers before applying it.
-
-**Since assumption 1 is not fully satisfied, we can not be 100% sure of the efficiency of Grubbs Test in our case.**
-
-#### Assumption 2: Exactly one outlier
-
-To check this assumption let us check the box plot of crime:
-
-```{r}
-boxplot(crime_df$Crime)
-```
-
-As seen in scatter plot, there are two points that seem pretty far away from the whiskers.
-
-Let us further examine these two points:
-
-```{r}
-mask = crime_df$Crime > 1800
-crime_df[mask,"Crime"]
-```
-
-```{r}
-crime_df[mask,]
-```
-
-We see that among these two points, there is disproportionate difference of population, $157 * 10^5$ vs $3 * 10^5$.
-
-While the Crime variable is already normalised, number of offenses per 100,000 population in 1960.
-such high crime rate for such small population seems counter intuitive, especially so when all other relevant variables like Po1, Po2, Prob are fairly similar.
-Let us examine this in detail by checking corelation:
-
-```{r}
-paste(cor(crime_df$Pop,crime_df$Crime),cor(crime_df$Prob,crime_df$Crime))
-```
-
-Basis population figures in 1960's, there are two candidate states for 1993 point: Wyoming & Nevada.
-
-And comparing data from [Disatster Center](https://www.disastercenter.com/crime/wycrime.htm), Wyoming does have roughly twice the national average crime rate, which is similar to our data as well (1993/ 905.1).
-Maybe there was some particular event in 1960 in Wyoming, which is causing such behavior.
-To understand this better, we need some expertise of Crime rate Trends in Wyoming.
-
-**Hence this point maybe outlier but there is high likelihood that the underlying data is true, and the decision of dropping it would depend on our exact modelling needs.**
-
-<https://en.wikipedia.org/wiki/1960_United_States_census#:~:text=The%20United%20States%20census%20of,enumerated%20during%20the%201950%20census.>
-
-**Since assumption 2 is again not fully satisfied, we can not be 100% sure of the efficiency of Grubbs Test in our case.**
-
-### Apply Grubbs Test
-
-```{r}
-grubbs.test(crime_df$Crime, type = 10, opposite = FALSE, two.sided = FALSE)
-```
-
-As per Grubbs test, we get p-value of \~8%.
-$\implies$ there is 8% chance that the given point is the effect of randomness, assuming Null Hypothesis is True.
-Generally we want p-value to be as low as possible to reject the Null hypothesis.In this case 8% value is slightly larger than typical thresholds of 5%, and warrants detailed investigation before classifying the point as outlier, which is in sync with analysis above.
-
-Let us also test what happens to the 2nd highest point, if the 1993 pt is removed
-
-```{r}
-mask1 = crime_df$Crime<1990
-grubbs.test(crime_df[mask1,]$Crime, type = 10, opposite = FALSE, two.sided = FALSE)
-```
-
-After removing 1993 point we get even lower p-value this time, hinting that 1969 in absence of 1993 is a much stronger candidate for being a outlier.
-
-Thus in conclusion, we have two candidate outliers in the data, 1993 and 1969 point.
-1993 point is the highest crime rate and presents weak evidence to discard it as outlier, however, to be definitive, we need to study the Wyoming angle in detail, especially in relevance to exact modelling needs at hand.
 
@@
 
-**Objective Function (Y)**: To predict the selling price of a property based in the predicting variables for that property
-
+**Objective Function (Y)**: To predict the selling price of a preperty based in the predicting variables for that property
 **Predicting Variables (X_i)**: Some of the key predicting variables are given below -
@@
 It can also be considered as the condition metric of the house.
-A house with good condition grade can worth more than one with lower condition grade as it might incur lower maintenance cost.
+A house with good condition grade can worth more than one with lower condition grade as it might incure lower maintenance cost.
 
@@
 The ability of the garage to hold more cars is an important factor which appreciates the value of a property
-
-## Question 8.2
-
-```{r crimes}
-crime_data<-read.table(file = "./uscrime.txt", sep = "\t",header=TRUE)
-head(crime_data)
-dim(crime_data)
-summary(crime_data)
-```
-
-### Data Distribution and Pairwise Corelation
-```{r}
-my_plots <- lapply(names(crime_data), function(var_x){
-  p <- 
-    ggplot(crime_data) +
-    aes_string(var_x)
-
-  if(var_x %in% list("So")) {
-    p <- p + geom_bar()
-
-  } else {
-    p <- p + geom_density()
-  }
-
-})
-
-plot_grid(plotlist = my_plots)
-
-### Correlation
-cormat <- round(cor(crime_data),2)
-cormat[upper.tri(cormat)] <- NA
-melted_cormat <- melt(cormat)
-# plotting the correlation heatmap
-
-ggplot(data = melted_cormat, aes(x=Var1, y=Var2,
-                                   fill=value)) +
-geom_tile() +
-geom_text(aes(Var1, Var2, label = value),
-          color = "white", size = 3)
-```
-
-[**#Analysis**]{style="color:red"}
-
-Wealth and Income Inequality are 2 highly correlated variables (-ve correlation), hence we can pick only 1 variable out of the 2.
-Po1 and Po2 are 2 highly correlated variables (+ve correlation), hence we can pick only 1 out of 2. 
-Same goes for U1 and U2, hence we can pick only 1 out of 2
-
-The dependent variable Crime Rate has high correlation with the following variables :
-Po1 and Po2 -> we see a surprising +ve correlation with crime rate.
-As regression cannot be used for defining a causal relation , we can infer the other way round that since the crime rate is higher, that could be the reason of a higher per capita expenditure on police protection.
-
-Prob -> makes sense to have -ve correlation in this case. As the crime rate is bound to decrease if probability of imprisonment increases.
-
-###Linearity in data between independent and dependent variables
-
-```{r}
-plot(crime_data$Po1 ~ crime_data$Crime)
-```
-
-[**#Analysis**]{style="color:red"}
-
-From the aove plot we can observe that there seems to be a linear relationship between the dependent variable and one of the highly correlated feature.
-Thus our assumption of linearity holds true and we can implement linear regression model.
-
-###Initial Model
-
-```{r}
-options(scipen=999)
-crime_data$So<-as.factor(crime_data$So)
-# Linear Regression model with all features
-lmodel_1<-lm(Crime~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data)
-summary(lmodel_1)
-# plotting the residuals to check for assumptions in the plain vanilla model with all features
-par(mfrow = c(2, 2))
-plot(lmodel_1)
-
-new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
-crime_rate_pred<-predict(lmodel_1,newdata = new_datapoint)
-crime_rate_pred<-round(crime_rate_pred,2)
-print(paste("Predicted Value:", crime_rate_pred))
-```
-
-[**#Analysis**]{style="color:red"}
-
-**Interpreting the output of the linear regression model** : 
-Based on the model output , it appears that the features Ed, Ineq, Prob, M, Po1 are significant variables based on the p-values . \
-* The lower the p-value , the more significant the variable  because the p-value signifies the probability of the estimated coefficients $\hat{\beta_{i}}'s$ coming from the underlying assumed distribution of null hypothesis (Null Hypothesis states that $\beta_{i}'s=0$).
-Thus , lower the p-value , higher the probability that the coefficient of the variable is not equal to 0 and hence that variable is a good predictor.\
-
-* For this model, the $R^{2}$ value is 0.8 which means that 80% of the variance in the dependent variable is explained by the set of these independent variables .\
-
-* The predicted crime_rate for the given data point is `r crime_rate_pred`.
-Since this prediction lies outside the range of or dependent variable, we can further tune the model as the features of the predicting data point lie withing the domain of independent variables.
-\
-
-* From the 4 diagnostic plots associated with the linear regression model, we can see that the assumptions are mostly in place, error terms are normally distributed i.e.
-constant variance assumption and residuals are homogeneous around the fitted line.
-However , we can see slight bends in the left and right end in the plot of residuals vs fitted values.
-Thus, a possible remedy can be thought of as transforming the dependent variable (taking a log transformation reduces the range of possible variables).
-
-* Further from the QQ Plot it can be inferred that the data is normally distributed. Thus our normality assumption also holds true.
-
-**Points for further exploration** : 
-* Taking a log transformation of the dependent variable and then create a regression model.
-* Create a model with lower number of features (remove highly correlated features).
-* Look for possible outliers/leverage points (based on the residuals vs leverage plot - cook's distance) and remove them to see changes in the model output.
-
-###Taking log transformation of the dependent variable with all features
-
-```{r}
-# Taking log transformation of the dependent variable with all features
-options(scipen=999)
-crime_data$So<-as.factor(crime_data$So)
-# Linear Regression model with all features
-lmodel_2<-lm(log(Crime)~M+So+Ed+Po1+Po2+LF+M.F+Pop+NW+U1+U2+Wealth+Ineq+Prob+Time,data=crime_data)
-summary(lmodel_2)
-# plotting the residuals to check for assumptions in the plain vanilla model with all features
-par(mfrow = c(2, 2))
-plot(lmodel_2)
-
-crime_rate_pred<-predict(lmodel_2,newdata = new_datapoint)
-crime_rate_pred<- exp(crime_rate_pred)
-print(paste("Predicted Value:", round(crime_rate_pred,2)))
-```
-
-[**#Analysis**]{style="color:red"}
-
-Although the $R^{2}$ takes a slight dip, but it appears that the model assumptions are better satisfied here .\ 
-Points 11,19,46 seems to have around 1.5*standard deviation, thus these could be tested as potential outliers. \
-
-Again, the predicted crime_rate for the given data point is `r crime_rate_pred`. The value has gotten within the range of dependent variable. But we can still work on our model.
-
-###Taking log transformation of the dependent variable with selected features
-
-```{r}
-# Taking log transformation of the dependent variable with selected features
-options(scipen=999)
-crime_data$So<-as.factor(crime_data$So)
-# Linear Regression model with reduced # of features
-lmodel_3<-lm(log(Crime)~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data)
-summary(lmodel_3)
-# plotting the residuals to check for assumptions in the plain vanilla model with all features
-par(mfrow = c(2, 2))
-plot(lmodel_3)
-
-crime_rate_pred<-predict(lmodel_3,newdata = new_datapoint)
-crime_rate_pred<- exp(crime_rate_pred)
-print(paste("Predicted Value:", round(crime_rate_pred,2)))
-```
-
-[**#Analysis**]{style="color:red"}
-
-* We can see that by removing a few predicting features, the value of $R_{2}$ has gotten worse. We are not able to explain much on the basis of our predicting feature.
-* We also observe that our predicted value for the new data point has gotten closer to the mean response.
-* Another thing to observe is the P-value for one feature ("Po1") has gotten significantly small and we can reject the null hypothesis. Thus it is a significant predictor for our model.
-
-```{r}
-plot(crime_data$Po1 ~ log(crime_data$Crime), main="Plot for one significant feature (Po1) v/s response (log(Crime))")
-abline(lm(crime_data$Po1 ~ log(crime_data$Crime)))
-```
-
-###Removing the potential outliers/influential points based on the diagnostic plots of linear regression model
-
-```{r}
-# removing the potential outliers/influential points based on the diagnostic plots of linear regression model
-crime_data_wo<-crime_data[-c(11, 19, 46),]
-
-# Taking log transformation of the dependent variable with all features
-options(scipen=999)
-crime_data$So<-as.factor(crime_data$So)
-# Linear Regression model with reduced # of features and removed potential influential points.
-lmodel_4<-lm(log(Crime)~M+So+Ed+Po1+LF+M.F+Pop+NW+U1+Ineq+Prob+Time,data=crime_data)
-summary(lmodel_4)
-# plotting the residuals to check for assumptions in the plain vanilla model with all features
-par(mfrow = c(2, 2))
-plot(lmodel_4)
-
-crime_rate_pred<-predict(lmodel_4,newdata = new_datapoint)
-crime_rate_pred<- exp(crime_rate_pred)
-print(paste("Predicted Value:", round(crime_rate_pred,2)))
-
-```
-
-[**#Analysis**]{style="color:red"}
-
-* As we can observe that after removing the potential outliers, there is no change in the model. Thus we can conclude that our assumption to classify the removed points as outliers is wrong.

### 2022-09-26 hw4-1/Submission.Rmd
box-cox addition

@@
 ---
-title: "Submission HW2"
+title: "Submission HW4"
 output:
@@
 
-From the aove plot we can observe that there seems to be a linear relationship between the dependent variable and one of the highly correlated feature.
-Thus our assumption of linearity holds true and we can implement linear regression model.
+From the above plot we can observe that there seems to be a linear relationship between the dependent variable and one of the highly correlated feature.
 
@@
 * The predicted crime_rate for the given data point is `r crime_rate_pred`.
-Since this prediction lies outside the range of or dependent variable, we can further tune the model as the features of the predicting data point lie withing the domain of independent variables.
+We noticed that this prediction lies outside the range of our dependent variable, and we decided to further tune our model and observe the changes.
 \
 
-* From the 4 diagnostic plots associated with the linear regression model, we can see that the assumptions are mostly in place, error terms are normally distributed i.e.
-constant variance assumption and residuals are homogeneous around the fitted line.
+* From the 4 diagnostic plots associated with the linear regression model , we can see that the assumptions are mostly in place although there are some slight deviations, error terms are normally distributed (deviation for a couple of points) and residuals are homogeneous around the fitted line (no heteroskedasticity).
 However , we can see slight bends in the left and right end in the plot of residuals vs fitted values.
-Thus, a possible remedy can be thought of as transforming the dependent variable (taking a log transformation reduces the range of possible variables).
+Thus , a possible remedy can be thought of as modeling the log of dependent variable based on independent variable.
+(because taking a log transformation reduces the range of possible variables).\
 
-* Further from the QQ Plot it can be inferred that the data is normally distributed. Thus our normality assumption also holds true.
+* Further from the QQ Plot it can be inferred that the data is mostly normally distributed (deviation for a couple of points) .
+Thus our normality assumption mostly holds true.
+However , we have also tried out box cox transformation technique to determine the best transformed y to regress on so that normality assumptions hold true for a couple of deviating points too.\
 
 **Points for further exploration** : 
-* Taking a log transformation of the dependent variable and then create a regression model.
-* Create a model with lower number of features (remove highly correlated features).
-* Look for possible outliers/leverage points (based on the residuals vs leverage plot - cook's distance) and remove them to see changes in the model output.
+* Taking a log transformation of the dependent variable and then create a regression model.\
+* Create a model with lower number of features (remove highly correlated features).\
+* Look for possible outliers/leverage points (based on the residuals vs leverage plot - cook's distance) and remove them to see changes in the model output.\ 
+* We can use Box-Cox transformation to identify the best possible transformation of y to regress on . This is usually done to handle the points deviating a bit from the normal Q-Q plot.\
+
+```{r}
+# Box Cox Transformation (we can try this out to find the best possible function of y to regress on , the value of lambda is determined through MLE)
+bc<-boxcox(lmodel_1, lambda =seq(-3,3))
+#extract the best lambda 
+best.lam<-bc$x[which(bc$y==max(bc$y))]
+
+```
+# Observations: 
+Since , the above graph gives us the best value of $\lambda$ to be around 0 only , therefore taking a log transformation of the dependent variable y makes more sense.
 
@@
 
-* As we can observe that after removing the potential outliers, there is no change in the model. Thus we can conclude that our assumption to classify the removed points as outliers is wrong.
+* As we can observe that after removing the potential outliers, there is no major change in the model.
+Thus we can conclude that our assumption to classify the removed these set of points as outliers maybe not be correctly justified.

### 2022-09-27 hw5-1/ashish/a.0_random_forest.Rmd
rf v1

---
title: "Random Forest | Ques 10.1"
author: "Ashish Dhiman"
date: "2022-09-26"
output: html_document
---

```{bash}
ls ../
```

```{r, warning=FALSE}
library(randomForest)
```

```{r}
crime_df = read.table("../uscrime.txt", sep="\t", header= TRUE)
dim(crime_df)
```

#### Data and variable meaning:

Criminologists are interested in the effect of punishment regimes on crime rates. This has been studied using aggregate data on 47 states of the USA for 1960.

| Variable | Description                                                                            |
|-----------------|-------------------------------------------------------|
| M        | percentage of males aged 14--24 in total state population                              |
| So       | indicator variable for a southern state                                                |
| Ed       | mean years of schooling of the population aged 25 years or over                        |
| Po1      | per capita expenditure on police protection in 1960                                    |
| Po2      | per capita expenditure on police protection in 1959                                    |
| LF       | labour force participation rate of civilian urban males in the age-group 14-24         |
| M.F      | number of males per 100 females                                                        |
| Pop      | state population in 1960 in hundred thousands                                          |
| NW       | percentage of nonwhites in the population                                              |
| U1       | unemployment rate of urban males 14--24                                                |
| U2       | unemployment rate of urban males 35--39                                                |
| Wealth   | wealth: median value of transferable assets or family income                           |
| Ineq     | income inequality: percentage of families earning below half the median income         |
| Prob     | probability of imprisonment: ratio of number of commitments to number of offenses      |
| Time     | average time in months served by offenders in state prisons before their first release |
| Crime    | crime rate: number of offenses per 100,000 population in 1960                          |

#Vanilla RF model

```{r}
lapply(crime_df, class)
```

```{r}
crime_df$So <- factor(crime_df$So)
```

```{r}
rf_model_v0 = randomForest(Crime ~ ., data = crime_df)
summary(rf_model_v0)
```

```{r}
rf_model_v0
```

```{r}
rmse_func = function(true,predicted){
  (mean((true-predicted)^2))**0.5
}

rmse_func(true = crime_df$Crime, predicted=rf_model_v0$predicted)
```

```{r}
rf_model_v1 = randomForest(Crime ~ ., data = crime_df, ntree = 100)
rmse_func(true = crime_df$Crime, predicted=rf_model_v1$predicted)
```

```{r}
rf_model_v2 = randomForest(log(Crime) ~ ., data = crime_df, ntree = 100)
rmse_func(true = crime_df$Crime, predicted=exp(rf_model_v2$predicted))
```

```{r}
varImpPlot(rf_model_v2)
```

```{r}
rmse_list = sapply(seq(100,1000,20), function(x)  rmse_func(true = crime_df$Crime, predicted=randomForest(Crime ~ ., data = crime_df, ntree = x)$predicted))
plot(x=seq(100,1000,20), y=rmse_list, main = "MSE vs ntrees")
```

```{r}
rmse_list = sapply(seq(2,8,1), function(x)  rmse_func(true = crime_df$Crime, predicted=randomForest(Crime ~ ., data = crime_df, mtry = x)$predicted))
plot(x=seq(2,8,1), y=rmse_list,main="RMSE vs varaibles split")
```

```{r}
rmse_list = sapply(seq(200,1000,20), function(x)  rmse_func(true = crime_df$Crime, predicted=randomForest(Crime ~ ., data = crime_df, ntree = x, mtry=3)$predicted))
plot(x=seq(200,1000,20), y=rmse_list, main = "MSE vs ntrees")
```

```{r}
rf_model_v3 = randomForest(Crime ~ ., data = crime_df, ntree = 250, mtry=3, proximity = TRUE)
rf_model_v3
```

```{r}
prox1=rf_model_v3$proximity
```

### CV

```{r}
n_folds = 5
folds <- sample(rep(1:n_folds, length.out = nrow(crime_df)), size = nrow(crime_df), replace = F)

table(folds)
```

```{r}
rmse_func = function(true,predicted){
  return (mean((true-predicted)^2)**0.5)
}

CV_rf_func = function(test_fold, ntree1, mtry1){
  model <- randomForest(Crime ~ ., data = crime_df[folds != test_fold,],
                        ntree = ntree1, mtry=mtry1)
  #print(model$ntree)
  preds <- predict(model,  crime_df[folds == test_fold,])
  return(rmse_func(true = crime_df[folds == test_fold,"Crime"],predicted = preds))
}
```

```{r}
rmse_list=c()

for (ntree_i in seq(10,400,50)) {
  for (mtry_i in seq(2,15,1)) {
    #RMSE on each CV Fold for ntree_i,mtry_i
    rmse_cv = mean(unlist(lapply(seq(1:5), function(fold) 
      CV_rf_func(test_fold = fold, ntree1 = ntree_i, mtry1 = mtry_i))))
    print (paste("For tree:",ntree_i,"mtry:",mtry_i,"rmse:",rmse_cv))
    rmse_list = c(rmse_list, rmse_cv)
}
}

rmse_df = cbind.data.frame(seq(10,400,50),seq(2,15,1),rmse_list)
```

```{r}
names(rmse_df) = c("ntree","mtry","rmse")
```

```{r}
rmse_df[which.min(rmse_df$rmse),]
```

```{r}
summary(rmse_df$rmse)
```

### 2022-09-28 hw5-1/ashish/a.0_random_forest.Rmd
submission RF v0

@@
 | Variable | Description                                                                            |
-|-----------------|-------------------------------------------------------|
+|-------------------|-----------------------------------------------------|
 | M        | percentage of males aged 14--24 in total state population                              |
@@
 ```{r}
-rf_model_v1 = randomForest(Crime ~ ., data = crime_df, ntree = 100)
+rf_model_v1 = randomForest(Crime ~ ., data = crime_df, ntree = 100,importance=TRUE)
 rmse_func(true = crime_df$Crime, predicted=rf_model_v1$predicted)
@@
 ```{r}
-rf_model_v2 = randomForest(log(Crime) ~ ., data = crime_df, ntree = 100)
+rf_model_v2 = randomForest(log(Crime) ~ ., data = crime_df, ntree = 100,importance=TRUE)
 rmse_func(true = crime_df$Crime, predicted=exp(rf_model_v2$predicted))
@@
 ```{r}
-varImpPlot(rf_model_v2)
+varImpPlot(rf_model_v1,sort = TRUE)
+```
+
+```{r}
+rf_model_v3 = randomForest(Crime ~ . - Ineq -So -Pop, data = crime_df, ntree = 100)
+rmse_func(true = crime_df$Crime, predicted=rf_model_v3$predicted)
 ```

### 2022-09-28 hw5-1/ashish/a.1_submission_rf.Rmd
submission RF v0

---
title: "Random Forest | Ques 10.1"
author: "Ashish Dhiman"
date: "2022-09-26"
output: html_document
---

### Read Data

```{r}
set.seed(77)
library(randomForest)

crime_df = read.table("../uscrime.txt", sep="\t", header= TRUE)
dim(crime_df)
```

#### Data and variable meaning:

Criminologists are interested in the effect of punishment regimes on crime rates. This has been studied using aggregate data on 47 states of the USA for 1960.

| Variable | Description                                                                            |
|--------------------|----------------------------------------------------|
| M        | percentage of males aged 14--24 in total state population                              |
| So       | indicator variable for a southern state                                                |
| Ed       | mean years of schooling of the population aged 25 years or over                        |
| Po1      | per capita expenditure on police protection in 1960                                    |
| Po2      | per capita expenditure on police protection in 1959                                    |
| LF       | labour force participation rate of civilian urban males in the age-group 14-24         |
| M.F      | number of males per 100 females                                                        |
| Pop      | state population in 1960 in hundred thousands                                          |
| NW       | percentage of nonwhites in the population                                              |
| U1       | unemployment rate of urban males 14--24                                                |
| U2       | unemployment rate of urban males 35--39                                                |
| Wealth   | wealth: median value of transferable assets or family income                           |
| Ineq     | income inequality: percentage of families earning below half the median income         |
| Prob     | probability of imprisonment: ratio of number of commitments to number of offenses      |
| Time     | average time in months served by offenders in state prisons before their first release |
| Crime    | crime rate: number of offenses per 100,000 population in 1960                          |

```{r}
crime_df$So <- factor(crime_df$So)
```

### Build Vanilla RF model

```{r}
rf_model_v0 = randomForest(Crime ~ ., data = crime_df)
summary(rf_model_v0)

rmse_func = function(true,predicted){
  (mean((true-predicted)^2))**0.5
}

rmse_func(true = crime_df$Crime, predicted=rf_model_v0$predicted)

varImpPlot(rf_model_v1,sort = TRUE)
```

From variable importance we can see that the some variables like, Ineq, So and Pop are not very helpful. In the CV stage below, we will test models without these variables.

```{r}
rf_model_v1 = randomForest(Crime ~ . -Ineq -U1, data = crime_df, importance = TRUE)
varImpPlot(rf_model_v1)
```

### Build RF and tune hyperparams with CV

#### Build CV folds

```{r}
n_folds = 5
folds <- sample(rep(1:n_folds, length.out = nrow(crime_df)), size = nrow(crime_df), replace = F)

table(folds)
```

We have created uniform CV folds here

#### Function for RF on CV fold

```{r}
CV_rf_func = function(test_fold, ntree1, mtry1){
  model1 <- randomForest(Crime ~ ., data = crime_df[folds != test_fold,],
                        ntree = ntree1, mtry=mtry1)
  model2 <- randomForest(Crime ~ . -Ineq -U1, data = crime_df[folds != test_fold,],
                        ntree = ntree1, mtry=mtry1)
  model3 <- randomForest(Crime ~ . -Ineq - Pop - U2, data = crime_df[folds != test_fold,],
                        ntree = ntree1, mtry=mtry1)

  preds1 <- predict(model1,  crime_df[folds == test_fold,])
  preds2 <- predict(model2,  crime_df[folds == test_fold,])
  preds3 <- predict(model3,  crime_df[folds == test_fold,])

  rmse1 = rmse_func(true = crime_df[folds == test_fold,"Crime"],predicted = preds1)
  rmse2 = rmse_func(true = crime_df[folds == test_fold,"Crime"],predicted = preds2)
  rmse3 = rmse_func(true = crime_df[folds == test_fold,"Crime"],predicted = preds3)

  return(c(rmse1,rmse2,rmse3))
}

```

```{r, warning=FALSE, message=FALSE}
rmse_list1=c()
rmse_list2=c()
rmse_list3=c()

for (ntree_i in seq(30,130,10)) {
  for (mtry_i in seq(2,12,1)) {
    #RMSE on each CV Fold for ntree_i,mtry_i
    rmse_cv_all = lapply(seq(1:5), function(fold) 
      CV_rf_func(test_fold = fold, ntree1 = ntree_i, mtry1 = mtry_i))
    rmse_cv_df = do.call(rbind.data.frame, rmse_cv_all)
    rmse_cv1 = mean(unlist(rmse_cv_df[,1]))
    rmse_cv2 = mean(unlist(rmse_cv_df[,2]))
    rmse_cv3 = mean(unlist(rmse_cv_df[,3]))
    print (paste("For tree:",ntree_i,"mtry:",mtry_i))
    rmse_list1 = c(rmse_list1, rmse_cv1)
    rmse_list2 = c(rmse_list2, rmse_cv2)
    rmse_list3 = c(rmse_list3, rmse_cv3)
}
}
```

```{r}
rmse_df = cbind.data.frame(seq(30,130,10),seq(2,12,1),
                           rmse_list1,rmse_list2,rmse_list3)
names(rmse_df)=c("ntree","mtry","rmse_model1","rmse_model2","rmse_model3")
```

```{r}
rmse_df
```

```{r}
print ("Min RMSE model 1")
rmse_df[which.min(rmse_df[,"rmse_model1"]),]

print ("Min RMSE model 2")
rmse_df[which.min(rmse_df[,"rmse_model2"]),]

print ("Min RMSE model 3")
rmse_df[which.min(rmse_df[,"rmse_model3"]),]
```

### Best Model is therefore given for:

1.  Dropping variables:
2.  number of trees = 60
3.  mtry = 5

Not only thus by RMSE, but by intution too this is a simpler model since it has lesser features, lower number of trees, as well as lower number of features to split on in RF.

Using the above hyper-parameters we can build the best model on full data.

```{r}
rf_final_model <- randomForest(Crime ~ . -Ineq -Pop, data = crime_df, ntree = 40, mtry=3, importance = TRUE)
rmse_func(true=crime_df$Crime,predicted = rf_final_model$predicted)
```

```{r}
varImpPlot(rf_final_model)
```

### 2022-09-28 hw5-1/ashish/a.1_submission_rf.Rmd
submission draft RF

@@
 
-```{r}
+```{r, warning=FALSE}
 set.seed(77)
 library(randomForest)
+library(ggplot2)
 
@@
 ```{r}
-rf_model_v0 = randomForest(Crime ~ ., data = crime_df)
+rf_model_v0 = randomForest(Crime ~ ., data = crime_df, importance=TRUE)
 summary(rf_model_v0)
@@
 
-varImpPlot(rf_model_v1,sort = TRUE)
+varImpPlot(rf_model_v0,sort = TRUE)
 ```
 
-From variable importance we can see that the some variables like, Ineq, So and Pop are not very helpful. In the CV stage below, we will test models without these variables.
+From variable importance we can see that the some variables like, So, U1 and U2 are not very helpful.
+Hence it might serve us better to drop these variables, bcos they might add more noise to our model instead of actual information capture.
+In the CV stage below, we will test the models without these variables.
 
 ```{r}
-rf_model_v1 = randomForest(Crime ~ . -Ineq -U1, data = crime_df, importance = TRUE)
+rf_model_v1 = randomForest(Crime ~ . -U1 -M.F -U2, data = crime_df, importance = TRUE)
 varImpPlot(rf_model_v1)
@@
                         ntree = ntree1, mtry=mtry1)
-  model2 <- randomForest(Crime ~ . -Ineq -U1, data = crime_df[folds != test_fold,],
+  model2 <- randomForest(Crime ~ . -U1 -M.F -U2, data = crime_df[folds != test_fold,],
                         ntree = ntree1, mtry=mtry1)
-  model3 <- randomForest(Crime ~ . -Ineq - Pop - U2, data = crime_df[folds != test_fold,],
+  model3 <- randomForest(Crime ~ . -U1 -U2 -Time, data = crime_df[folds != test_fold,],
                         ntree = ntree1, mtry=mtry1)
@@
 rmse_list3=c()
+ntree_list=c()
+mtry_list=c()
 
 for (ntree_i in seq(30,130,10)) {
-  for (mtry_i in seq(2,12,1)) {
+  for (mtry_i in seq(2,6,1)) {
     #RMSE on each CV Fold for ntree_i,mtry_i
@@
     rmse_list3 = c(rmse_list3, rmse_cv3)
+    ntree_list = c(ntree_list, ntree_i)
+    mtry_list = c(mtry_list, mtry_i)
 }
@@
 ```{r}
-rmse_df = cbind.data.frame(seq(30,130,10),seq(2,12,1),
+rmse_df = cbind.data.frame(ntree_list,mtry_list,
                            rmse_list1,rmse_list2,rmse_list3)
@@
 ```
+
+### Let us study the impact of these two hyper-parameters on RMSE
+
+#### Impact of ntree
+
+```{r}
+ggplot(rmse_df[rmse_df$mtry<4,], aes(x = ntree, y = rmse_model1, group = mtry, color = mtry)) +
+  geom_point() + geom_line()
+
+ggplot(rmse_df[rmse_df$mtry<4,], aes(x = ntree, y = rmse_model2, group = mtry, color = mtry)) +
+  geom_point() + geom_line()
+
+ggplot(rmse_df[rmse_df$mtry<4,], aes(x = ntree, y = rmse_model3, group = mtry, color = mtry)) +
+  geom_point() + geom_line()
+```
+
+Because our data is so small; the effect of ntree and mtry is very sporadic, with high interaction between them. Also the trend varies as per the choice of the model(i.e the choice of predictors).
+
+Ideally we would have expected the RMSE to exhibit a 'U' shaped trend with out hyper parameters, where adding complexity to model (i.e.
+more num of trees to average upon or more features to split nodes upon) first decreases RMSE, but after a point it starts overfitting, and RMSE starts increasing again.
 
@@
 
+As
+
 ### Best Model is therefore given for:
 
-1.  Dropping variables:
-2.  number of trees = 60
-3.  mtry = 5
+1.  Dropping variables: U1, M.F and U2
+2.  number of trees = 40
+3.  mtry = 4
 
@@
 ```{r}
-rf_final_model <- randomForest(Crime ~ . -Ineq -Pop, data = crime_df, ntree = 40, mtry=3, importance = TRUE)
+rf_final_model <- randomForest(Crime ~ . -U1 -M.F -U2, data = crime_df, ntree = 40, mtry=4, importance = TRUE)
 rmse_func(true=crime_df$Crime,predicted = rf_final_model$predicted)
@@
 ```
+
+From the variable importance plot we see, that the following variable feature very highly in Importance:
+
+1.  Po2
+2.  NW
+3.  Prob
+
+### Comparison between Decision Tree and Random Forest
+
+Note: while the above 3 variables are also significant features here is the variable importance of variable "Pop".
+While this variable is very important for Decision trees, it is not that important for Random Forest.
+One possible reason for this could be the notion of randomness involved in RF.
+
+Another point to note here is that after Hyperparameter tuning, Decision tree is giving lower RMSE, and is most likely over-fitting the data-set, in comparison to Random Forest, and if we had a test data, RF model would most likely perform better, because of better generalization.

### 2022-09-28 hw5-1/abhinav/tree.Rmd
Decision Tree

```{r}
library(cowplot)
library(ggplot2)
library(reshape2)
library(grid)
library(MASS)
library(rpart)
library(rpart.plot)
library(mlr)
```

```{r,include=FALSE}
#install.packages("mlr")
```

```{r crimes}
crime_data<-read.table(file = "../uscrime.txt", sep = "\t",header=TRUE)
head(crime_data)
dim(crime_data)
summary(crime_data)
```
```{r}
### Data Distribution and Pairwise Corelation

my_plots <- lapply(names(crime_data), function(var_x){
  p <- 
    ggplot(crime_data) +
    aes_string(var_x)

  if(var_x %in% list("So")) {
    p <- p + geom_bar()

  } else {
    p <- p + geom_density()
  }

})

plot_grid(plotlist = my_plots)

### Correlation
cormat <- round(cor(crime_data),2)
cormat[upper.tri(cormat)] <- NA
melted_cormat <- melt(cormat)
# plotting the correlation heatmap

ggplot(data = melted_cormat, aes(x=Var1, y=Var2,
                                   fill=value)) +
geom_tile() +
geom_text(aes(Var1, Var2, label = value),
          color = "white", size = 3)
```

# Training and evaluating on the same training dataset (not a good method for generalizing) , better to do cross validation 
```{r}
# Vanilla decision tree with a fixed value of complexity parameter of 0.01 and minsplit of 20
reg_tree<-rpart(Crime~.,data=crime_data,method="anova",control=c(cp=0.01,minsplit=20))
rmse<-sqrt(sum((crime_data$Crime - predict(reg_tree,type=c("vector")))**2)/nrow(crime_data))
rmse
```

```{r}
rpart.plot(reg_tree)
```

* Observations & Points for further exploration:

+ The above decision tree uses 3 variables NW , Pop, LF and Po1 to create a decision tree. (2 of them Pop and Po1 have a linear correlation with dependent variable i.e. Crime)

+ One food for thought is that this is a regression tree and we are imputing the mean values of all the points that fall in a leaf or terminal node , so incase if we have an outlier , then those will drive the value of mean and hence cause higher RMSE or SSE. So , it would be good to see the effect of removing or capping outliers and then performing decision tree.

+ The smaller the value of complex parameter , the model is expected to have high variance as a split would be performed if we get such a marginal increase in entropy . Therefore , to prevent our model from overfitting amd having high variance , it is better if we not choose too small values of cp i.e. complexity parameter.

+ The default value of max depth parameter here is 30 but given that we have only 47 nodes in our dataset , it dfoes not make much sense to have tree models greater than or equal to 3 or 4 levels of depth

+ minsplit hyperparameter : The default value of this parameter is 20 i.e. a split would be performed if we have atleast 20 datapoints in a node , although it is generally good to have higher values of "minsplit" parameter to avoid overfitting but since in this problem statement , the total # of datapoints is only 47 , so this is one parameter where we can tweak in a bit .

+ minbucket hyperparameter : Similarly , this hyperparameter means  the number of datpoints in a terminal node and thus here also , it makes sense to have a smaller value for this hyperparameter as we do not have large dataset for training our model.

#Perform hyperparameter tuning minsplit and cp(complexity hyperparameter)
```{r}
folds <- sample(rep(1:5, length.out = nrow(crime_data)), size = nrow(crime_data), replace = F)
pred_table<-data.frame(min_split=numeric(0),cp=double(0),rmse=double(0))
for(j in seq(0.001,0.02,0.001)){
for(i in seq(5,25,2)){
  CV_mse <- lapply(1:5, function(x){ #5 corresponds to the number of folds defined earlier
  model <- rpart(Crime ~ ., data = crime_data[folds != x,],method="anova",control=c(minsplit=i,cp=j))
  preds <- predict(model,  crime_data[folds == x,], type="vector")
  sum_sse<-sum((crime_data[folds==x,"Crime"]-preds)**2)
  rmse<-sqrt(sum_sse/nrow(crime_data[folds==x,]))
  #total_rmse<-total_rmse+rmse
  return(rmse)
   })
pred_table[nrow(pred_table)+1,]=c(i,j,mean(unlist(CV_mse)))
}
}  
#rpart.plot(rpart(Crime ~ ., data = crime_data,method="anova",minsplit=10,cp=0.01))
#plot(pred_table)
```

# Identifying the optimal value of the hyperparameters cp and min_split
```{r}
options(scipen=999)
pred_table[which.min(pred_table$rmse),c("min_split","cp")]
```
* From the above output , we get the the optimal value of cp as 0.001 and min_split as 15
```{r}
reg_cv_tree<-rpart(Crime~.,data=crime_data,method="anova",control=c(cp=0.001,minsplit=15))
rpart.plot(reg_cv_tree)
```

* Note: The above set of optimal parameters reproduces the same decision tree that we had done initially with fixed values of these 2 parameters.This can be because of less # of datapoints. We can try reducing the number of minsplits which will lead to a higher variance model and we can try iterating to get the optimal maxDepth hyperparameter.

```{r}
pred_table<-data.frame(min_split=numeric(0),maxDepth=double(0),rmse=double(0))
for(j in seq(2,15,1)){
for(i in seq(5,25,2)){
  CV_mse <- lapply(1:5, function(x){ #5 corresponds to the number of folds defined earlier
  model <- rpart(Crime ~ ., data = crime_data[folds != x,],method="anova",control=c(minsplit=i,maxdepth=j))
  preds <- predict(model,  crime_data[folds == x,], type="vector")
  sum_sse<-sum((crime_data[folds==x,"Crime"]-preds)**2)
  rmse<-sqrt(sum_sse/nrow(crime_data[folds==x,]))
  #total_rmse<-total_rmse+rmse
  return(rmse)
   })
pred_table[nrow(pred_table)+1,]=c(i,j,mean(unlist(CV_mse)))
}
}  
#rpart.plot(rpart(Crime ~ ., data = crime_data,method="anova",minsplit=10,cp=0.01))
#plot(pred_table)
```

* Identifying the optimal value of the hyperparameter maxDepth (which comes out to be around 2 with min_split within range of 5-15 (rmse is same for different values of min_split))
```{r}
options(scipen=999)
pred_table[which.min(pred_table$rmse),c("maxDepth")]
```

```{r}
reg_tree_final<-rpart(Crime~.,data=crime_data,method="anova",control=c(cp=0.001,minsplit=15,maxdepth=2))
rmse_final<-sqrt(sum((crime_data$Crime - predict(reg_tree_final,type=c("vector")))**2)/nrow(crime_data))
rmse_final
```
#Conclusion : The RMSE value for the decision tree model using the optimal values of the hyperparameter is `r round(rmse_final,2)`

### 2022-09-28 hw5-1/submission.Rmd
submission file created

--- 2022-09-28 hw5-1/ashish/a.1_submission_rf.Rmd
@@
 ---
-title: "Random Forest | Ques 10.1"
-author: "Ashish Dhiman"
-date: "2022-09-26"
-output: html_document
+title: "ISYE 6501"
+output:
+  html_document:
+    df_print: paged
+  pdf_document: default
+date: "2022-09-04"
 ---
 
-### Read Data
+# Submission HW4 \| Fall 22
 
-```{r, warning=FALSE}
+-   Ashish Dhiman \| [ashish.dhiman\@gatech.edu](mailto:ashish.dhiman@gatech.edu){.email}
+
+-   Abhinav Arun \| [aarun60\@gatech.edu](mailto:aarun60gatech.edu){.email}
+
+-   Anshit Verma \| [averma373\@gatech.edu](mailto:averma373@gatech.edu){.email}
+
+[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
+
+```{r, warning=FALSE, message=FALSE}
 set.seed(77)
+
 library(randomForest)
 library(ggplot2)
-
-crime_df = read.table("../uscrime.txt", sep="\t", header= TRUE)
-dim(crime_df)
+library(cowplot)
+library(ggplot2)
+library(reshape2)
+library(grid)
+library(MASS)
+library(rpart)
+library(rpart.plot)
 ```
 
-#### Data and variable meaning:
+```{r}
+crime_data<-read.table(file = "./uscrime.txt", sep = "\t",header=TRUE)
+head(crime_data)
+dim(crime_data)
+summary(crime_data)
+```
 
-Criminologists are interested in the effect of punishment regimes on crime rates. This has been studied using aggregate data on 47 states of the USA for 1960.
+## EDA
 
-| Variable | Description                                                                            |
-|--------------------|----------------------------------------------------|
-| M        | percentage of males aged 14--24 in total state population                              |
-| So       | indicator variable for a southern state                                                |
-| Ed       | mean years of schooling of the population aged 25 years or over                        |
-| Po1      | per capita expenditure on police protection in 1960                                    |
-| Po2      | per capita expenditure on police protection in 1959                                    |
-| LF       | labour force participation rate of civilian urban males in the age-group 14-24         |
-| M.F      | number of males per 100 females                                                        |
-| Pop      | state population in 1960 in hundred thousands                                          |
-| NW       | percentage of nonwhites in the population                                              |
-| U1       | unemployment rate of urban males 14--24                                                |
-| U2       | unemployment rate of urban males 35--39                                                |
-| Wealth   | wealth: median value of transferable assets or family income                           |
-| Ineq     | income inequality: percentage of families earning below half the median income         |
-| Prob     | probability of imprisonment: ratio of number of commitments to number of offenses      |
-| Time     | average time in months served by offenders in state prisons before their first release |
-| Crime    | crime rate: number of offenses per 100,000 population in 1960                          |
+### Data Distribution and Pairwise Corelation
 
 ```{r}
+my_plots <- lapply(names(crime_data), function(var_x){
+  p <- 
+    ggplot(crime_data) +
+    aes_string(var_x)
+
+  if(var_x %in% list("So")) {
+    p <- p + geom_bar()
+
+  } else {
+    p <- p + geom_density()
+  }
+
+})
+
+plot_grid(plotlist = my_plots)
+
+### Correlation
+cormat <- round(cor(crime_data),2)
+cormat[upper.tri(cormat)] <- NA
+melted_cormat <- melt(cormat)
+# plotting the correlation heatmap
+
+ggplot(data = melted_cormat, aes(x=Var1, y=Var2,
+                                   fill=value)) +
+geom_tile() +
+geom_text(aes(Var1, Var2, label = value),
+          color = "white", size = 3)
+```
+
+## Decision Tree
+
+#### Training and evaluating on the same training dataset (not a good method for generalizing) , better to do cross validation
+
+```{r}
+# Vanilla decision tree with a fixed value of complexity parameter of 0.01 and minsplit of 20
+reg_tree<-rpart(Crime~.,data=crime_data,method="anova",control=c(cp=0.01,minsplit=20))
+rmse<-sqrt(sum((crime_data$Crime - predict(reg_tree,type=c("vector")))**2)/nrow(crime_data))
+rmse
+```
+
+```{r}
+rpart.plot(reg_tree)
+```
+
+[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
+
+-   Observations & Points for further exploration:
+
+-   The above decision tree uses 3 variables NW , Pop, LF and Po1 to create a decision tree. (2 of them Pop and Po1 have a linear correlation with dependent variable i.e. Crime)
+
+-   One food for thought is that this is a regression tree and we are imputing the mean values of all the points that fall in a leaf or terminal node , so incase if we have an outlier , then those will drive the value of mean and hence cause higher RMSE or SSE.
+So , it would be good to see the effect of removing or capping outliers and then performing decision tree.
+
+-   The smaller the value of complex parameter , the model is expected to have high variance as a split would be performed if we get such a marginal increase in entropy .
+Therefore , to prevent our model from overfitting amd having high variance , it is better if we not choose too small values of cp i.e.
+complexity parameter.
+
+-   The default value of max depth parameter here is 30 but given that we have only 47 nodes in our dataset , it dfoes not make much sense to have tree models greater than or equal to 3 or 4 levels of depth
+
+-   minsplit hyperparameter : The default value of this parameter is 20 i.e.
+a split would be performed if we have atleast 20 datapoints in a node , although it is generally good to have higher values of "minsplit" parameter to avoid overfitting but since in this problem statement , the total \# of datapoints is only 47 , so this is one parameter where we can tweak in a bit .
+
+-   minbucket hyperparameter : Similarly , this hyperparameter means the number of datpoints in a terminal node and thus here also , it makes sense to have a smaller value for this hyperparameter as we do not have large dataset for training our model.
+
+#### Perform hyperparameter tuning minsplit and cp(complexity hyperparameter)
+
+```{r}
+folds <- sample(rep(1:5, length.out = nrow(crime_data)), size = nrow(crime_data), replace = F)
+pred_table<-data.frame(min_split=numeric(0),cp=double(0),rmse=double(0))
+for(j in seq(0.001,0.02,0.001)){
+for(i in seq(5,25,2)){
+  CV_mse <- lapply(1:5, function(x){ #5 corresponds to the number of folds defined earlier
+  model <- rpart(Crime ~ ., data = crime_data[folds != x,],method="anova",control=c(minsplit=i,cp=j))
+  preds <- predict(model,  crime_data[folds == x,], type="vector")
+  sum_sse<-sum((crime_data[folds==x,"Crime"]-preds)**2)
+  rmse<-sqrt(sum_sse/nrow(crime_data[folds==x,]))
+  #total_rmse<-total_rmse+rmse
+  return(rmse)
+   })
+pred_table[nrow(pred_table)+1,]=c(i,j,mean(unlist(CV_mse)))
+}
+}
+```
+
+# Identifying the optimal value of the hyperparameters cp and min_split
+
+```{r}
+options(scipen=999)
+pred_table[which.min(pred_table$rmse),c("min_split","cp")]
+```
+
+-   From the above output , we get the the optimal value of cp as 0.001 and min_split as 15
+
+```{r}
+reg_cv_tree<-rpart(Crime~.,data=crime_data,method="anova",control=c(cp=0.001,minsplit=15))
+rpart.plot(reg_cv_tree)
+```
+
+-   [**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
+-   Note: The above set of optimal parameters reproduces the same decision tree that we had done initially with fixed values of these 2 parameters.This can be because of less \# of datapoints.
+We can try reducing the number of minsplits which will lead to a higher variance model and we can try iterating to get the optimal maxDepth hyperparameter.
+
+```{r}
+pred_table<-data.frame(min_split=numeric(0),maxDepth=double(0),rmse=double(0))
+for(j in seq(2,15,1)){
+for(i in seq(5,25,2)){
+  CV_mse <- lapply(1:5, function(x){ #5 corresponds to the number of folds defined earlier
+  model <- rpart(Crime ~ ., data = crime_data[folds != x,],method="anova",control=c(minsplit=i,maxdepth=j))
+  preds <- predict(model,  crime_data[folds == x,], type="vector")
+  sum_sse<-sum((crime_data[folds==x,"Crime"]-preds)**2)
+  rmse<-sqrt(sum_sse/nrow(crime_data[folds==x,]))
+  #total_rmse<-total_rmse+rmse
+  return(rmse)
+   })
+pred_table[nrow(pred_table)+1,]=c(i,j,mean(unlist(CV_mse)))
+}
+}  
+#rpart.plot(rpart(Crime ~ ., data = crime_data,method="anova",minsplit=10,cp=0.01))
+#plot(pred_table)
+```
+
+-   [**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
+-   Identifying the optimal value of the hyperparameter maxDepth (which comes out to be around 2 with min_split within range of 5-15 (rmse is same for different values of min_split))
+
+```{r}
+options(scipen=999)
+pred_table[which.min(pred_table$rmse),c("maxDepth")]
+```
+
+```{r}
+reg_tree_final<-rpart(Crime~.,data=crime_data,method="anova",control=c(cp=0.001,minsplit=15,maxdepth=2))
+rmse_final<-sqrt(sum((crime_data$Crime - predict(reg_tree_final,type=c("vector")))**2)/nrow(crime_data))
+rmse_final
+```
+
+**Conclusion : The RMSE value for the decision tree model using the optimal values of the hyperparameter is `r round(rmse_final,2)`**
+
+## Random Forest
+
+```{r}
+crime_df <- crime_data
 crime_df$So <- factor(crime_df$So)
@@
 
+[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
+
 From variable importance we can see that the some variables like, So, U1 and U2 are not very helpful.
@@
 In the CV stage below, we will test the models without these variables.
+
+Variable Importance here is given basis two metrics, drop in MSE when the variable is not included in feature set, and second the Information gain (or purity) of a node, if the variable is used to split it.
 
@@
 
-We have created uniform CV folds here
+[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
+
+We have created uniform CV folds here.
 
@@
 
-#### Impact of ntree
-
 ```{r}
@@
 ```
+
+[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
 
@@
 ```
-
-As
 
@@
 
+[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
+
 From the variable importance plot we see, that the following variable feature very highly in Importance:
@@
 
+[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
+
 Note: while the above 3 variables are also significant features here is the variable importance of variable "Pop".

### 2022-09-28 hw5-1/submission.Rmd
Submission.Rmd

@@
 -   Anshit Verma \| [averma373\@gatech.edu](mailto:averma373@gatech.edu){.email}
+
+**Q10.1**
 
@@
     rmse_cv3 = mean(unlist(rmse_cv_df[,3]))
-    print (paste("For tree:",ntree_i,"mtry:",mtry_i))
+    #print (paste("For tree:",ntree_i,"mtry:",mtry_i))
     rmse_list1 = c(rmse_list1, rmse_cv1)
@@
 Another point to note here is that after Hyperparameter tuning, Decision tree is giving lower RMSE, and is most likely over-fitting the data-set, in comparison to Random Forest, and if we had a test data, RF model would most likely perform better, because of better generalization.
+
+**Q10.2**

### 2022-09-28 hw5-1/anshit/Homework.Rmd

---
title: "Homework-5"
date: "26 September 2022"
---

## Question 10.3

```{r, include=FALSE}
library(caret)
library(ROCR)
```

```{r}
german_data = read.table("../germancredit.txt")
german_data
```

```{r}
cols = colnames(german_data)
for(i in seq(1, length(cols))){
  if(class(german_data[, cols[i]]) == "character"){
    german_data[, cols[i]]= as.factor(german_data[, cols[i]])
  }
}
```

### Exploratory Data Analysis

```{r}
summary(german_data)
```

We can see that the data is uniformly distributed, both for the numerical features as well as categorical features.

```{r}
par(mfrow=(c(1,2)))
plot(german_data$V5)
plot(german_data$V2)
```

2 features has suspiciously large max values (comparing to the 3rd quartile).

```{r}
# (1 = Good, 2 = Bad) -> (0 = Good, 1 = Bad)
german_data['V21'] = lapply(german_data['V21'], function(x){
  x-1
})
german_data[,"V21"] = as.factor(german_data[,"V21"])
german_data
```

```{r}
split_data = function(dataset){
  ## Sampling train and test data

  ## 75% of the sample size
  smp_size <- floor(0.75 * nrow(dataset))

  ## set the seed to make your partition reproducible
  set.seed(123)
  train_ind <- sample(seq_len(nrow(dataset)), size = smp_size)

  train <- dataset[train_ind, ]
  test <- dataset[-train_ind, ]
  return(list("train"=train, "test"=test))
}
```

### Splitting the data into train and test

```{r}
df_list = split_data(german_data)
train = df_list$train
test = df_list$test
summary(test$V21)
```

### Training vanilla model with all features

```{r}
model <- glm(V21 ~.,family=binomial(link='logit'),data=train)
model
```

```{r}
summary(model)
```

```{r}
get_roc_plot_and_auc = function(model, test){
  p <- predict(model, newdata=test[, 1:(length(test)-1)], type="response")
  pr <- prediction(p, test$V21)
  prf <- performance(pr, measure = "tpr", x.measure = "fpr")
  auc <- performance(pr, measure = "auc")
  auc <- auc@y.values[[1]]
  return (list("prf"=prf, "auc"=auc, "prediction"=p))
}
```

### Plotting the ROC Curve and calculating AUC

```{r}
roc = get_roc_plot_and_auc(model, test)
plot(roc$prf)
roc$auc
```

As we can see that AUC value is 0.7846446 that is close to 1. Hence there is our model is above average in being able to separate between classes.

```{r}
predicted = ifelse(roc$prediction > 0.5,1,0)
cor = confusionMatrix(data=as.factor(predicted), reference=as.factor(test$V21), positive="1")
cor
```

```{r}
cor$byClass
```

With and initial (default) threshold of 0.5, we can see that our model achieves an accuracy of 75.6% with a recall of 54.3% and precision of 64.7%. We can try dropping some statistically insignificant features to better improve the meterics of our model.

### Dropping Features

```{r}
# We can drop statistically insignificant features
drop.features = german_data
drops <- c("V7","V17", "V11", "V13", "V16", "V12")
drop.features = drop.features[ , !(names(drop.features) %in% drops)]
```

```{r}
df.list.1 = split_data(drop.features)
train.1 = df.list.1$train
test.1 = df.list.1$test
```

```{r}
model.1 <- glm(V21 ~.,family=binomial(link='logit'),data=train.1)
model.1
```

```{r}
summary(model.1)
```

```{r}
roc.1 = get_roc_plot_and_auc(model.1, test.1)
plot(roc.1$prf)
roc.1$auc
```

We can observe that dropping insignificant features help us achieve a higher AUC value. Hence this increases our model's ability to seperate between the classes.

```{r}
predicted.1 = ifelse(roc.1$prediction > 0.5,1,0)
cor.1 = confusionMatrix(data=as.factor(predicted.1), reference=as.factor(test.1$V21), positive="1")
cor.1
```

```{r}
cor.1$byClass
```

Again we can observe that with the initial (default) value of threshold - 0.5, we have increased the accuracy of our model. While the recall remains unchanged, we were able to increase the precision.

### Determining Threshold by F-Beta Value

```{r}
get_recall_and_precision = function(cor){
  cor_table = data.frame(cor$table)
  tn = cor_table[cor_table$Prediction==0 & cor_table$Reference==0, "Freq"]
  fn = cor_table[cor_table$Prediction==0 & cor_table$Reference==1, "Freq"]
  tp = cor_table[cor_table$Prediction==1 & cor_table$Reference==1, "Freq"]
  fp = cor_table[cor_table$Prediction==1 & cor_table$Reference==0, "Freq"]
  recall = tp / (tp + fn)
  precision = tp / (tp + fp)
  return (c("recall"=recall, "precision"=precision))
}
```

We can use the F-Beta value for determining the best value of threshold. As we are given that a classifying a "bad" customer as "good" is 5 times worse than classifying "good" customer as "bad". We need to determine the best value of threshold that will help us increase recall and accordingly lower the value of false negatives (positive class is 1 i.e. "bad").

```{r}
rp.df = data.frame("threshold"=double(), "recall"=double(), "precision"=double(), f_beta=double())
for(i in seq(0, 100, 1)){
  beta=2
  predicted.2 = ifelse(roc.1$prediction > i/100,1,0)
  cor.2 = confusionMatrix(data=as.factor(predicted.2), reference=as.factor(test.1$V21), positive = "1")
  rp = get_recall_and_precision(cor.2)
  f_beta = ((1 + beta^2) * rp[2] * rp[1]) / (beta^2 * rp[2] + rp[1])
  rp.df[nrow(rp.df)+1,] = c("threshold"=i/100, rp, f_beta)
}
```

We are using F2 specifically which helps us weight recall more than precision. Since we want to maximize recall and minimize false negative due to higher cost associated with it, F2 will help us determine the best threshold value. Since F2 can range from 0 to 2, the higher the value of F2 - the better the threshold to minimize false negative.

```{r}
rp.df
```

```{r}
# Max value of F Beta Score
rp.df[which.max(rp.df$f_beta),]
```
***Plotting the best value of threshold for maximum F2 value***

```{r}
plot(x=rp.df$threshold, y=rp.df$f_beta, type="o", xlab="Probability Threshold", ylab="F Beta Score", main="F2 v/s Threshold")
points(x=rp.df$threshold[which.max(rp.df$f_beta)], y=rp.df$f_beta[which.max(rp.df$f_beta)], pch = 18, col = "red", type = "b", lty = 2)
```

We can see from the above plot that best value that minimizes false negatives is 0.11.

### Determining Threshold by Calculating Error Rate

```{r}
get_cm_value = function(cor){
  cor_table = data.frame(cor.2$table)
  tn = cor_table[cor_table$Prediction==0 & cor_table$Reference==0, "Freq"]
  fn = cor_table[cor_table$Prediction==0 & cor_table$Reference==1, "Freq"]
  tp = cor_table[cor_table$Prediction==1 & cor_table$Reference==1, "Freq"]
  fp = cor_table[cor_table$Prediction==1 & cor_table$Reference==0, "Freq"]
  return(list(tn=tn,fn=fn,tp=tp,fp=fp))
}
```

Another way determine the threshold is by calculating the error value by weighing the false negatives more than false positive. As we know that false negatives are 5 time more costly than false positive, we can use that information while calculating our error value.
We need to choose the value of threshold which minimizes the error.

```{r}
## Minimizing the error
## Since we know FN is 5 times worse than FP, we minimize the error that gives us the best threshold

error.df = data.frame("threshold"=double(), "error"=double())
for(i in seq(0, 100, 1)){
  beta=2
  predicted.2 = ifelse(roc.1$prediction > i/100,1,0)
  cor.2 = confusionMatrix(data=as.factor(predicted.2), reference=as.factor(test.1$V21), positive = "1")
  cm.values = get_cm_value(cor.2)
  total_error = (5*cm.values$fn + cm.values$fp)/nrow(german_data)
  error.df[nrow(error.df)+1,] = c("threshold"=i/100, "error"=total_error)
}
```

```{r}
error.df[which.min(error.df$error),]
```

We can again see that our error value is minimum at the threshold value of 0.11. This is in sync with our value of threshold which we determined via F-beta method. This this value of threshold is optimal.

```{r}
plot(error.df)
points(x=error.df$threshold[which.min(error.df$error)], y=error.df$error[which.min(error.df$error)], pch = 18, col = "red", type = "b", lty = 2)
```

### Final Model

```{r}
final.predicted = ifelse(roc.1$prediction > 0.11,1,0)
final.cor = confusionMatrix(data=as.factor(final.predicted), reference=as.factor(test.1$V21), positive="1")
final.cor
```

```{r}
final.cor$byClass
```

By choosing our threshold to be 0.11, we can achieve a recall value of 0.91. While our precision has gone down, we are not making mistakes with false negatives which are more costly to us.

In our test data, the no. of "bad" classified data points was ~33%. We were able to achieve a threshold value of 0.11 which is 1/3rd of that value.

### Bucketing and New Feature

```{r}
## Creating feature - Credit Amount(V5) / Duration in months(V2)

feature_data.1 = german_data
feature_data.1['V22'] = with(feature_data.1, V5/V2)
feature_data.1
```

```{r}
summary(feature_data.1$V22)
```

```{r}
feature_data.1["Bucket"] = ""
feature_data.1[feature_data.1$V22 <= 89.60, "Bucket"] = "<=89.60"
feature_data.1[feature_data.1$V22 <= 130.33 & feature_data.1$V22 > 89.60, "Bucket"] = ">89.60 & <=130.33"
feature_data.1[feature_data.1$V22 <= 206.18 & feature_data.1$V22 > 130.33, "Bucket"] = ">130.33 & <=206.18"
feature_data.1[feature_data.1$V22 > 206.18, "Bucket"] = ">206.18"
```

```{r}
feature_data.1
```

```{r}
plot(as.factor(x=feature_data.1$Bucket), y=feature_data.1$V21)
```

```{r}
feature_data.1 = feature_data.1[,1:length(feature_data.1)-1]
feature_data.1 = feature_data.1[,c(seq(1,20), 22, 21)]
drops <- c("V5","V2")
feature_data.1[ , !(names(feature_data.1) %in% drops)]
```

```{r}
## Sampling train and test data

## 75% of the sample size
smp_size <- floor(0.75 * nrow(feature_data.1))

## set the seed to make your partition reproducible
set.seed(123)
train_ind <- sample(seq_len(nrow(feature_data.1)), size = smp_size)

feature.1.train <- feature_data.1[train_ind, ]
feature.1.test <- feature_data.1[-train_ind, ]

model.1 <- glm(V21 ~.,family=binomial(link='logit'),data=feature.1.train)
```

```{r}
model.1
```

```{r}
p.1 <- predict(model.1, newdata=feature.1.test[, 1:(length(feature.1.test)-1)], type="response")
pr.1 <- prediction(p.1, feature.1.test$V21)
prf.1 <- performance(pr.1, measure = "tpr", x.measure = "fpr")
plot(prf.1)
auc.1 <- performance(pr.1, measure = "auc")
auc.1 <- auc.1@y.values[[1]]
auc.1
```

```{r}
predicted.1 = ifelse(p.1 > 0.35,1,0)
confusionMatrix(data=as.factor(predicted.1), reference=as.factor(feature.1.test$V21))
```

### 2022-09-28 hw5-1/submission.Rmd

@@
 
-**Q10.1**
+[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
 
-[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
+## Question 10.1
 
@@
 
-## EDA
+### EDA
 
@@
 
-## Decision Tree
+### Decision Tree
 
@@
 
-[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
+[**#Analysis**]{style="color:red"}
 
@@
 
-# Identifying the optimal value of the hyperparameters cp and min_split
+#### Identifying the optimal value of the hyperparameters cp and min_split
 
@@
 
--   [**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
+[**#Analysis**]{style="color:red"}
+
 -   Note: The above set of optimal parameters reproduces the same decision tree that we had done initially with fixed values of these 2 parameters.This can be because of less \# of datapoints.
@@
 
--   [**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}
+[**#Analysis**]{style="color:red"}
+
 -   Identifying the optimal value of the hyperparameter maxDepth (which comes out to be around 2 with min_split within range of 5-15 (rmse is same for different values of min_split))
@@
 
-## Random Forest
+### Random Forest
 
@@
 
-### Build Vanilla RF model
+#### Build Vanilla RF model
 
@@
 
-**Q10.2**
+## Question 10.2
+
+## Question 10.3
+
+```{r, include=FALSE}
+library(caret)
+library(ROCR)
+```
+
+```{r}
+german_data = read.table("./germancredit.txt")
+german_data
+```
+
+```{r}
+cols = colnames(german_data)
+for(i in seq(1, length(cols))){
+  if(class(german_data[, cols[i]]) == "character"){
+    german_data[, cols[i]]= as.factor(german_data[, cols[i]])
+  }
+}
+```
+
+### Exploratory Data Analysis
+
+```{r}
+summary(german_data)
+```
+
+We can see that the data is uniformly distributed, both for the numerical features as well as categorical features.
+
+```{r}
+par(mfrow=(c(1,2)))
+plot(german_data$V5)
+plot(german_data$V2)
+```
+
+2 features have suspiciously large max values (comparing to the 3rd quartile).
+
+```{r}
+# (1 = Good, 2 = Bad) -> (0 = Good, 1 = Bad)
+german_data['V21'] = lapply(german_data['V21'], function(x){
+  x-1
+})
+german_data[,"V21"] = as.factor(german_data[,"V21"])
+german_data
+```
+
+```{r}
+split_data = function(dataset){
+  ## Sampling train and test data
+
+  ## 75% of the sample size
+  smp_size <- floor(0.75 * nrow(dataset))
+
+  ## set the seed to make your partition reproducible
+  set.seed(123)
+  train_ind <- sample(seq_len(nrow(dataset)), size = smp_size)
+
+  train <- dataset[train_ind, ]
+  test <- dataset[-train_ind, ]
+  return(list("train"=train, "test"=test))
+}
+```
+
+### Splitting the data into train and test
+
+```{r}
+df_list = split_data(german_data)
+train = df_list$train
+test = df_list$test
+summary(test$V21)
+```
+
+### Training vanilla model with all features
+
+```{r}
+model <- glm(V21 ~.,family=binomial(link='logit'),data=train)
+model
+```
+
+```{r}
+summary(model)
+```
+
+```{r}
+get_roc_plot_and_auc = function(model, test){
+  p <- predict(model, newdata=test[, 1:(length(test)-1)], type="response")
+  pr <- prediction(p, test$V21)
+  prf <- performance(pr, measure = "tpr", x.measure = "fpr")
+  auc <- performance(pr, measure = "auc")
+  auc <- auc@y.values[[1]]
+  return (list("prf"=prf, "auc"=auc, "prediction"=p))
+}
+```
+
+### Plotting the ROC Curve and calculating AUC
+
+```{r}
+roc = get_roc_plot_and_auc(model, test)
+plot(roc$prf)
+roc$auc
+```
+
+[**#Analysis**]{style="color:red"}
+
+As we can see that AUC value is 0.7846446 that is close to 1. Hence our model is above average in being able to separate between classes.
+
+```{r}
+predicted = ifelse(roc$prediction > 0.5,1,0)
+cor = confusionMatrix(data=as.factor(predicted), reference=as.factor(test$V21), positive="1")
+cor
+```
+
+```{r}
+cor$byClass
+```
+
+[**#Analysis**]{style="color:red"}
+
+With and initial (default) threshold of 0.5, we can see that our model achieves an accuracy of 75.6% with a recall of 54.3% and precision of 64.7%.
+We can try dropping some statistically insignificant features to better improve the metrics of our model.
+
+### Dropping Features
+
+```{r}
+# We can drop statistically insignificant features
+drop.features = german_data
+drops <- c("V7","V17", "V11", "V13", "V16", "V12")
+drop.features = drop.features[ , !(names(drop.features) %in% drops)]
+```
+
+```{r}
+df.list.1 = split_data(drop.features)
+train.1 = df.list.1$train
+test.1 = df.list.1$test
+```
+
+```{r}
+model.1 <- glm(V21 ~.,family=binomial(link='logit'),data=train.1)
+model.1
+```
+
+```{r}
+summary(model.1)
+```
+
+```{r}
+roc.1 = get_roc_plot_and_auc(model.1, test.1)
+plot(roc.1$prf)
+roc.1$auc
+```
+
+[**#Analysis**]{style="color:red"}
+
+We can observe that dropping statistically insignificant features help us achieve a higher AUC value. Hence this increases our model's ability to separate between the classes.
+
+```{r}
+predicted.1 = ifelse(roc.1$prediction > 0.5,1,0)
+cor.1 = confusionMatrix(data=as.factor(predicted.1), reference=as.factor(test.1$V21), positive="1")
+cor.1
+```
+
+```{r}
+cor.1$byClass
+```
+
+[**#Analysis**]{style="color:red"}
+
+Again we can observe that with the initial (default) value of threshold - 0.5, we have increased the accuracy of our model. While the recall remains unchanged, we were able to increase the precision.
+
+### Determining Threshold by F-Beta Value
+
+```{r}
+get_recall_and_precision = function(cor){
+  cor_table = data.frame(cor$table)
+  tn = cor_table[cor_table$Prediction==0 & cor_table$Reference==0, "Freq"]
+  fn = cor_table[cor_table$Prediction==0 & cor_table$Reference==1, "Freq"]
+  tp = cor_table[cor_table$Prediction==1 & cor_table$Reference==1, "Freq"]
+  fp = cor_table[cor_table$Prediction==1 & cor_table$Reference==0, "Freq"]
+  recall = tp / (tp + fn)
+  precision = tp / (tp + fp)
+  return (c("recall"=recall, "precision"=precision))
+}
+```
+
+[**#Analysis**]{style="color:red"}
+
+We can use the F-Beta value for determining the best value of threshold.
+As we are given that a classifying a "bad" customer as "good" is 5 times worse than classifying "good" customer as "bad".
+We need to determine the best value of threshold that will help us increase recall and accordingly lower the value of false negatives (positive class is 1 i.e.
+"bad").
+
+```{r}
+rp.df = data.frame("threshold"=double(), "recall"=double(), "precision"=double(), f_beta=double())
+for(i in seq(0, 100, 1)){
+  beta=2
+  predicted.2 = ifelse(roc.1$prediction > i/100,1,0)
+  cor.2 = confusionMatrix(data=as.factor(predicted.2), reference=as.factor(test.1$V21), positive = "1")
+  rp = get_recall_and_precision(cor.2)
+  f_beta = ((1 + beta^2) * rp[2] * rp[1]) / (beta^2 * rp[2] + rp[1])
+  rp.df[nrow(rp.df)+1,] = c("threshold"=i/100, rp, f_beta)
+}
+```
+
+[**#Analysis**]{style="color:red"}
+
+We are using F2 specifically which helps us weight recall more than precision.
+Since we want to maximize recall and minimize false negative due to higher cost associated with it, F2 will help us determine the best threshold value.
+Since F2 can range from 0 to 2, the higher the value of F2 - the better the threshold to minimize false negative.
+
+```{r}
+rp.df
+```
+
+```{r}
+# Max value of F Beta Score
+rp.df[which.max(rp.df$f_beta),]
+```
+
+***Plotting the best value of threshold for maximum F2 value***
+
+```{r}
+plot(x=rp.df$threshold, y=rp.df$f_beta, type="o", xlab="Probability Threshold", ylab="F Beta Score", main="F2 v/s Threshold")
+points(x=rp.df$threshold[which.max(rp.df$f_beta)], y=rp.df$f_beta[which.max(rp.df$f_beta)], pch = 18, col = "red", type = "b", lty = 2)
+```
+
+[**#Analysis**]{style="color:red"}
+
+We can see from the above plot that best value that minimizes false negatives is 0.11.
+
+### Determining Threshold by Calculating Error Rate
+
+```{r}
+get_cm_value = function(cor){
+  cor_table = data.frame(cor.2$table)
+  tn = cor_table[cor_table$Prediction==0 & cor_table$Reference==0, "Freq"]
+  fn = cor_table[cor_table$Prediction==0 & cor_table$Reference==1, "Freq"]
+  tp = cor_table[cor_table$Prediction==1 & cor_table$Reference==1, "Freq"]
+  fp = cor_table[cor_table$Prediction==1 & cor_table$Reference==0, "Freq"]
+  return(list(tn=tn,fn=fn,tp=tp,fp=fp))
+}
+```
+
+[**#Analysis**]{style="color:red"}
+
+Another way determine the threshold is by calculating the error value by weighing the false negatives more than false positive.
+As we know that false negatives are 5 time more costly than false positive, we can use that information while calculating our error value.
+We need to choose the value of threshold which minimizes the error.
+
+```{r}
+## Minimizing the error
+## Since we know FN is 5 times worse than FP, we minimize the error that gives us the best threshold
+
+error.df = data.frame("threshold"=double(), "error"=double())
+for(i in seq(0, 100, 1)){
+  beta=2
+  predicted.2 = ifelse(roc.1$prediction > i/100,1,0)
+  cor.2 = confusionMatrix(data=as.factor(predicted.2), reference=as.factor(test.1$V21), positive = "1")
+  cm.values = get_cm_value(cor.2)
+  total_error = (5*cm.values$fn + cm.values$fp)/nrow(german_data)
+  error.df[nrow(error.df)+1,] = c("threshold"=i/100, "error"=total_error)
+}
+```
+
+```{r}
+error.df[which.min(error.df$error),]
+```
+
+[**#Analysis**]{style="color:red"}
+
+We can again see that our error value is minimum at the threshold value of 0.11.
+This is in sync with our value of threshold which we determined via F-beta method.
+This this value of threshold is optimal.
+
+```{r}
+plot(error.df)
+points(x=error.df$threshold[which.min(error.df$error)], y=error.df$error[which.min(error.df$error)], pch = 18, col = "red", type = "b", lty = 2)
+```
+
+### Final Model
+
+```{r}
+final.predicted = ifelse(roc.1$prediction > 0.11,1,0)
+final.cor = confusionMatrix(data=as.factor(final.predicted), reference=as.factor(test.1$V21), positive="1")
+final.cor
+```
+
+```{r}
+final.cor$byClass
+```
+
+[**#Analysis**]{style="color:red"}
+
+By choosing our threshold to be 0.11, we can achieve a recall value of 0.91. While our precision has gone down, we are not making mistakes with false negatives which are more costly to us.
+
+In our test data, the no. of "bad" classified data points was ~33%. We were able to achieve a threshold value of 0.11 which is 1/3rd of that value.

### 2022-09-28 hw5-1/submission.Rmd
submission v2

@@
 2.  number of trees = 40
-3.  mtry = 4
+3.  mtry = 3
 
@@
 ```{r}
-rf_final_model <- randomForest(Crime ~ . -U1 -M.F -U2, data = crime_df, ntree = 40, mtry=4, importance = TRUE)
+rf_final_model <- randomForest(Crime ~ . -U1 -M.F -U2, data = crime_df, ntree = 40, mtry=3, importance = TRUE)
 rmse_func(true=crime_df$Crime,predicted = rf_final_model$predicted)
@@
 3.  Prob
+
+If we want we can still test if drop the lowest variables: Time and Pop is helpful.
 
@@
 
-We can observe that dropping statistically insignificant features help us achieve a higher AUC value. Hence this increases our model's ability to separate between the classes.
+We can observe that dropping statistically insignificant features help us achieve a higher AUC value, which translates to a better risk discrimination.
+In other words, this increases our model's ability to separate between the classes.
 
@@
 As we are given that a classifying a "bad" customer as "good" is 5 times worse than classifying "good" customer as "bad".
-We need to determine the best value of threshold that will help us increase recall and accordingly lower the value of false negatives (positive class is 1 i.e.
+We need to determine the best value of threshold that will help us increase recall (positive class is 1 i.e.
 "bad").
@@
 ```{r}
+ggplot(rp.df, aes(threshold)) + 
+  geom_line(aes(y = precision, colour = "precision")) + 
+  geom_line(aes(y = recall, colour = "recall"))
+```
+
+[**#Analysis**]{style="color:red"}
+
+As expected, decreasing threshold, makes it easier to classify data points as bad, and so our recall value increases.
+On the other hand, because on longer threshold, we might no longer be tagging the actual bads, our model precision decreases.
+
+Also the Precision and Recall curver intersect somewhere between 0.3 to 0.4, which is near the bad rate in our dataset.
+
+```{r}
 # Max value of F Beta Score
@@
 ### Determining Threshold by Calculating Error Rate
+
+$$
+Error Rate^* = \frac{5.FP}{TP+FP+FN+TN} + \frac{FN}{TP+FP+FN+TN}
+$$
+
+\*Note we are deviating here from conventional deviation of error rate, in order to better apply to our use case.
 
@@
 
-By choosing our threshold to be 0.11, we can achieve a recall value of 0.91. While our precision has gone down, we are not making mistakes with false negatives which are more costly to us.
+By choosing our threshold to be 0.11, we can achieve a recall value of 0.91, which implies if there are 100 "bad" points we are able to capture 91 of those.
 
-In our test data, the no. of "bad" classified data points was ~33%. We were able to achieve a threshold value of 0.11 which is 1/3rd of that value.
+It is also interesting to note here, that the we have \~33% "bad" data points. We
+
+In our test data, the no. of "bad" classified data points was \~33%. We were able to achieve a threshold value of 0.11 which is 1/3rd of that value.

### 2022-09-28 hw5-1/submission.Rmd
final push

@@
 
-#### Training and evaluating on the same training dataset (not a good method for generalizing) , better to do cross validation
+#### Training and evaluating on the same training dataset (not a good method for generalizing) , better to do cross validation to get optimal hyperparameters.
+Splitting into training and testing set does not make sense here as we have only 47 datapoints.
 
@@
 -   One food for thought is that this is a regression tree and we are imputing the mean values of all the points that fall in a leaf or terminal node , so incase if we have an outlier , then those will drive the value of mean and hence cause higher RMSE or SSE.
-So , it would be good to see the effect of removing or capping outliers and then performing decision tree.
 
@@
 
--   The default value of max depth parameter here is 30 but given that we have only 47 nodes in our dataset , it dfoes not make much sense to have tree models greater than or equal to 3 or 4 levels of depth
+-   The default value of max depth parameter here is 30 but given that we have only 47 nodes in our dataset , it does not make much sense to have tree models greater than or equal to 3 or 4 levels of depth
 
@@
 
--   Identifying the optimal value of the hyperparameter maxDepth (which comes out to be around 2 with min_split within range of 5-15 (rmse is same for different values of min_split))
+-   Identifying the optimal value of the hyperparameter maxDepth (which comes out to be 3 with min_split within range of 5-15 (rmse is same for different values of min_split))
 
@@
 ## Question 10.2
+
+### Logistic Regression finds many use cases in the healthcare analytics domain , one of them being predicting the probability of readmission within 30 days of being discharged .
+The above problem depends on lot of factors , the primary ones being discussed below :
+
++ Dicharge Disposition : The discharge disposition/location code is an important factor for determining the probability of readmission .
+People with a relatively high risk score and requiring post acute care when  discharged directly to home are more prone to getting readmitted.
+
++ Length of Stay : The period/length of stay also has a strong predictive power on the probability of readmission.
+Both low and high values are correlated with higher readmission rate.Low length of stay can lead to readmission due to patient not being treated properly without proper diagnosis and people with higher length of stay get readmitted due to Hospital acquired diseases.
+
++ Risk Score : The overall risk score of a patient based on past medical history and demographics is also a good prediction variable for the probability of readmission.
+
++ Change in medication/procedure during hospitalization : This is also a significant feature related with higher readmission rates.
+People who have seen greater changes in procedure/prescribed medicine are more prone to getting readmitted.
+
++ Number of ED/IP visits within the rolling 1 year : This utilization metric is also a good predictor as it encapsulates the overall profile of a patient .
+Patients with higher ED/IP visits are more prone to getting readmitted within 30 days of being discharged.
 

### 2022-10-10 hw5-1/submission.Rmd
hw 7

@@
 
-If we want we can still test if drop the lowest variables: Time and Pop is helpful.
+If we want we can still test if dropping the lowest variables, Time and Pop is helpful. This shall further simplify our model, which is especially helpful with small amount of data.
 
@@
 
-+ Dicharge Disposition : The discharge disposition/location code is an important factor for determining the probability of readmission .
+-   Dicharge Disposition : The discharge disposition/location code is an important factor for determining the probability of readmission .
 People with a relatively high risk score and requiring post acute care when discharged directly to home are more prone to getting readmitted.
 
-+ Length of Stay : The period/length of stay also has a strong predictive power on the probability of readmission.
+-   Length of Stay : The period/length of stay also has a strong predictive power on the probability of readmission.
 Both low and high values are correlated with higher readmission rate.Low length of stay can lead to readmission due to patient not being treated properly without proper diagnosis and people with higher length of stay get readmitted due to Hospital acquired diseases.
 
-+ Risk Score : The overall risk score of a patient based on past medical history and demographics is also a good prediction variable for the probability of readmission.
+-   Risk Score : The overall risk score of a patient based on past medical history and demographics is also a good prediction variable for the probability of readmission.
 
-+ Change in medication/procedure during hospitalization : This is also a significant feature related with higher readmission rates.
+-   Change in medication/procedure during hospitalization : This is also a significant feature related with higher readmission rates.
 People who have seen greater changes in procedure/prescribed medicine are more prone to getting readmitted.
 
-+ Number of ED/IP visits within the rolling 1 year : This utilization metric is also a good predictor as it encapsulates the overall profile of a patient .
+-   Number of ED/IP visits within the rolling 1 year : This utilization metric is also a good predictor as it encapsulates the overall profile of a patient .
 Patients with higher ED/IP visits are more prone to getting readmitted within 30 days of being discharged.

### 2022-10-11 hw7/ashish/lasso.Rmd
hw 7: lasso

```{bash}
ls ../
```

```{r, warning=FALSE}
library(glmnet)
```

```{r}
crime_df = read.table("../uscrime.txt", sep="\t", header= TRUE)
dim(crime_df)
```

```{r}
names(crime_df)
```

```{r}
x = as.matrix(crime_df[,names(crime_df) != "Crime"])
y = as.matrix(crime_df[,"Crime"])
dim(x)
dim(y)
```

```{r}
fit <- glmnet(x, y, alpha = 1, nlambda = 20, standardize = TRUE)
```

```{r}
print(fit)
```

```{r}
plot(fit, xvar = "lambda", label = TRUE)
```

```{r}
cvfit <- cv.glmnet(x = x, y = y, alpha = 1, nlambda = 30, standardize = TRUE, type.measure = "mse", nfolds = 5)
```

Because we have only 47 points in the data, we don't want fit too many models during the CV stage

```{r}
plot(cvfit)
```

Above plot gives CV error against values of lambda. With increasing Lambda the weight of regularization factor increases, and therefore the model is encouraged, to select lower number of predictors.

```{r}
print (cvfit)
```

Two special values along the 𝜆λ sequence are indicated by the vertical dotted lines. `lambda.min` is the value of λ that gives minimum mean cross-validated error, while `lambda.1se` is the value of λ that gives the most regularized model such that the cross-validated error is within one standard error of the minimum.

```{r}
coef(cvfit, s = "lambda.min")
```

```{r}
coef(cvfit, s = "lambda.1se")
```

```{r}
cvfit$lambda.1se
```

```{r}
fit_1se <- glmnet(x, y, alpha = 1, lambda = cvfit$lambda.1se, standardize = TRUE)
coef(fit_1se)

fit_min <- glmnet(x, y, alpha = 1, lambda = cvfit$lambda.min, standardize = TRUE)
coef(fit_min)
```

```{r}
predict_min = predict(fit_min, newx = x, s = "lambda.min")
predict_1se = predict(fit_1se, newx = x, s = "lambda.1se")
```

```{r}
rmse_func = function(true,predicted){
  (mean((true-predicted)^2))**0.5
}
```

### RMSE for

```{r}
rmse_func(true=y,predicted = predict_min)
```

```{r}
rmse_func(true=y,predicted = predict_1se)
```

### 2022-10-12 hw7/abhinav/stepwise.Rmd
STEPWSIE REGRESSION

---
title: "ISYE 6501"
output:
  html_document:
    df_print: paged
  pdf_document: default
date: "2022-10-11"
---

# Submission HW4 \| Fall 22

-   Ashish Dhiman \| [ashish.dhiman\@gatech.edu](mailto:ashish.dhiman@gatech.edu){.email}

-   Abhinav Arun \| [aarun60\@gatech.edu](mailto:aarun60gatech.edu){.email}

-   Anshit Verma \| [averma373\@gatech.edu](mailto:averma373@gatech.edu){.email}

[**\*Analysis Notes are marked with Red header: #Analysis**]{style="color:red"}

## Question 11.1

```{r, warning=FALSE, message=FALSE}
library(ggplot2)
library(qqplotr)
library(cowplot)
library(reshape2)
library(grid)
library(MASS)
#install.packages("car")
library(car)
```

```{r}
crime_data<-read.table(file = "./uscrime.txt", sep = "\t",header=TRUE)
head(crime_data)
dim(crime_data)
```

```{r}
# Fitting the full model
# The full model has 15 independent variables 
full_model <- lm(Crime ~., data = crime_data)
summary(full_model)
```

```{r}
### Data Distribution and Pairwise Corelation
my_plots <- lapply(names(crime_data), function(var_x){
  p <- 
    ggplot(crime_data) +
    aes_string(var_x)
  if(var_x %in% list("So")) {
    p <- p + geom_bar()
  } else {
    p <- p + geom_density()
  } 
})
plot_grid(plotlist = my_plots)
### Correlation
cormat <- round(cor(crime_data),2)
cormat[upper.tri(cormat)] <- NA
melted_cormat <- melt(cormat)
# plotting the correlation heatmap
ggplot(data = melted_cormat, aes(x=Var1, y=Var2,
                                   fill=value)) +
geom_tile() +
geom_text(aes(Var1, Var2, label = value),
          color = "white", size = 3)
```

```{r}
#The "maximal" model is a linear regression model which assumes independent model errors and includes only main effects for the predictor variables.

#The "maximal" model is used as input for the stepAIC() function

# Stepwise regression model to choose the best model by AIC in a stepwise algorithm (shown with iterations)
step_model <- stepAIC(full_model, direction = "both")
summary(step_model)

```

```{r}
#Final model at the end of stepwise regression model
final_model<-stepAIC(full_model, direction = "both",trace=FALSE)
summary(final_model)

# Computing the Variance Inflation factor to see if multicolinearity issue exists in the final model
vif(final_model)

```

# Observations/Inferences :

1. The initial full model has 15 independent predictor variables.  Whereas , our final model output of the Stepwise Regression Model has 8 important predictor variables that minimizes the AIC value . These variables are evaluated on using the combination of forward selection and backward elimination methods iteratively. The lower the AIC value , the better the model as AIC value estimates the relative amount of information lost by a given model in an effort to estimate the actual model (Info Theory).

2. Another point to note is that the Multiple R-squared of the final model is 0.789 which is lower than the initial model which has a value of 0.8031 but this is the pitfall of R squared value i.e. It will always increases with the increase in the # of variables without any penalty/adjustment for model complexity/overfitting unlike AIC or Adjusted R-squared.

3. If we observe Adjusted R squared value , it is higher for our final model with 8 independent predictor variables , thus the final model is validated to be better than the initial model on the basis of Adjusted R squared value too.

![](adjs-R.png)

Based on the above formula , we can infer that the Adjusted R squared value decreases with the increase in # of independent variables(k) until and unless if there is a significant increase in the R squared value on adding a new variable. Thus both AIC and Adjusted R squared provide a technique for selecting simpler model which prevents overfitting.

4. Also , based on the correlation plot it could be seen that variables Wealth and Income Inequality are 2 highly correlated variables (-ve correlation), hence we should pick only 1 variable out of the 2. Same goes for Po1 and Po2. So , if we see our final model , it has just Ineq (Income Inequality) and Po1 (one out of each pair of correlated variables ) and thus our stepwise regression model handles it implicitly.

5. Since , the Variance Inflation factor for all the predictor variables in the final output is less than 5 , we can infer that the problem of multicollinearity also does not reside in our final model.

### 2022-10-12 hw7/anshit/Submission.Rmd

---
title: "ISYE-6501 HW 7"
author: "Anshit Verma"
date: "11/10/2022"
---

## Question 11.1

## Elastic Net

* Elastic net is a combination of both Lasso and Ridge regression. It uses both L1 and L2 penalty to restrict the regression solution.
* It involves two hyperparameters - Lambda and Alpha.
* Alpha is used to give direction between Ridge and Lasso. Alpha = 0 is Ridge and Alpha = 1 is Lasso.
* Lambda is amount of penalty in the regression.

```{r}
library(glmnet)
```

```{r}
data = read.table("../uscrime.txt", sep="\t", header=TRUE)
head(data)
```

```{r}
x = as.matrix(data[, 1:ncol(data)-1])
y = data[,"Crime"]
```

```{r}
x = scale(x)
```

### Train Test Split

```{r}
n=nrow(data)
train_rows <- sample(1:n, .8*n, replace = F)

x.train <- x[train_rows,]
y.train <- y[train_rows]

x.test <- x[-train_rows,]
y.test <- y[-train_rows]
```

* Splitting the data to get a value of MSE for deciding a value of alpha.

### Estimating Lambda and Alpha - hyperparameters

```{r}
models <- list()
for (i in 0:20) {
  name <- paste0("alpha", i/20)

  models[[name]] <-
    cv.glmnet(as.matrix(x.train), as.matrix(y.train), alpha=i/20, family="gaussian", nfold = 5)
}
```

* We are fixing the value of alpha and using cross validation for find the value of lambda which minimizes the value of MSE - our metric for validation.
* We find a minimum value of lambda which minimizes the MSE and another value of lambda one standard deviation apart as it is more practical. We are accepting some performance loss and incorporating more bias in the model.

```{r}
results <- data.frame()
for (i in 0:20) {
  name <- paste0("alpha", i/20)

  ## Use each model to predict 'y' given the Testing dataset
  predicted <- predict(models[[name]], 
                       s=models[[name]]$lambda.1se, newx=x.test)

  ## Calculate the Mean Squared Error...
  mse <- mean((y.test - predicted)^2)

  ## Store the results
  temp <- data.frame(alpha=i/20, mse=mse, name=name)
  results <- rbind(results, temp)
}
```

* After we have selected our lambda, we predict on the test data for getting the value of alpha.
* Alpha is the parameter for deciding between ridge (L2 norm) penalty and lasso (L1 penalty) penalty.

```{r}
print(results)
```

```{r}
plot(results$alpha, results$mse)
```

```{r}
results[which.min(results$mse),]
```

It can be observed from the above outputs that the value of alpha at which we get the best fit and lowest value of MSE is at alpha = 0.3.

--------  DO NOT INCLUDE AFTER THIS LINE ---------

```{r}
library(MASS)
library(caret)
```

```{r}
custom <- trainControl(method = "repeatedcv", number = 10, repeats = 5, verboseIter = TRUE)
```

```{r}
set.seed(1234)
en <- train(x = x.train, y = y.train,
            method='glmnet',
            tuneGrid =expand.grid(alpha=seq(0,1,length=10),
                                  lambda = seq(0.0001,0.2,length=5)),
            trControl=custom)
```

```{r}
en
```

```{r}
mean(en$resample$RMSE)
```

```{r}
plot(en, main = "Elastic Net Regression")
```

```{r}
plot(varImp(en,scale=TRUE))
```

### 2022-10-12 hw7/submission-hw7.Rmd
submission file created

--- 2022-10-12 hw7/abhinav/stepwise.Rmd
@@
   pdf_document: default
-date: "2022-10-11"
+date: "2022-10-12"
 ---
@@
 
+## Part 1 : Stepwise Regression
+
 ```{r, warning=FALSE, message=FALSE}
@@
 library(MASS)
+library(glmnet)
 #install.packages("car")
@@
 Since , the Variance Inflation factor for all the predictor variables in the final output is less than 5 , we can infer that the problem of multicollinearity also does not reside in our final model.
+
+## Part 2: Lasso Regression
+
+```{r}
+crime_df = read.table("./uscrime.txt", sep="\t", header= TRUE)
+dim(crime_df)
+names(crime_df)
+```
+
+```{r}
+x = as.matrix(crime_df[,names(crime_df) != "Crime"])
+y = as.matrix(crime_df[,"Crime"])
+dim(x)
+dim(y)
+```
+
+```{r}
+fit <- glmnet(x, y, alpha = 1, nlambda = 20, standardize = TRUE)
+```
+
+```{r}
+print(fit)
+```
+
+```{r}
+plot(fit, xvar = "lambda", label = TRUE)
+```
+
+```{r}
+cvfit <- cv.glmnet(x = x, y = y, alpha = 1, nlambda = 30, standardize = TRUE, type.measure = "mse", nfolds = 5)
+```
+
+Because we have only 47 points in the data, we don't want fit too many models during the CV stage
+
+```{r}
+plot(cvfit)
+```
+
+Above plot gives CV error against values of lambda. With increasing Lambda the weight of regularization factor increases, and therefore the model is encouraged, to select lower number of predictors.
+
+```{r}
+print (cvfit)
+```
+
+Two special values along the 𝜆 sequence are indicated by the vertical dotted lines.
+`lambda.min` is the value of λ that gives minimum mean cross-validated error, while `lambda.1se` is the value of λ that gives the most regularized model such that the cross-validated error is within one standard error of the minimum.
+
+```{r}
+coef(cvfit, s = "lambda.min")
+```
+
+```{r}
+coef(cvfit, s = "lambda.1se")
+```
+
+```{r}
+cvfit$lambda.1se
+```
+
+```{r}
+fit_1se <- glmnet(x, y, alpha = 1, lambda = cvfit$lambda.1se, standardize = TRUE)
+coef(fit_1se)
+
+fit_min <- glmnet(x, y, alpha = 1, lambda = cvfit$lambda.min, standardize = TRUE)
+coef(fit_min)
+```
+
+```{r}
+predict_min = predict(fit_min, newx = x, s = "lambda.min")
+predict_1se = predict(fit_1se, newx = x, s = "lambda.1se")
+```
+
+```{r}
+rmse_func = function(true,predicted){
+  (mean((true-predicted)^2))**0.5
+}
+```
+
+### RMSE for
+
+```{r}
+rmse_func(true=y,predicted = predict_min)
+```
+
+```{r}
+rmse_func(true=y,predicted = predict_1se)
+```
+
+## Part 3 : ElasticNet Regression
+
+* Elastic net is a combination of both Lasso and Ridge regression. It uses both L1 and L2 penalty to restrict the regression solution.
+* It involves two hyperparameters - Lambda and Alpha.
+* Alpha is used to give direction between Ridge and Lasso. Alpha = 0 is Ridge and Alpha = 1 is Lasso.
+* Lambda is amount of penalty in the regression.
+
+```{r}
+data = read.table("./uscrime.txt", sep="\t", header=TRUE)
+head(data)
+```
+
+```{r}
+x = as.matrix(data[, 1:ncol(data)-1])
+y = data[,"Crime"]
+```
+
+```{r}
+x = scale(x)
+```
+
+### Train Test Split
+
+```{r}
+n=nrow(data)
+train_rows <- sample(1:n, .8*n, replace = F)
+
+x.train <- x[train_rows,]
+y.train <- y[train_rows]
+
+x.test <- x[-train_rows,]
+y.test <- y[-train_rows]
+```
+
+* Splitting the data to get a value of MSE for deciding a value of alpha.
+
+### Estimating Lambda and Alpha - hyperparameters
+
+```{r}
+models <- list()
+for (i in 0:20) {
+  name <- paste0("alpha", i/20)
+
+  models[[name]] <-
+    cv.glmnet(as.matrix(x.train), as.matrix(y.train), alpha=i/20, family="gaussian", nfold = 5)
+}
+```
+
+* We are fixing the value of alpha and using cross validation for find the value of lambda which minimizes the value of MSE - our metric for validation.
+* We find a minimum value of lambda which minimizes the MSE and another value of lambda one standard deviation apart as it is more practical.
+We are accepting some performance loss and incorporating more bias in the model.
+
+```{r}
+results <- data.frame()
+for (i in 0:20) {
+  name <- paste0("alpha", i/20)
+
+  ## Use each model to predict 'y' given the Testing dataset
+  predicted <- predict(models[[name]], 
+                       s=models[[name]]$lambda.1se, newx=x.test)
+
+  ## Calculate the Mean Squared Error...
+  mse <- mean((y.test - predicted)^2)
+
+  ## Store the results
+  temp <- data.frame(alpha=i/20, mse=mse, name=name)
+  results <- rbind(results, temp)
+}
+```
+
+* After we have selected our lambda, we predict on the test data for getting the value of alpha.
+* Alpha is the parameter for deciding between ridge (L2 norm) penalty and lasso (L1 penalty) penalty.
+
+```{r}
+print(results)
+```
+
+```{r}
+plot(results$alpha, results$mse)
+```
+
+```{r}
+results[which.min(results$mse),]
+```
+
+It can be observed from the above outputs that the value of alpha at which we get the best fit and lowest value of MSE is at alpha = 0.3.
+
+--------  DO NOT INCLUDE AFTER THIS LINE ---------
+
+```{r}
+library(MASS)
+library(caret)
+```
+
+```{r}
+custom <- trainControl(method = "repeatedcv", number = 10, repeats = 5, verboseIter = TRUE)
+```
+
+```{r}
+set.seed(1234)
+en <- train(x = x.train, y = y.train,
+            method='glmnet',
+            tuneGrid =expand.grid(alpha=seq(0,1,length=10),
+                                  lambda = seq(0.0001,0.2,length=5)),
+            trControl=custom)
+```
+
+```{r}
+en
+```
+
+```{r}
+mean(en$resample$RMSE)
+```
+
+```{r}
+plot(en, main = "Elastic Net Regression")
+```
+
+```{r}
+plot(varImp(en,scale=TRUE))
+```

### 2022-10-19 hw7/submission-hw7.Rmd
backlog

@@
 Another point to note is that the Multiple R-squared of the final model is 0.789 which is lower than the initial model which has a value of 0.8031 but this is the pitfall of R squared value i.e.
-It will always increases with the increase in the # of variables without any penalty/adjustment for model complexity/overfitting unlike AIC or Adjusted R-squared.
+It will always increases with the increase in the \# of variables without any penalty/adjustment for model complexity/overfitting unlike AIC or Adjusted R-squared.
 
@@
 
-Based on the above formula , we can infer that the Adjusted R squared value decreases with the increase in # of independent variables(k) until and unless if there is a significant increase in the R squared value on adding a new variable.
+Based on the above formula , we can infer that the Adjusted R squared value decreases with the increase in \# of independent variables(k) until and unless if there is a significant increase in the R squared value on adding a new variable.
 Thus both AIC and Adjusted R squared provide a technique for selecting simpler model which prevents overfitting.
@@
 ```
+
+Lasso Regression is fit with glmnet function. Here we have two parameters:
+
+-    alpha = 1 gives 1 weight to Lasso and 0 to Ridge
+
+-   lambda is another parameter is the weight for regularization in cost function
+
+Note the parameter standardizes the data, which is a requirement for Lasso
 
@@
 
-Two special values along the 𝜆 sequence are indicated by the vertical dotted lines.
+Two special values along the λ sequence are indicated by the vertical dotted lines.
 `lambda.min` is the value of λ that gives minimum mean cross-validated error, while `lambda.1se` is the value of λ that gives the most regularized model such that the cross-validated error is within one standard error of the minimum.
@@
 
+Let us fit full models for these parameter values. While ideally we should pick model from CV stage, just to understand the functioning better we are selecting two models here.
+
+The nuance here is that while lambda.min minimizes CV error, We employ lasso for feature selection and hence one might argue lambda.1se is the parameter we should choose, since it minimizes the number of features selected.
+
 ```{r}
@@
 ```
+
+We also note here that the variables being dropped are similar to what we hypothesized in previous iterations of this exercise, say in RF and Linear Regression.
+With highly correlated features being dropped.
 
@@
 
-### RMSE for
+### RMSE for above two models
 
@@
 
-* Elastic net is a combination of both Lasso and Ridge regression. It uses both L1 and L2 penalty to restrict the regression solution.
-* It involves two hyperparameters - Lambda and Alpha.
-* Alpha is used to give direction between Ridge and Lasso. Alpha = 0 is Ridge and Alpha = 1 is Lasso.
-* Lambda is amount of penalty in the regression.
+-   Elastic net is a combination of both Lasso and Ridge regression. It uses both L1 and L2 penalty to restrict the regression solution.
+-   It involves two hyperparameters - Lambda and Alpha.
+-   Alpha is used to give direction between Ridge and Lasso. Alpha = 0 is Ridge and Alpha = 1 is Lasso.
+-   Lambda is amount of penalty in the regression.
 
@@
 
-* Splitting the data to get a value of MSE for deciding a value of alpha.
+-   Splitting the data to get a value of MSE for deciding a value of alpha.
 
@@
 
-* We are fixing the value of alpha and using cross validation for find the value of lambda which minimizes the value of MSE - our metric for validation.
-* We find a minimum value of lambda which minimizes the MSE and another value of lambda one standard deviation apart as it is more practical.
+-   We are fixing the value of alpha and using cross validation for find the value of lambda which minimizes the value of MSE - our metric for validation.
+-   We find a minimum value of lambda which minimizes the MSE and another value of lambda one standard deviation apart as it is more practical.
 We are accepting some performance loss and incorporating more bias in the model.
@@
 
-* After we have selected our lambda, we predict on the test data for getting the value of alpha.
-* Alpha is the parameter for deciding between ridge (L2 norm) penalty and lasso (L1 penalty) penalty.
+-   After we have selected our lambda, we predict on the test data for getting the value of alpha.
+-   Alpha is the parameter for deciding between ridge (L2 norm) penalty and lasso (L1 penalty) penalty.
 
@@
 It can be observed from the above outputs that the value of alpha at which we get the best fit and lowest value of MSE is at alpha = 0.3.
-
---------  DO NOT INCLUDE AFTER THIS LINE ---------
-
-```{r}
-library(MASS)
-library(caret)
-```
-
-```{r}
-custom <- trainControl(method = "repeatedcv", number = 10, repeats = 5, verboseIter = TRUE)
-```
-
-```{r}
-set.seed(1234)
-en <- train(x = x.train, y = y.train,
-            method='glmnet',
-            tuneGrid =expand.grid(alpha=seq(0,1,length=10),
-                                  lambda = seq(0.0001,0.2,length=5)),
-            trControl=custom)
-```
-
-```{r}
-en
-```
-
-```{r}
-mean(en$resample$RMSE)
-```
-
-```{r}
-plot(en, main = "Elastic Net Regression")
-```
-
-```{r}
-plot(varImp(en,scale=TRUE))
-```

### 2022-10-19 hw8-1/ashish_ques12.2.Rmd
backlog

---
title: "ad_ques12.2"
author: "Ashish Dhiman"
date: "2022-10-19"
output: html_document
---

```{r}
library(FrF2)
```

## Question 12.2

We want to design a experiment such that we can find the effect of 10 Binary features on the price of a house. Some of the sample features could be:

1.  House has more than 4 bedrooms
2.  Park nearby
3.  School nearby
4.  If Pets are allowed in Locality, etc

With the full factorial design, we would land with $$2^{10} = 1024$$ combinations, which is a very large number.

Hence we want to design a fractional factorial design experiment here, with 16 combinations.

```{r}
n_combinations = 16
n_variables = 10
```

```{r}
factorial_doe = FrF2(nruns = n_combinations,nfactors =  n_variables)
factorial_doe
```

### Properties of Factorial Design

#### Each choice should be checked same number of times

```{r}
table(factorial_doe$A)
```

```{r}
print (paste("Shape of original df: ",dim(factorial_doe)[1]))
print (paste("Shape of de-deuplicated df: ",dim(unique(as.data.frame(factorial_doe)))[1]))
```

The above test shows that the table is unique without any duplicates. And, since A is tested only 8 times, the information can be combined to assert the above test.

#### All pair of choices should be chosen the same number of times

```{r}
table(factorial_doe$A,factorial_doe$B)
```

```{r}
table(factorial_doe$A,factorial_doe$D)
```

Above Behavior can again be generalized to prove the second assumption.

### 2022-10-19 hw8-1/ashish_ques12.2.Rmd
completed ques12.2

@@
 ```{r}
-factorial_doe = FrF2(nruns = n_combinations,nfactors =  n_variables)
+factorial_doe = FrF2(nruns = n_combinations,nfactors =  n_variables,
+                     default.levels = c(0,1))
 factorial_doe
 ```
+
+Note here 1 means binary variable present, and 0 means otherwise
+
+```{r}
+summary(factorial_doe)
+```
+
+The #generators =10-4 = 6. B'cos 16 = 2\^4.
 
@@
 Above Behavior can again be generalized to prove the second assumption.
+
+### Feature distribution
+
+```{r}
+factorial_doe1 = as.data.frame(factorial_doe)
+```
+
+```{r}
+factorial_doe1$sum = as.numeric(factorial_doe1$A)+as.numeric(factorial_doe1$B)+as.numeric(factorial_doe1$C)+as.numeric(factorial_doe1$D)+as.numeric(factorial_doe1$E)+as.numeric(factorial_doe1$F)+as.numeric(factorial_doe1$G)+as.numeric(factorial_doe1$H)+as.numeric(factorial_doe1$J)+as.numeric(factorial_doe1$K)-10
+table(factorial_doe1$sum)
+```

### 2022-10-19 hw8-1/Submission.Rmd
added hw8 submission.rmd

--- 2022-10-19 hw8-1/ashish_ques12.2.Rmd
@@
 ---
-title: "ad_ques12.2"
-author: "Ashish Dhiman"
-date: "2022-10-19"
-output: html_document
+title: "ISYE 6501"
+output:
+  html_document:
+    df_print: paged
+  pdf_document: default
+date: "2022-09-12"
 ---
 
-```{r}
+# Submission HW8 \| Fall 22
+
+-   Ashish Dhiman \| [ashish.dhiman\@gatech.edu](mailto:ashish.dhiman@gatech.edu){.email}
+
+-   Abhinav Arun \| [aarun60\@gatech.edu](mailto:aarun60gatech.edu){.email}
+
+-   Anshit Verma \| [averma373\@gatech.edu](mailto:averma373@gatech.edu){.email}
+
+## Question 12.1
+
+## Question 12.2
+
+```{r, include=FALSE}
 library(FrF2)
 ```
-
-## Question 12.2
 
@@
 ```
+
+## Question 13.1
+
+a.	**Binomial** - In finding defective items, the outcome could be either success (item is defective) or failure (item is non-defective).
+b.	**Geometric** - For example, you ask people outside a polling station who they voted for until you find someone that voted for the independent candidate in a local election.
+The geometric distribution would represent the number of people who you had to poll before you found someone who voted independent.
+You would need to get a certain number of failures before you got your first success.
+c.	**Poisson** - It gives the probability of an event happening a certain number of times (k) within a given interval of time or space. One example is no. of website visitors per month.
+d.	**Exponential** - One example is the amount of time, in months, a car battery lasts.
+e.	**Weibull** - One example is Warranty analysis. We are using data to model time-to-failure.

### 2022-10-19 hw8-1/Submission.Rmd
hw8: ques 12.1
Part 1 added

@@
 ## Question 12.1
+
+Design of Experiments finds its use cases in various domains :
+
+One particular area where it is widely used is in :
+
+***Website design prototyping and Call to Action (CTA) Optimization***
+
+We can use it to evaluate the impact of adding a new functionality on a website to evaluate user interaction/user traffic on a website :
+
+There are various aspects which we could experiment on :
+
+* Location of the new feature on the website : 4 possible categories (Top Left , Top Right , Bottom Left , Bottom Right) to evaluate whether the new feature location would increase user engagement.
+
+* Color of the feature added : 3 possible categories (red,blue,green) - to see if a the color of the feature affects user interaction
+
+* Feature mobility: 2 categories (dynamic, static) - whether the feature should be static or dynamic (can help us evaluate if people are liking simplistic feature or they like more interactive/dynamic features).
+
+* Having couple of icons on the new feature : One good use case of this is when creating a "Pay Now" icon on the payment gateway website .
+Should we add an additional icon of "Return Back" adjacent to Pay now button because it may so happen that many people might not go through payment transactions and as a result of it the company's profitability might get impacted.So it's better to test out it's impact with some real data (Yes/No option , so 2 possible categories).
+
+* Size of the feature added : 3 possible categories (Small, Medium, Large) and it's impact on user liking/interaction.
+
+Thus , we will have a total of $4\cdot 3\cdot 2\cdot 2 \cdot 3$ = 144 combinations to test out .
+In order to get a better sampling distribution across each combination , we can think on reducing the total number of combinations using various factorial design techniques so that we do not have very less data corresponding to a possible combination and so that we could make some statistically significant inference about the user behavior.
 

### 2022-10-26 hw8-1/Submission.Rmd
final hw8

@@
 
-a.	**Binomial** - In finding defective items, the outcome could be either success (item is defective) or failure (item is non-defective).
-b.	**Geometric** - For example, you ask people outside a polling station who they voted for until you find someone that voted for the independent candidate in a local election.
+Examples of data for the following distributions:
+
+a.
+**Binomial** - In finding defective items, the outcome could be either success (item is defective) or failure (item is non-defective).
+If you are doing this for multiple items and each item being selected is at random i.e.
+iid, then this example follows a Binomial distribution.
+b.
+**Geometric** - For example, you ask people outside a polling station who they voted for until you find someone that voted for the independent candidate in a local election.
 The geometric distribution would represent the number of people who you had to poll before you found someone who voted independent.
 You would need to get a certain number of failures before you got your first success.
-c.	**Poisson** - It gives the probability of an event happening a certain number of times (k) within a given interval of time or space. One example is no. of website visitors per month.
-d.	**Exponential** - One example is the amount of time, in months, a car battery lasts.
-e.	**Weibull** - One example is Warranty analysis. We are using data to model time-to-failure.
+This example follows a Geometric distribution.
+c.
+**Poisson** - It gives the probability of an event happening a certain number of times (k) within a given interval of time or space.
+One example is no.
+of website visitors per month.
+We are using the arrivals of vistors to a website on a per month basis.
+This follows a Poisson distribution.
+d. **Exponential** - One example is the amount of time until an earthquake occurs. The time interval between earthquake follows an exponential distribution.
+e.
+**Weibull** - One example is lifetime of dental implants.
+We are using data to model time-to-failure.
+If you get new dental implants, after how long will the implants breaks.
+This follows a Weibull distribution.

### 2022-11-28 hw13/homework_13.Rmd
hw13 submission

---
title: "Homework_13"
output: html_document
date: "2022-11-28"
---

## Shelf Space Optimization - Retail Industry - Recommendations ##

The shelf space optimization problem is widely used by various retail chains to streamline revenue growth and to enhance sales velocity for various categories of products.

Data sources that could be useful :

Computing the % share of shelf space . It could be quanitified in a variety of ways. 
-> Linear Occupancy (shelf space) -> % of physical space occupied by selected products as a % of total # products.

-> Eye-Level Facing -> How many facings of a product are placed at the eye level (buy level) as compared to the total # of products of that same category.

Key variables impacting objective function : 
-> Sales Price of products
-> Replenishing rate
-> Location of products (Vertical, Horizontal positioning)
-> Size/Area of products
-> Sizes/Areas of placement shelves (Vertical, Horizontal – important for space optimization)
-> Average amount of products sold in a period (over time)
-> Logistics costs (shipment to store)
-> Historical purchase data

A wide range of techniques are being currently deployed by various retail chains in order to counter the problem of shelf space optimization but such practices need to be evaluated properly before rolling them into production and that is where a lot of thinking and planning is needed to avoid any pitfalls and successfully parse through caveats.

-\> Increasing the shelf area assigned to a given product and then evaluating its effect on the sales of that product over a period of time. But it is very tricky to evaluate the effect in an unbiased manner as there are various controlling factors that we need to account for (like seasonality , global macroeconomic events like supply side constraints, economic headwinds to counter inflation , global pandemic like covid etc.).

For e.g. Suppose we are evaluating the effect of increased shelf size on the sales of products like icecream, cold drinks during the time when summer season is about to kick in , then our analysis might not be a true reflection of the actual impact of increased shelf size . So , it is imperative that we account for seasonality/demand trends. Also , we need to control for bias inherent in human behavior , for e.g. as per various research studies , products that are placed at eye level and those that are placed in clockwise direction(customer flow through stores) have comparatively higher sales velocity. Therefore , these factors must be accounted for when analyzing the impact of increased shelf size alone on sales growth.

Moreover , the days since the launch of the product, brand of the product and product type also need to be controlled for in order to have an unbiased evaluation.

We can use A/B testing to evaluate efficacy of increasing shelf size. But we need to be cognizant of the exploration vs exploitation conundrum here as one retail chain might not want to experiment on a product at a time when it's demand and supply is highly fluctuating , also the logistic and effort for replenishment aspect of things needs to be handled as increasing the shelf size of the product whose supply is highly constrained might not be a good idea as the same space can be utilized for some other in-demand product with slightly more robust supply chain . Bigger retail stores have an edge here over the newer stores as they have specialized supply chain and inventory management teams to look into these things and model accordingly.

Also , the evaluation must be done in a similar store location .

We can build a regression model to model the effect of all these features on the increase in sales over a period of time . Thus , it can be modeled as a combination of time series and regression model.

-\> Analyzing the interaction affects- whether placing 2 complementary products together has an influence on the sales of each product. Various data mining algorithms used extensively in market basket analysis can be used here. Association Rule Mining techniques like Apriori algorithm can be used to evaluate the efficacy of interaction effect. Defining appropriate threshold support and confidence helps in clearly finding those products which greatly influence each other's sales.

Analysis techniques to be used given that we have the requisite data :

1. Time Series Analysis : We can use exponential smoothing to model the sales forecast of a product. It would help us account for both seasonality and trend. The time series technique would also help in inventory management and division of shelf space among the product types.

2. Optimization models : linear/integer programming. Optimizing the overall store space because if we start reaching saturation in sales with 80% of the total store space occupied , then it doesn't make sense to stack up the entire store as it would lead to inventory getting piled up.
Objective : Maximize Sales/ Revenue
Constraints to be considered : 
-> minimum amount of shelf space required
-> maximum amount of shelf space that can be devoted
-> total amount of shelf space that has to be used

3. Data Mining Techniques: Apriori , FP Tree Algorithm -> Used to find hidden interactive patterns in the underlying data.

4. Clustering technique : To create clusters of similar products (help us in product placement).

5. Genetic Algorithm

### 2022-11-29 project/submission_project.Rmd
project report started

---
title: "ISYE 6501"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2022-11-28"
urlcolor: blue
---

# Project Submission \| Fall 22

-   Ashish Dhiman \| [ashish.dhiman\@gatech.edu](mailto:ashish.dhiman@gatech.edu){.email}

-   Abhinav Arun \| [aarun60\@gatech.edu](mailto:aarun60gatech.edu){.email}

-   Anshit Verma \| [averma373\@gatech.edu](mailto:averma373@gatech.edu){.email}

#### Table of Contents

-   [Summary](#Summary)
-   [Backfround on Bharat Pe](#background)
-   [Analytics Usecases](#usecases)
-   [Usecase 1: Lending Activity](#usecase1)
-   [Usecase 2:](#usecase2)
-   [Usecase 3:](#usecase3)

## Summary {#summary}

BharatPe is an Indian fin-tech company operating mainly in the payments space and targeting Small Business Owners(SBOs). With the digital transformation of Indian economy in the recent years, the scale of digital payments has touched \~6M monthly transactions. BharatPe was one of the key players to ride this wave of digital (UPI) payments.

Access to this rich payments data lends itself to a variety of analytics use-cases, and as part of this course project, we try to discuss our ideas on some of those use-cases.

## Backfround on BharatPe and UPI {#background}

### BharatPe

[BharatPe](https://bharatpe.com/companyprofile)

### UPI

[UPI](https://www.npci.org.in/what-we-do/upi/product-overview)

monthly transactions. BharatPe was one of the key players to ride this wave of digital (UPI) payments.

Access to this rich payments data lends itself to a variety of analytics use-cases, and as part of this course project, we try to discuss our ideas on some of those use-cases.

monthly transactions. BharatPe was one of the key players to ride this wave of digital (UPI) payments.

Access to this rich payments data lends itself to a variety of analytics use-cases, and as part of this course project, we try to discuss our ideas on some of those use-cases.

### Analytics Usecases {#usecases}

## Usecase 1: Lending Activity {#usecase1}

One of the core competitive advantages of BharatPe at its inception was that it created a multiplatform UPI (QR code), which could be used with any 3rd party app to accept payments by the SBO. This greatly accelerated the on-boarding of SBOs to BharatPe, who were otherwise under serviced by the traditional lending players like Banks and NBFCs on account of high risk.

With access to payments made to the SBO, now we can reasonably approximate the revenue of the SBO, and therefore its credit worthiness. There are two parts to this problem as given in the schematic below:

### Classification Model: To Lend or not ?

### Regression Model: How much to lend ?

## Usecase 2: Personalisation Services

### 2022-12-05 project/submission_project.Rmd
project report v0
hw4

@@
 -   [Summary](#Summary)
--   [Backfround on Bharat Pe](#background)
+-   [Backfround on Zepto](#background)
 -   [Analytics Usecases](#usecases)
@@
 
-## Backfround on BharatPe and UPI {#background}
+## Backfround on Zepto {#background}
 
-### BharatPe
+## Problem Definition {#usecases}
 
-[BharatPe](https://bharatpe.com/companyprofile)
-
-### UPI
-
-[UPI](https://www.npci.org.in/what-we-do/upi/product-overview)
-
-monthly transactions. BharatPe was one of the key players to ride this wave of digital (UPI) payments.
-
-Access to this rich payments data lends itself to a variety of analytics use-cases, and as part of this course project, we try to discuss our ideas on some of those use-cases.
-
-monthly transactions. BharatPe was one of the key players to ride this wave of digital (UPI) payments.
-
-Access to this rich payments data lends itself to a variety of analytics use-cases, and as part of this course project, we try to discuss our ideas on some of those use-cases.
-
-### Analytics Usecases {#usecases}
-
-## Usecase 1: Lending Activity {#usecase1}
+## Modelling Methodology {#usecase1}
 
@@
 
-### Classification Model: To Lend or not ?
+```{r,echo=FALSE, out.width = "80%", fig.align = "center"}
+knitr::include_graphics("./6501_usecase1.drawio.png")
+```
 
-### Regression Model: How much to lend ?
-
-## Usecase 2: Personalisation Services
+###

### 2022-12-05 hw14/submission.Rmd
hw14

---
title: "hw14"
author: "Ashish Dhiman"
date: "2022-12-05"
output:
  pdf_document: default
  html_document: default
---

# HW 14 Submission \| Fall 22

-   Ashish Dhiman \| [ashish.dhiman\@gatech.edu](mailto:ashish.dhiman@gatech.edu){.email}

-   Abhinav Arun \| [aarun60\@gatech.edu](mailto:aarun60gatech.edu){.email}

-   Anshit Verma \| [averma373\@gatech.edu](mailto:averma373@gatech.edu){.email}

##

### Use-case 1: Recommend Repeat Purchase Products (Only Data Set 3)

In this case we wan to build a model to recommend repeat purchases by a user, say Shampoo etc. The hypothesis here is that there exists certain type of products which a customer might purchase/consume repeatedly. In such a scenario we can use analytics to reccommend the user to a higher end product in the same category. Say a user buys a shampoo costing 10USD every other month, then we can recommend him a shampoo costing 20 USD.

**So Given Data (X) on:**

-   User

-   Product

-   Time spent on the product page the first time

-   Time user spends on the product page (if he spends less time, might show repeat purchase)

-   ratio of above two

-   Purchase History:

    -   Times bought in last n days , with variable n

    -   time since last purchase

    -   (days since purchase(k) - days since purchase(k+1))(days since purchase(k+1) - days since purchase(k+2)), with variable k

**We can model:**

Y: If particular user product is a repeat combination.

Again this will be a classification model and training data can be created with the help of rules capturing above information.

**This will result in:**

In a probability score basis which we can decide if the user, product combination is a repeat purchase. If so, we can calculate his average repeat frequency as purchases in last year/1365 days. Using these two pieces of information we can reccomend him higher end products in the same category.

### Use-case 2: Buy Now Pay Later(BNPL)

With this case we want to use dataset 2 and dataset 3 to, predict if the merchant should offer services like Buy Now Pay later (or EMI) to the customer. Buy Now pay later (or Monthly installment) is an attractive manner for the customer to buy stuff out of his reach instantly and if it does not bring additional risk, it's hugely helpful in increasing the revenue.

**So Given Data (X) on:**

-   Credit History: Low risk ==\> More BNPL

-   Balance on Credit Cards ==\> Should not have very high balance

-   Credit Limit on Credit Cards ==\> Low credit limit would be more likely to access BNPL

-   Monthly Commitments (Utilities Rent etc):

-   Time spent on a product page for products not bought (Might measure want of a product)

**We can model:**

Y: If present options like BNPL to the user, translates to product purchase AND customer pays back money

This will be a classification model. We want the model to predict not only if showing BNPL option transtates to a purchase, but also if the customer pays back the amount. To train this model we will first need to create the training data using experiments (possibly with A/B Testing). We can do a test where we randomly present BNPL options to some people vs a control group. If the test shows BNPL results in purchase, we can track them to see who paid back and thus create our final Y label. With this we can train our classification model.

**This will result in:**

In a probability score basis which we can decide if we should present BNPL options to a certain customer.
